---
title: "Calibrating SSC de novo data with Roulette (from Trost Cell 2022)"
author: "XSun"
date: "2023-07-17"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(data.table)
library(ggplot2)
```


Data source: 

[https://www.sciencedirect.com/science/article/pii/S0092867422013241?via=ihub#abs0020 [Cell 2022]](https://www.sciencedirect.com/science/article/abs/pii/S0092867422013241?via=ihub)


# Data overview

 Simons Simplex Collection (SSC; n = 9,205)
 
 326,669 de novo SNVs.



| mutation types | # of de nove SNVs (total =  326,669)        |
|----------------|-----------------------|
| a->c t->g      | 12253 + 12166 = 22419   |
| a->g t->c      | 44124 + 43634 = 87758 |
| a->t t->a      | 11257 + 11605 = 22862   |
| c->a g->t      | 15885 + 16206 = 32091 |
| c->g g->c      | 15834 + 15781 = 31615 |
| c->t g->a      | 68209 + 68512 = 136721 |


```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=5}

data <- data.frame(
  category = c("a->c", "a->g", "a->t", "c->a", "c->g", "c->t"),
  percentage = c(7.48,26.86,7.00, 9.82, 9.68,41.85)
)
data$angle <- cumsum(data$percentage) - (0.5 * data$percentage)
ggplot(data, aes(x = "", y = percentage, fill = category)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  theme_void() +
  theme(legend.position = "bottom") +
  # Adjust the position of the labels
  geom_label(
    aes(label = paste0(category, "\n", percentage, "%")),
    position = position_stack(vjust = 0.5),
    size = 3.5,
    show.legend = FALSE
  )
```


# Rescaling 

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=5}

load("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/results/scaling_factor.rdata")


print("fitting: observed (SSC) ~ expected (Roulette) + 0")
fit <- lm(denovo_per_chr ~ rl_per_chr + 0)
print(summary(fit))

data <- data.frame(expected = rl_per_chr, observed = denovo_per_chr)
print(data)

# Create the plot
ggplot(data, aes(x = expected, y = observed)) +
  geom_point() +
  geom_abline(slope = fit$coefficients[1], intercept = 0, color = "red") +
  xlab("Expected_Roulette") +
  ylab("Observed")

print(paste0("the scalling factor is ",round(fit$coefficients[1],digits = 5)))
print(paste0("obs = ",round(fit$coefficients[1],digits = 5),"* exp"))
```


# Random effect distribution (after rescaling)

## Window size: 10kb

### chr1

All 6 mutation types, random effect: obs/exp

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

data <- fread("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr1/rl_ssc_trost_combine_chr1_10000.txt")
colnames(data) <- c("chr","start","end","obs","exp")
data$Mb <- data$start/1000000
data$randeff <- data$obs / data$exp

print("the number of observed de novo mutations per window")
table(data$obs)

ggplot(data, aes(x = Mb, y = obs)) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of observed de novo mutations") +
  ggtitle("Distribution of observed de novo mutations") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


ggplot(data, aes(x = Mb, y = randeff)) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("Random effects") +
  ggtitle("Distribution of random effects") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


```

### chr22

All 6 mutation types, random effect: obs/exp

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

data <- fread("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr22/rl_ssc_trost_combine_chr22_10000.txt")
colnames(data) <- c("chr","start","end","obs","exp")
data$Mb <- data$start/1000000
data$randeff <- data$obs / data$exp

print("the number of observed de novo mutations per window")
table(data$obs)

ggplot(data, aes(x = Mb, y = obs)) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of observed de novo mutations") +
  ggtitle("Distribution of observed de novo mutations") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


ggplot(data, aes(x = Mb, y = randeff)) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("Random effects") +
  ggtitle("Distribution of random effects") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


```

## Window size: 50kb

### chr1

All 6 mutation types, random effect: obs/exp

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

data <- fread("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr1/rl_ssc_trost_combine_chr1_50000.txt")
colnames(data) <- c("chr","start","end","obs","exp")
data$Mb <- data$start/1000000
data$randeff <- data$obs / data$exp

print("the number of observed de novo mutations per window")
table(data$obs)

ggplot(data, aes(x = Mb, y = obs)) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of observed de novo mutations") +
  ggtitle("Distribution of observed de novo mutations") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


ggplot(data, aes(x = Mb, y = randeff)) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("Random effects") +
  ggtitle("Distribution of random effects") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


```

### chr22

All 6 mutation types, random effect: obs/exp

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

data <- fread("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr22/rl_ssc_trost_combine_chr22_50000.txt")
colnames(data) <- c("chr","start","end","obs","exp")
data$Mb <- data$start/1000000
data$randeff <- data$obs / data$exp

print("the number of observed de novo mutations per window")
table(data$obs)

ggplot(data, aes(x = Mb, y = obs)) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of observed de novo mutations") +
  ggtitle("Distribution of observed de novo mutations") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


ggplot(data, aes(x = Mb, y = randeff)) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("Random effects") +
  ggtitle("Distribution of random effects") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


```
