---
title: "Simulation for non-spatial random effects"
author: "XSun"
date: "2024-04-22"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
---

# Introduction

## Modeling the non-spatial random effects with gamma distribution

The model is described [here](https://xsun1229.github.io/mutation_rate/denovo_0422_models.html)

## Simulation setups

1. Use Gamma prior to sample $\tilde \theta_i$ in the simulation: $$\tilde \theta_{i} \sim Gamma(\alpha,\alpha)$$  for all windows. The prior $\alpha \in \{5,10,30,50\}$

```{r}
randeff_simulated <- rgamma(n = 3000, shape = 3, rate =3)
plot(randeff_simulated)
print(sprintf("mean value of randeff_simulated = %s",mean(randeff_simulated)))
```


2. Sampling the observed mutation count $y_i$ From Poisson Distribution $y_i \sim \text{pois}(\mu_i\tilde\theta_i)$, $\mu_i$ is the Roulette estimated mutation rate

3. Estimating $\hat{\alpha}$ using MLE Simplex (Nelder-Mead) method (see the details above)

4. Computing the estimated $\hat{\theta_i}$ from $$\dfrac{y_i+\hat{\alpha}}{\mu_i+\hat{\alpha}} $$


## Functions used

```{r} 
library(ggplot2)
library(gridExtra)
plot_randeff <- function(data) {
  
  
  ggplot(data, aes(x = x)) +
    geom_line(aes(y = randeff_simulated, color = "Simulated"), alpha = 0.5) +
    geom_line(aes(y = randeff_est, color = "Estimated"), alpha = 0.7) +
    labs(x = "Window index",
         y = "random effects",
         color = "Type") +
    scale_color_manual(values = c("Simulated" = "blue", "Estimated" = "red")) +
    theme_minimal()
}

scatter_plot_randeff <- function(data) {
  # Fit the linear model
  fit <- lm(randeff_simulated ~ randeff_est + 0, data = data)
  adj_rsq <- summary(fit)$adj.r.squared  # Extract the adjusted R-squared
  
  # Generate the plot
  ggplot(data, aes(x = randeff_est, y = randeff_simulated)) +
    geom_point(color = "black") +
    geom_abline(intercept = 0, slope = 1, color = "red") +
    geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
              x = Inf, y = -Inf, hjust = 1, vjust = 0,
              color = "blue") +
    labs(x = "Estimated random effect",
         y = "True random effect theta_i") +
    theme_minimal()
}

plot_dnm <- function(data, windowsize) {
   ggplot(data, aes(x = x)) +
    geom_point(aes(y = y_sample), color = "grey") +
    geom_line(aes(y = randeff_simulated), color = "blue") +
    geom_line(aes(y = randeff_est), color = "red", alpha = 0.5) +
    labs(x = "Window index",
         y = paste0("# of DNM/window (windowsize = ", as.numeric(windowsize) / 1000, "kb)")) +
    theme_minimal()
}

scatter_plot_mutct <- function(data, xcol, ycol) {
  # Fit the linear model using user-defined columns
  formula <- as.formula(paste(ycol, "~", xcol,"+0"))
  fit <- lm(formula, data = data)
  adj_rsq <- summary(fit)$adj.r.squared  # Extract the adjusted R-squared
  
  # Generate the plot with user-defined columns
  ggplot(data, aes_string(x = xcol, y = ycol)) +
    geom_point(color = "black") +
    geom_abline(intercept = 0, slope = 1, color = "red") +
    geom_text(aes(label = paste0("Adj R-sq = ", round(adj_rsq, 3))),
              x = Inf, y = Inf, hjust = 1.1, vjust = 1.1,
              color = "blue", parse = FALSE) +  # Ensure label is treated as text
    labs(x = paste("Predicted y:", xcol),
         y = paste("True y (sample):", ycol)) +
    theme_minimal()
}

```


# Results

The estimated mutation rates are from Roulette chr1 

## 10 kb

### Gamma prior $\alpha=5$

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/non_spatial/simu_nonspatial_chr1_10kb_gammaprior5.rdata")

print(sprintf("true alpha = %s", alpha_prior))
print(sprintf("estimated alpha = %s", alpha_est))

plot_randeff(data)
plot_dnm(data,windowsize)
scatter_plot_randeff(data)

data$y_pred <- data$randeff_est * data$exp_rl
p1 <- scatter_plot_mutct(data = data, xcol = "y_pred", ycol = "y_sample")
p2 <- scatter_plot_mutct(data = data, xcol ="exp_rl" , ycol = "y_sample")
grid.arrange(p1, p2, ncol = 2)
```


### Gamma prior $\alpha=10$

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/non_spatial/simu_nonspatial_chr1_10kb_gammaprior10.rdata")

print(sprintf("true alpha = %s", alpha_prior))
print(sprintf("estimated alpha = %s", alpha_est))
plot_randeff(data)
plot_dnm(data,windowsize)
scatter_plot_randeff(data)

data$y_pred <- data$randeff_est * data$exp_rl
p1 <- scatter_plot_mutct(data = data, xcol = "y_pred", ycol = "y_sample")
p2 <- scatter_plot_mutct(data = data, xcol ="exp_rl" , ycol = "y_sample")
grid.arrange(p1, p2, ncol = 2)
```


### Gamma prior $\alpha=30$

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/non_spatial/simu_nonspatial_chr1_10kb_gammaprior30.rdata")

print(sprintf("true alpha = %s", alpha_prior))
print(sprintf("estimated alpha = %s", alpha_est))
plot_randeff(data)
plot_dnm(data,windowsize)
scatter_plot_randeff(data)

data$y_pred <- data$randeff_est * data$exp_rl
p1 <- scatter_plot_mutct(data = data, xcol = "y_pred", ycol = "y_sample")
p2 <- scatter_plot_mutct(data = data, xcol ="exp_rl" , ycol = "y_sample")
grid.arrange(p1, p2, ncol = 2)
```


### Gamma prior $\alpha=50$

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/non_spatial/simu_nonspatial_chr1_10kb_gammaprior50.rdata")

print(sprintf("true alpha = %s", alpha_prior))
print(sprintf("estimated alpha = %s", alpha_est))
plot_randeff(data)
plot_dnm(data,windowsize)
scatter_plot_randeff(data)

data$y_pred <- data$randeff_est * data$exp_rl
p1 <- scatter_plot_mutct(data = data, xcol = "y_pred", ycol = "y_sample")
p2 <- scatter_plot_mutct(data = data, xcol ="exp_rl" , ycol = "y_sample")
grid.arrange(p1, p2, ncol = 2)
```

## 30 kb

### Gamma prior $\alpha=5$

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/non_spatial/simu_nonspatial_chr1_30kb_gammaprior5.rdata")

print(sprintf("true alpha = %s", alpha_prior))
print(sprintf("estimated alpha = %s", alpha_est))
plot_randeff(data)
plot_dnm(data,windowsize)
scatter_plot_randeff(data)

data$y_pred <- data$randeff_est * data$exp_rl
p1 <- scatter_plot_mutct(data = data, xcol = "y_pred", ycol = "y_sample")
p2 <- scatter_plot_mutct(data = data, xcol ="exp_rl" , ycol = "y_sample")
grid.arrange(p1, p2, ncol = 2)
```


### Gamma prior $\alpha=10$

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/non_spatial/simu_nonspatial_chr1_30kb_gammaprior10.rdata")

print(sprintf("true alpha = %s", alpha_prior))
print(sprintf("estimated alpha = %s", alpha_est))
plot_randeff(data)
plot_dnm(data,windowsize)
scatter_plot_randeff(data)

data$y_pred <- data$randeff_est * data$exp_rl
p1 <- scatter_plot_mutct(data = data, xcol = "y_pred", ycol = "y_sample")
p2 <- scatter_plot_mutct(data = data, xcol ="exp_rl" , ycol = "y_sample")
grid.arrange(p1, p2, ncol = 2)
```


### Gamma prior $\alpha=30$

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/non_spatial/simu_nonspatial_chr1_30kb_gammaprior30.rdata")

print(sprintf("true alpha = %s", alpha_prior))
print(sprintf("estimated alpha = %s", alpha_est))
plot_randeff(data)
plot_dnm(data,windowsize)
scatter_plot_randeff(data)

data$y_pred <- data$randeff_est * data$exp_rl
p1 <- scatter_plot_mutct(data = data, xcol = "y_pred", ycol = "y_sample")
p2 <- scatter_plot_mutct(data = data, xcol ="exp_rl" , ycol = "y_sample")
grid.arrange(p1, p2, ncol = 2)
```


### Gamma prior $\alpha=50$

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/non_spatial/simu_nonspatial_chr1_30kb_gammaprior50.rdata")

print(sprintf("true alpha = %s", alpha_prior))
print(sprintf("estimated alpha = %s", alpha_est))
plot_randeff(data)
plot_dnm(data,windowsize)
scatter_plot_randeff(data)

data$y_pred <- data$randeff_est * data$exp_rl
p1 <- scatter_plot_mutct(data = data, xcol = "y_pred", ycol = "y_sample")
p2 <- scatter_plot_mutct(data = data, xcol ="exp_rl" , ycol = "y_sample")
grid.arrange(p1, p2, ncol = 2)
```

## 50 kb

### Gamma prior $\alpha=5$

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/non_spatial/simu_nonspatial_chr1_50kb_gammaprior5.rdata")

print(sprintf("true alpha = %s", alpha_prior))
print(sprintf("estimated alpha = %s", alpha_est))
plot_randeff(data)
plot_dnm(data,windowsize)
scatter_plot_randeff(data)

data$y_pred <- data$randeff_est * data$exp_rl
p1 <- scatter_plot_mutct(data = data, xcol = "y_pred", ycol = "y_sample")
p2 <- scatter_plot_mutct(data = data, xcol ="exp_rl" , ycol = "y_sample")
grid.arrange(p1, p2, ncol = 2)
```


### Gamma prior $\alpha=10$

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/non_spatial/simu_nonspatial_chr1_50kb_gammaprior10.rdata")

print(sprintf("true alpha = %s", alpha_prior))
print(sprintf("estimated alpha = %s", alpha_est))
plot_randeff(data)
plot_dnm(data,windowsize)
scatter_plot_randeff(data)

data$y_pred <- data$randeff_est * data$exp_rl
p1 <- scatter_plot_mutct(data = data, xcol = "y_pred", ycol = "y_sample")
p2 <- scatter_plot_mutct(data = data, xcol ="exp_rl" , ycol = "y_sample")
grid.arrange(p1, p2, ncol = 2)
```


### Gamma prior $\alpha=30$

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/non_spatial/simu_nonspatial_chr1_50kb_gammaprior30.rdata")

print(sprintf("true alpha = %s", alpha_prior))
print(sprintf("estimated alpha = %s", alpha_est))
plot_randeff(data)
plot_dnm(data,windowsize)
scatter_plot_randeff(data)

data$y_pred <- data$randeff_est * data$exp_rl
p1 <- scatter_plot_mutct(data = data, xcol = "y_pred", ycol = "y_sample")
p2 <- scatter_plot_mutct(data = data, xcol ="exp_rl" , ycol = "y_sample")
grid.arrange(p1, p2, ncol = 2)
```


### Gamma prior $\alpha=50$

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/non_spatial/simu_nonspatial_chr1_50kb_gammaprior50.rdata")

print(sprintf("true alpha = %s", alpha_prior))
print(sprintf("estimated alpha = %s", alpha_est))
plot_randeff(data)
plot_dnm(data,windowsize)
scatter_plot_randeff(data)

data$y_pred <- data$randeff_est * data$exp_rl
p1 <- scatter_plot_mutct(data = data, xcol = "y_pred", ycol = "y_sample")
p2 <- scatter_plot_mutct(data = data, xcol ="exp_rl" , ycol = "y_sample")
grid.arrange(p1, p2, ncol = 2)
```


## 100 kb

### Gamma prior $\alpha=5$

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/non_spatial/simu_nonspatial_chr1_100kb_gammaprior5.rdata")

print(sprintf("true alpha = %s", alpha_prior))
print(sprintf("estimated alpha = %s", alpha_est))
plot_randeff(data)
plot_dnm(data,windowsize)
scatter_plot_randeff(data)

data$y_pred <- data$randeff_est * data$exp_rl
p1 <- scatter_plot_mutct(data = data, xcol = "y_pred", ycol = "y_sample")
p2 <- scatter_plot_mutct(data = data, xcol ="exp_rl" , ycol = "y_sample")
grid.arrange(p1, p2, ncol = 2)
```


### Gamma prior $\alpha=10$

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/non_spatial/simu_nonspatial_chr1_100kb_gammaprior10.rdata")

print(sprintf("true alpha = %s", alpha_prior))
print(sprintf("estimated alpha = %s", alpha_est))
plot_randeff(data)
plot_dnm(data,windowsize)
scatter_plot_randeff(data)

data$y_pred <- data$randeff_est * data$exp_rl
p1 <- scatter_plot_mutct(data = data, xcol = "y_pred", ycol = "y_sample")
p2 <- scatter_plot_mutct(data = data, xcol ="exp_rl" , ycol = "y_sample")
grid.arrange(p1, p2, ncol = 2)
```


### Gamma prior $\alpha=30$

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/non_spatial/simu_nonspatial_chr1_100kb_gammaprior30.rdata")

print(sprintf("true alpha = %s", alpha_prior))
print(sprintf("estimated alpha = %s", alpha_est))
plot_randeff(data)
plot_dnm(data,windowsize)
scatter_plot_randeff(data)

data$y_pred <- data$randeff_est * data$exp_rl
p1 <- scatter_plot_mutct(data = data, xcol = "y_pred", ycol = "y_sample")
p2 <- scatter_plot_mutct(data = data, xcol ="exp_rl" , ycol = "y_sample")
grid.arrange(p1, p2, ncol = 2)
```


### Gamma prior $\alpha=50$

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/non_spatial/simu_nonspatial_chr1_100kb_gammaprior50.rdata")

print(sprintf("true alpha = %s", alpha_prior))
print(sprintf("estimated alpha = %s", alpha_est))
plot_randeff(data)
plot_dnm(data,windowsize)
scatter_plot_randeff(data)

data$y_pred <- data$randeff_est * data$exp_rl
p1 <- scatter_plot_mutct(data = data, xcol = "y_pred", ycol = "y_sample")
p2 <- scatter_plot_mutct(data = data, xcol ="exp_rl" , ycol = "y_sample")
grid.arrange(p1, p2, ncol = 2)
```