---
title: "Calibrating Trost's de novo data with Roulette (Cell 2022)"
author: "XSun"
date: "2023-07-04"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(data.table)
library(ggplot2)
```


Data source: 

[https://www.sciencedirect.com/science/article/pii/S0092867422013241?via=ihub#abs0020 [Cell 2022]](https://www.sciencedirect.com/science/article/abs/pii/S0092867422013241?via=ihub)


# Data overview

The paper provides a collection of 167,760 de novo mutations from MSSNG, SPARK, and ASC. 145,237 of them are de novo SNV. 

| data set    | trios | cases | control | # of de novo mutations |
|-------------|-------|-------|---------|------------------------|
| ASC         | 6430  | 5556  | 8809    | 15772  [source](https://www.sciencedirect.com/science/article/pii/S0092867419313984)|
| MSSNG       | 3586  | -     | -       | 126788                 |
| SPARK pilot | 465   | -     | -       | 615    [source](https://www.nature.com/articles/s41525-019-0093-8)                |
| SPARK WES 1 | 6577  | -     | -       | 19825  [source](https://www.sciencedirect.com/science/article/pii/S0896627318300187?via%3Dihub)|
| SPARK WES 2 | 2167  | -     | -       | 4760   [source](https://www.sciencedirect.com/science/article/pii/S0896627318300187?via%3Dihub)|


## summary for MSSNG de novo SNVs 

| mutation types | # of de nove SNVs (total = 107377)        |
|----------------|-----------------------|
| a->c t->g      | 4014 + 4120 = 8134   |
| a->g t->c      | 14229 + 14094 = 28323 |
| a->t t->a      | 3515 + 3636 = 7141   |
| c->a g->t      | 5030 + 4843 = 9873 |
| c->g g->c      | 4830 + 5038 = 9868 |
| c->t g->a      | 21747 + 22281 = 44028 |


```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=5}

data <- data.frame(
  category = c("a->c", "a->g", "a->t", "c->a", "c->g", "c->t"),
  percentage = c(7.58,26.37,6.66, 9.20, 9.19,41.01)
)
data$angle <- cumsum(data$percentage) - (0.5 * data$percentage)
ggplot(data, aes(x = "", y = percentage, fill = category)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  theme_void() +
  theme(legend.position = "bottom") +
  # Adjust the position of the labels
  geom_label(
    aes(label = paste0(category, "\n", percentage, "%")),
    position = position_stack(vjust = 0.5),
    size = 3.5,
    show.legend = FALSE
  )
```

## summary for all de novo SNVs collected by the paper

```{r echo=FALSE, message=FALSE, warning=FALSE}
print("all 6 mutation types")
an <- fread("/project/xinhe/xsun/mutation_rate/data/trost_denovo/processed/alltype.bed")
colnames(an) <-c("chr","pos","ref","alt")
an[1:10,]
```

| mutation types | # of mutations        |
|----------------|-----------------------|
| a->c t->g      | 6613 + 6662 = 13275   |
| a->g t->c      | 17381 + 17332 = 34713 |
| a->t t->a      | 4354 + 4537 = 8891   |
| c->a g->t      | 6634 + 6474 = 13108 |
| c->g g->c      | 6681 + 6944 = 13625 |
| c->t g->a      | 30553 + 31072 = 61625 |


```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=5}

data <- data.frame(
  category = c("a->c", "a->g", "a->t", "c->a", "c->g", "c->t"),
  percentage = c(9.14,23.90, 6.12, 9.03, 9.38,42.43)
)
data$angle <- cumsum(data$percentage) - (0.5 * data$percentage)
ggplot(data, aes(x = "", y = percentage, fill = category)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  theme_void() +
  theme(legend.position = "bottom") +
  # Adjust the position of the labels
  geom_label(
    aes(label = paste0(category, "\n", percentage, "%")),
    position = position_stack(vjust = 0.5),
    size = 3.5,
    show.legend = FALSE
  )
```


<!-- # Rescaling -->

<!-- ```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=5} -->

<!-- load("/project/xinhe/xsun/mutation_rate/4.de_novo_an/results/scaling_factor.rdata") -->


<!-- print("fitting: observed (An et al.) ~ expected (Roulette) + 0") -->
<!-- fit <- lm(denovo_per_chr ~ rl_per_chr + 0) -->
<!-- print(summary(fit)) -->

<!-- data <- data.frame(expected = rl_per_chr, observed = denovo_per_chr) -->
<!-- print(data) -->

<!-- # Create the plot -->
<!-- ggplot(data, aes(x = expected, y = observed)) + -->
<!--   geom_point() + -->
<!--   geom_abline(slope = fit$coefficients[1], intercept = 0, color = "red") + -->
<!--   xlab("Expected_Roulette") + -->
<!--   ylab("Observed_An et al.")  -->

<!-- print(paste0("the scalling factor is ",round(fit$coefficients[1],digits = 5))) -->
<!-- print(paste0("obs = ",round(fit$coefficients[1],digits = 5),"* exp")) -->
<!-- ``` -->


<!-- # Random effect distribution (after rescaling) -->

<!-- ## Window size: 10kb -->

<!-- ### chr1 -->

<!-- All 6 mutation types, random effect: obs/exp -->

<!-- ```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10} -->

<!-- data <- fread("/project/xinhe/xsun/mutation_rate/4.de_novo_an/data/per_chr_processed/chr1/rl_an_combine_chr1_10000.txt") -->
<!-- colnames(data) <- c("chr","start","end","obs","exp") -->
<!-- data$Mb <- data$start/1000000 -->
<!-- data$randeff <- data$obs / data$exp -->

<!-- print("the number of observed de novo mutations per window") -->
<!-- table(data$obs) -->

<!-- ggplot(data, aes(x = Mb, y = obs)) + -->
<!--   geom_point(size=0.2) + -->
<!--   scale_x_continuous(name = "Genomic position (Mb)") + -->
<!--   ylab("# of observed de novo mutations") + -->
<!--   ggtitle("Distribution of observed de novo mutations") +  -->
<!--   theme_bw() + -->
<!--   theme(plot.title = element_text(hjust = 0.5)) -->


<!-- ggplot(data, aes(x = Mb, y = randeff)) + -->
<!--   geom_point(size=0.2) + -->
<!--   scale_x_continuous(name = "Genomic position (Mb)") + -->
<!--   ylab("Random effects") + -->
<!--   ggtitle("Distribution of random effects") +  -->
<!--   theme_bw() + -->
<!--   theme(plot.title = element_text(hjust = 0.5)) -->


<!-- ``` -->

<!-- ### chr22 -->

<!-- All 6 mutation types, random effect: obs/exp -->

<!-- ```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10} -->

<!-- data <- fread("/project/xinhe/xsun/mutation_rate/4.de_novo_an/data/per_chr_processed/chr22/rl_an_combine_chr22_10000.txt") -->
<!-- colnames(data) <- c("chr","start","end","obs","exp") -->
<!-- data$Mb <- data$start/1000000 -->
<!-- data$randeff <- data$obs / data$exp -->

<!-- print("the number of observed de novo mutations per window") -->
<!-- table(data$obs) -->

<!-- ggplot(data, aes(x = Mb, y = obs)) + -->
<!--   geom_point(size=0.2) + -->
<!--   scale_x_continuous(name = "Genomic position (Mb)") + -->
<!--   ylab("# of observed de novo mutations") + -->
<!--   ggtitle("Distribution of observed de novo mutations") +  -->
<!--   theme_bw() + -->
<!--   theme(plot.title = element_text(hjust = 0.5)) -->


<!-- ggplot(data, aes(x = Mb, y = randeff)) + -->
<!--   geom_point(size=0.2) + -->
<!--   scale_x_continuous(name = "Genomic position (Mb)") + -->
<!--   ylab("Random effects") + -->
<!--   ggtitle("Distribution of random effects") +  -->
<!--   theme_bw() + -->
<!--   theme(plot.title = element_text(hjust = 0.5)) -->


<!-- ``` -->

<!-- ## Window size: 50kb -->

<!-- ### chr1 -->

<!-- All 6 mutation types, random effect: obs/exp -->

<!-- ```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10} -->

<!-- data <- fread("/project/xinhe/xsun/mutation_rate/4.de_novo_an/data/per_chr_processed/chr1/rl_an_combine_chr1_50000.txt") -->
<!-- colnames(data) <- c("chr","start","end","obs","exp") -->
<!-- data$Mb <- data$start/1000000 -->
<!-- data$randeff <- data$obs / data$exp -->

<!-- print("the number of observed de novo mutations per window") -->
<!-- table(data$obs) -->

<!-- ggplot(data, aes(x = Mb, y = obs)) + -->
<!--   geom_point(size=0.2) + -->
<!--   scale_x_continuous(name = "Genomic position (Mb)") + -->
<!--   ylab("# of observed de novo mutations") + -->
<!--   ggtitle("Distribution of observed de novo mutations") +  -->
<!--   theme_bw() + -->
<!--   theme(plot.title = element_text(hjust = 0.5)) -->


<!-- ggplot(data, aes(x = Mb, y = randeff)) + -->
<!--   geom_point(size=0.2) + -->
<!--   scale_x_continuous(name = "Genomic position (Mb)") + -->
<!--   ylab("Random effects") + -->
<!--   ggtitle("Distribution of random effects") +  -->
<!--   theme_bw() + -->
<!--   theme(plot.title = element_text(hjust = 0.5)) -->


<!-- ``` -->

<!-- ### chr22 -->

<!-- All 6 mutation types, random effect: obs/exp -->

<!-- ```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10} -->

<!-- data <- fread("/project/xinhe/xsun/mutation_rate/4.de_novo_an/data/per_chr_processed/chr22/rl_an_combine_chr22_50000.txt") -->
<!-- colnames(data) <- c("chr","start","end","obs","exp") -->
<!-- data$Mb <- data$start/1000000 -->
<!-- data$randeff <- data$obs / data$exp -->

<!-- print("the number of observed de novo mutations per window") -->
<!-- table(data$obs) -->

<!-- ggplot(data, aes(x = Mb, y = obs)) + -->
<!--   geom_point(size=0.2) + -->
<!--   scale_x_continuous(name = "Genomic position (Mb)") + -->
<!--   ylab("# of observed de novo mutations") + -->
<!--   ggtitle("Distribution of observed de novo mutations") +  -->
<!--   theme_bw() + -->
<!--   theme(plot.title = element_text(hjust = 0.5)) -->


<!-- ggplot(data, aes(x = Mb, y = randeff)) + -->
<!--   geom_point(size=0.2) + -->
<!--   scale_x_continuous(name = "Genomic position (Mb)") + -->
<!--   ylab("Random effects") + -->
<!--   ggtitle("Distribution of random effects") +  -->
<!--   theme_bw() + -->
<!--   theme(plot.title = element_text(hjust = 0.5)) -->


<!-- ``` -->
