---
title: "Summary table for all the real analyses"
author: "XSun"
date: "2024-06-19"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
---

```{r warning=F, message=F}
library(ggplot2)
library(dplyr)
library(gridExtra)

create_pairwise_scatter_plots <- function(SSC_an, SSC_trost, mssng_trost, ASC_MSSNG_SPARK_trost) {
  # Extract randeff_est columns
  randeff_SSC_an <- SSC_an$randeff_est
  randeff_SSC_trost <- SSC_trost$randeff_est
  randeff_mssng_trost <- mssng_trost$randeff_est
  randeff_ASC_MSSNG_SPARK_trost <- ASC_MSSNG_SPARK_trost$randeff_est

  # Combine into a single data frame for pairwise comparison
  combined_df <- data.frame(
    randeff_SSC_an,
    randeff_SSC_trost,
    randeff_mssng_trost,
    randeff_ASC_MSSNG_SPARK_trost
  )

  # Define pairs for comparison
  pairs <- list(
    c("randeff_SSC_an", "randeff_SSC_trost"),
    c("randeff_SSC_an", "randeff_mssng_trost"),
    c("randeff_SSC_an", "randeff_ASC_MSSNG_SPARK_trost"),
    c("randeff_SSC_trost", "randeff_mssng_trost"),
    c("randeff_SSC_trost", "randeff_ASC_MSSNG_SPARK_trost"),
    c("randeff_mssng_trost", "randeff_ASC_MSSNG_SPARK_trost")
  )

  # Create scatter plots
  plots <- lapply(pairs, function(pair) {
    ggplot(combined_df, aes_string(x = pair[1], y = pair[2])) +
      geom_point() +
      geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
      theme_minimal() +
      labs(title = "Estimated nugget effects",
           x = pair[1],
           y = pair[2])
  })

  # Arrange plots in a 2x3 grid
  grid.arrange(grobs = plots, ncol = 3, nrow = 2)
}
```

# Fitting rsq in cross validation

Below is a table summarizing the fitting rsq for `observed_mutation_count ~ predicted_mutation_count + 0` in cross validation. 

by_segments means, we partitioned the chromosome if there is large gaps (>20mb)

by_chr means, we treat the whole chromosome as one segment and run smashgen on it. 

|                   | Roulette baseline&SSC | SSC, Trost Cell 2022 | Roulette baseline&MSSNG | MSSNG, Trost Cell 2022 | Roulette baseline&ASC+MSSNG+SPARK | ASC+MSSNG+SPARK, Trost Cell 2022 | Roulette baseline&SSC | SSC, An Science 2018 |
|-------------------|---------------------------|----------------------|-----------------------------|------------------------|---------------------------------------|----------------------------------|---------------------------|----------------------|
| 10kb_by segments  | 0.084                     | 0.083                | 0.032                       | 0.029                  | 0.034                                 | 0.027                            | 0.072                     | 0.073                |
| 10kb_by chr       | 0.084                     | 0.083                | 0.032                       | 0.029                  | 0.034                                 | 0.027                            | 0.072                     | 0.072                |
| 30kb_by segments  | 0.236                     | 0.231                | 0.086                       | 0.096                  | 0.102                                 | 0.107                            | 0.187                     | 0.19                 |
| 30kb_by chr       | 0.236                     | 0.231                | 0.086                       | 0.096                  | 0.102                                 | 0.107                            | 0.187                     | 0.19                 |
| 50kb_by segments  | 0.356                     | 0.358                | 0.136                       | 0.175                  | 0.154                                 | 0.193                            | 0.287                     | 0.293                |
| 50kb_by chr       | 0.356                     | 0.358                | 0.136                       | 0.176                  | 0.154                                 | 0.194                            | 0.287                     | 0.293                |
| 100kb_by segments | 0.466                     | 0.461                | 0.209                       | 0.25                   | 0.088                                 | 0.09                             | 0.419                     | 0.427                |
| 100kb_by chr      | 0.466                     | 0.461                | 0.209                       | 0.25                   | 0.088                                 | 0.091                            | 0.419                     | 0.426                |


Quick links to the detailed results

SSC, Trost Cell 2022 https://xsun1229.github.io/mutation_rate/denovo0422_realdata_CV.html

MSSNG, Trost Cell 2022 https://xsun1229.github.io/mutation_rate/denovo0422_realdata_CV_mssng.html

ASC+MSSNG+SPARK, Trost Cell 2022 https://xsun1229.github.io/mutation_rate/denovo0422_realdata_CV_integrate_trost.html

SSC, An Science 2018 https://xsun1229.github.io/mutation_rate/denovo0422_realdata_CV_ssc_an.html


# Comparing the estimated nugget effects -- estimated by segments

## 10kb

```{r fig.width=12, fig.height=8, warning=F, message=F}

SSC_an <- get(load("/project/xinhe/xsun/mutation_rate/12.denovo_real_ssc_an/results/ssc_an_per_window_chr1_10kb_fitted.rdata"))
SSC_trost <- get(load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_10kb_fitted.rdata"))
mssng_trost <- get(load("/project/xinhe/xsun/mutation_rate/11.denovo_real_mssng/results/mssng_trost_per_window_chr1_10kb_fitted.rdata"))
ASC_MSSNG_SPARK_trost <- get(load("/project/xinhe/xsun/mutation_rate/13.denovo_real_integration_trost/results/integration_trost_per_window_chr1_10kb_fitted.rdata"))

create_pairwise_scatter_plots(SSC_an, SSC_trost, mssng_trost, ASC_MSSNG_SPARK_trost)
```


## 30kb

```{r fig.width=12, fig.height=8, warning=F, message=F}

SSC_an <- get(load("/project/xinhe/xsun/mutation_rate/12.denovo_real_ssc_an/results/ssc_an_per_window_chr1_30kb_fitted.rdata"))
SSC_trost <- get(load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_30kb_fitted.rdata"))
mssng_trost <- get(load("/project/xinhe/xsun/mutation_rate/11.denovo_real_mssng/results/mssng_trost_per_window_chr1_30kb_fitted.rdata"))
ASC_MSSNG_SPARK_trost <- get(load("/project/xinhe/xsun/mutation_rate/13.denovo_real_integration_trost/results/integration_trost_per_window_chr1_30kb_fitted.rdata"))

create_pairwise_scatter_plots(SSC_an, SSC_trost, mssng_trost, ASC_MSSNG_SPARK_trost)
```


## 50kb

```{r fig.width=12, fig.height=8, warning=F, message=F}

SSC_an <- get(load("/project/xinhe/xsun/mutation_rate/12.denovo_real_ssc_an/results/ssc_an_per_window_chr1_50kb_fitted.rdata"))
SSC_trost <- get(load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_50kb_fitted.rdata"))
mssng_trost <- get(load("/project/xinhe/xsun/mutation_rate/11.denovo_real_mssng/results/mssng_trost_per_window_chr1_50kb_fitted.rdata"))
ASC_MSSNG_SPARK_trost <- get(load("/project/xinhe/xsun/mutation_rate/13.denovo_real_integration_trost/results/integration_trost_per_window_chr1_50kb_fitted.rdata"))

create_pairwise_scatter_plots(SSC_an, SSC_trost, mssng_trost, ASC_MSSNG_SPARK_trost)
```

## 100kb

```{r fig.width=12, fig.height=8, warning=F, message=F}

SSC_an <- get(load("/project/xinhe/xsun/mutation_rate/12.denovo_real_ssc_an/results/ssc_an_per_window_chr1_100kb_fitted.rdata"))
SSC_trost <- get(load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_100kb_fitted.rdata"))
mssng_trost <- get(load("/project/xinhe/xsun/mutation_rate/11.denovo_real_mssng/results/mssng_trost_per_window_chr1_100kb_fitted.rdata"))
ASC_MSSNG_SPARK_trost <- get(load("/project/xinhe/xsun/mutation_rate/13.denovo_real_integration_trost/results/integration_trost_per_window_chr1_100kb_fitted.rdata"))

create_pairwise_scatter_plots(SSC_an, SSC_trost, mssng_trost, ASC_MSSNG_SPARK_trost)
```