---
title: "Spatial random effect distribution"
author: "XSun"
date: "2023-07-18"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(data.table)
library(ggplot2)
```

# Introduction

We consider spatial random effect here: 

For each window $i$, the random effect $\theta_i$ is: 

$$ \theta_i = y_i / \mu_i$$
where $y_i$ is the number of observed SNV in each window, $\mu_i$ is the sum of Roulette estimation (after calibration) in this window


# SSC (Cell 2022)

## Window size: 10kb

### chr1

All 6 mutation types, random effect: obs/exp

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

data <- fread("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr1/rl_ssc_trost_combine_chr1_10000.txt")
colnames(data) <- c("chr","start","end","obs","exp")
data$Mb <- data$start/1000000
data$randeff <- data$obs / data$exp

print("the number of observed de novo mutations per window")
table(data$obs)

ggplot(data, aes(x = Mb, y = obs)) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of observed de novo mutations") +
  ggtitle("Distribution of observed de novo mutations") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


ggplot(data, aes(x = Mb, y = randeff)) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("Random effects") +
  ggtitle("Distribution of random effects") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


```

### chr22

All 6 mutation types, random effect: obs/exp

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

data <- fread("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr22/rl_ssc_trost_combine_chr22_10000.txt")
colnames(data) <- c("chr","start","end","obs","exp")
data$Mb <- data$start/1000000
data$randeff <- data$obs / data$exp

print("the number of observed de novo mutations per window")
table(data$obs)

ggplot(data, aes(x = Mb, y = obs)) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of observed de novo mutations") +
  ggtitle("Distribution of observed de novo mutations") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


ggplot(data, aes(x = Mb, y = randeff)) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("Random effects") +
  ggtitle("Distribution of random effects") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


```

## Window size: 50kb

### chr1

All 6 mutation types, random effect: obs/exp

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

data <- fread("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr1/rl_ssc_trost_combine_chr1_50000.txt")
colnames(data) <- c("chr","start","end","obs","exp")
data$Mb <- data$start/1000000
data$randeff <- data$obs / data$exp

print("the number of observed de novo mutations per window")
table(data$obs)

ggplot(data, aes(x = Mb, y = obs)) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of observed de novo mutations") +
  ggtitle("Distribution of observed de novo mutations") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


ggplot(data, aes(x = Mb, y = randeff)) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("Random effects") +
  ggtitle("Distribution of random effects") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


```

### chr22

All 6 mutation types, random effect: obs/exp

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

data <- fread("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr22/rl_ssc_trost_combine_chr22_50000.txt")
colnames(data) <- c("chr","start","end","obs","exp")
data$Mb <- data$start/1000000
data$randeff <- data$obs / data$exp

print("the number of observed de novo mutations per window")
table(data$obs)

ggplot(data, aes(x = Mb, y = obs)) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of observed de novo mutations") +
  ggtitle("Distribution of observed de novo mutations") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


ggplot(data, aes(x = Mb, y = randeff)) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("Random effects") +
  ggtitle("Distribution of random effects") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


```


# MSSNG (Cell 2022)

## Window size: 10kb

### chr1


```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

data <- fread("/project/xinhe/xsun/mutation_rate/5.de_novo_trost/data/per_chr_processed/chr1/rl_trost_combine_chr1_10000_mssng.txt")
colnames(data) <- c("chr","start","end","obs","exp")
data$Mb <- data$start/1000000
data$randeff <- data$obs / data$exp

print("the number of observed de novo mutations per window")
table(data$obs)

ggplot(data, aes(x = Mb, y = obs)) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of observed de novo mutations") +
  ggtitle("Distribution of observed de novo mutations") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


ggplot(data, aes(x = Mb, y = randeff)) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("Random effects") +
  ggtitle("Distribution of random effects") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


```

### chr22

All 6 mutation types, random effect: obs/exp

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

data <- fread("/project/xinhe/xsun/mutation_rate/5.de_novo_trost/data/per_chr_processed/chr22/rl_trost_combine_chr22_10000_mssng.txt")
colnames(data) <- c("chr","start","end","obs","exp")
data$Mb <- data$start/1000000
data$randeff <- data$obs / data$exp

print("the number of observed de novo mutations per window")
table(data$obs)

ggplot(data, aes(x = Mb, y = obs)) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of observed de novo mutations") +
  ggtitle("Distribution of observed de novo mutations") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


ggplot(data, aes(x = Mb, y = randeff)) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("Random effects") +
  ggtitle("Distribution of random effects") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


```

## Window size: 50kb

### chr1

All 6 mutation types, random effect: obs/exp

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

data <- fread("/project/xinhe/xsun/mutation_rate/5.de_novo_trost/data/per_chr_processed/chr1/rl_trost_combine_chr1_50000_mssng.txt")
colnames(data) <- c("chr","start","end","obs","exp")
data$Mb <- data$start/1000000
data$randeff <- data$obs / data$exp

print("the number of observed de novo mutations per window")
table(data$obs)

ggplot(data, aes(x = Mb, y = obs)) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of observed de novo mutations") +
  ggtitle("Distribution of observed de novo mutations") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


ggplot(data, aes(x = Mb, y = randeff)) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("Random effects") +
  ggtitle("Distribution of random effects") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


```

### chr22

All 6 mutation types, random effect: obs/exp

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

data <- fread("/project/xinhe/xsun/mutation_rate/5.de_novo_trost/data/per_chr_processed/chr22/rl_trost_combine_chr22_50000_mssng.txt")
colnames(data) <- c("chr","start","end","obs","exp")
data$Mb <- data$start/1000000
data$randeff <- data$obs / data$exp

print("the number of observed de novo mutations per window")
table(data$obs)

ggplot(data, aes(x = Mb, y = obs)) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of observed de novo mutations") +
  ggtitle("Distribution of observed de novo mutations") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


ggplot(data, aes(x = Mb, y = randeff)) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("Random effects") +
  ggtitle("Distribution of random effects") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


```
