---
title: "SMASHgen estimation: spatial and non-spatial random effects, random effects_spatial from normal distribution"
author: "XSun"
date: "2024-10-03"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

# Introduction

## Simulation setup 

We combined the spatial and non-spatial models here

$y_i = \text{pois}(u_i\theta_i)$

where $u_i$ is expected rate from calibrated roulette. $y_i$ is the observed mutation rate. $\theta_i$ is the random effect.

1. Simulating the random effects

We simulated spatial random effects as we did in [spatial-only model](https://xsun1229.github.io/mutation_rate/denovo0422_simulation_spatial_norm.html) 
We simulated spatial random effects as we did in [non-spatial-only model](https://xsun1229.github.io/mutation_rate/denovo0422_simulation_nonspatial.html)

Then we multiply the spatial and non-spatial random effects, and normalized them, making $\text{mean}(\tilde\theta_i) =1$

```{r}

generate_randeff_nonspatial_gammaprior_spatial_normal <-  function(num_windows_all, num_windows_per_seg_min, num_windows_per_seg_max, randeff_max, num_seg_spatial,alpha,normal_sigma) {
  
  # Create the vector
  randeff <- rep(1,num_windows_all)
  
  #spatial
  segment_lengths_spatial <- ceiling(runif(num_seg_spatial, min = num_windows_per_seg_min, max = num_windows_per_seg_max))
  
  # Calculate the total length covered by spatial random effects
  total_segment_length <- sum(segment_lengths_spatial) 
  
  # Calculate the number of zeros needed
  num_zeros <- num_windows_all - total_segment_length
  
  # Randomly select positions for spatial segment starts 
  segment_starts_spatial <- sample(1:num_zeros, num_seg_spatial, replace = FALSE)
  segment_ends_spatial <- segment_starts_spatial + segment_lengths_spatial - 1
  
  # Assign random values to segments
  for (i in 1:(num_seg_spatial)) {
    start_pos <- segment_starts_spatial[i]
    end_pos <- start_pos + segment_lengths_spatial[i] - 1
    
    log_eff <- rnorm(1,mean = 0,sd=normal_sigma)
    eff <- exp(log_eff)
    randeff[start_pos:end_pos] <- eff
  }
  
  
  ## non-spatial
  alpha <- alpha
  randeff_nonspatial <- rgamma(n = num_windows_all, shape = alpha, rate = alpha)
  
  
  randeff <- randeff * randeff_nonspatial
  
  # Adjust values to meet the mean of 1
  mean_adjustment <- 1 - mean(randeff)
  randeff <- randeff + mean_adjustment
  # Clip values between 0 and randeff_max
  randeff[randeff < 0] <- 0
  #randeff[randeff > randeff_max] <- randeff_max
  
  
  return(randeff)
}

set.seed(2)
randeff <- generate_randeff_nonspatial_gammaprior_spatial_normal(num_windows_all = 3000,num_windows_per_seg_min = 20,num_windows_per_seg_max = 50,randeff_max = 3,num_seg_spatial = 10,alpha = 30,normal_sigma = 1)

plot(randeff,type = "l",ylab = "simulated random effect")
```

2. Then, we estimated $\hat\theta_i$ using smashgen

3. The results are compared with the estimation from [2steps method](https://xsun1229.github.io/mutation_rate/denovo0422_simulation_spatial_norm_nonspatial.html), [non-spatial method](https://xsun1229.github.io/mutation_rate/denovo0422_simulation_nonspatial.html), and [spatial smoothed method](https://xsun1229.github.io/mutation_rate/denovo0422_simulation_spatial_norm.html)

## Functions used

```{r}
library(ggplot2)

plot_randeff_windows_smashgen <- function(data) {
  
  ggplot(data, aes(x = window)) +
  geom_line(aes(y = randeff_est_spatial_smashgen, color = "Estimated random effects -- spatial"), alpha = 0.4) +
  #geom_line(aes(y = randeff_est_nonspatial_smashgen, color = "Estimated random effects -- nonspatial"), alpha = 0.2) +
  geom_line(aes(y = randeff_est_total_smashgen, color = "Estimated random effects -- total"), alpha = 0.2) +
  #geom_line(aes(y = roulette, color = "Roulette predicted mutation count"), alpha = 0.2) +
  geom_line(aes(y = randeff_simulated, color = "Simulated random effects"), alpha = 0.1) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "black", size = 0.3) +  # y=1 line
  annotate("text", x = Inf, y = 1, label = "y = 1", vjust = -1, hjust = 1.1, color = "black", size = 2.5) +  # y=1 text
  labs(x = "Genomic position (Mb)",
       y = "Random effects") +
  scale_color_manual(name = "Group",
                     values = c("Estimated random effects -- spatial" = "purple",
                                #"Estimated random effects -- nonspatial" = "orange",
                                "Estimated random effects -- total" = "blue",
                                #"Roulette predicted mutation count" = "green", 
                                "Simulated random effects" = "red")) +
  guides(color = guide_legend(override.aes = list(alpha = 1))) +
  ggtitle("smashgen estimation")+ 
  theme_minimal()
  
}


```


# Gamma alpha = 0.5

## Fixed normal_var_spatial =0.5, different spatial window size window

### 200kb ~ 1mb

```{r fig.width=16, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_0.5_minseg4_maxseg20.rdata")


plot_randeff_windows_smashgen(dat)
```

### 500kb ~ 1mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_0.5_minseg10_maxseg20.rdata")



plot_randeff_windows_smashgen(dat)
```

### 1mb ~ 2mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_0.5_minseg20_maxseg40.rdata")



plot_randeff_windows_smashgen(dat)
```


### 2mb ~ 4mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_0.5_minseg40_maxseg80.rdata")



plot_randeff_windows_smashgen(dat)
```

### 4mb ~ 6mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_0.5_minseg80_maxseg120.rdata")



plot_randeff_windows_smashgen(dat)
```

### 6mb ~ 8mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_0.5_minseg120_maxseg160.rdata")



plot_randeff_windows_smashgen(dat)
```

### 8mb ~ 10mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_0.5_minseg160_maxseg200.rdata")



plot_randeff_windows_smashgen(dat)
```






## Fixed normal_var_spatial =1, different spatial window size window

### 200kb ~ 1mb

```{r fig.width=16, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_0.5_minseg4_maxseg20.rdata")


plot_randeff_windows_smashgen(dat)
```

### 500kb ~ 1mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_0.5_minseg10_maxseg20.rdata")



plot_randeff_windows_smashgen(dat)
```

### 1mb ~ 2mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_0.5_minseg20_maxseg40.rdata")



plot_randeff_windows_smashgen(dat)
```


### 2mb ~ 4mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_0.5_minseg40_maxseg80.rdata")



plot_randeff_windows_smashgen(dat)
```

### 4mb ~ 6mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_0.5_minseg80_maxseg120.rdata")



plot_randeff_windows_smashgen(dat)
```

### 6mb ~ 8mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_0.5_minseg120_maxseg160.rdata")



plot_randeff_windows_smashgen(dat)
```

### 8mb ~ 10mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_0.5_minseg160_maxseg200.rdata")


plot_randeff_windows_smashgen(dat)
```









# Gamma alpha = 1

## Fixed normal_var_spatial =0.5, different spatial window size window

### 200kb ~ 1mb

```{r fig.width=16, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_1_minseg4_maxseg20.rdata")


plot_randeff_windows_smashgen(dat)
```

### 500kb ~ 1mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_1_minseg10_maxseg20.rdata")



plot_randeff_windows_smashgen(dat)
```

### 1mb ~ 2mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_1_minseg20_maxseg40.rdata")



plot_randeff_windows_smashgen(dat)
```


### 2mb ~ 4mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_1_minseg40_maxseg80.rdata")



plot_randeff_windows_smashgen(dat)
```

### 4mb ~ 6mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_1_minseg80_maxseg120.rdata")



plot_randeff_windows_smashgen(dat)
```

### 6mb ~ 8mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_1_minseg120_maxseg160.rdata")



plot_randeff_windows_smashgen(dat)
```

### 8mb ~ 10mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_1_minseg160_maxseg200.rdata")



plot_randeff_windows_smashgen(dat)
```






## Fixed normal_var_spatial =1, different spatial window size window

### 200kb ~ 1mb

```{r fig.width=16, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_1_minseg4_maxseg20.rdata")


plot_randeff_windows_smashgen(dat)
```

### 500kb ~ 1mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_1_minseg10_maxseg20.rdata")



plot_randeff_windows_smashgen(dat)
```

### 1mb ~ 2mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_1_minseg20_maxseg40.rdata")



plot_randeff_windows_smashgen(dat)
```


### 2mb ~ 4mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_1_minseg40_maxseg80.rdata")



plot_randeff_windows_smashgen(dat)
```

### 4mb ~ 6mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_1_minseg80_maxseg120.rdata")



plot_randeff_windows_smashgen(dat)
```

### 6mb ~ 8mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_1_minseg120_maxseg160.rdata")



plot_randeff_windows_smashgen(dat)
```

### 8mb ~ 10mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_1_minseg160_maxseg200.rdata")


plot_randeff_windows_smashgen(dat)
```










# Gamma alpha = 2

## Fixed normal_var_spatial =0.5, different spatial window size window

### 200kb ~ 1mb

```{r fig.width=16, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_2_minseg4_maxseg20.rdata")


plot_randeff_windows_smashgen(dat)
```

### 500kb ~ 1mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_2_minseg10_maxseg20.rdata")



plot_randeff_windows_smashgen(dat)
```

### 1mb ~ 2mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_2_minseg20_maxseg40.rdata")



plot_randeff_windows_smashgen(dat)
```


### 2mb ~ 4mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_2_minseg40_maxseg80.rdata")



plot_randeff_windows_smashgen(dat)
```

### 4mb ~ 6mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_2_minseg80_maxseg120.rdata")



plot_randeff_windows_smashgen(dat)
```

### 6mb ~ 8mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_2_minseg120_maxseg160.rdata")



plot_randeff_windows_smashgen(dat)
```

### 8mb ~ 10mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_2_minseg160_maxseg200.rdata")



plot_randeff_windows_smashgen(dat)
```






## Fixed normal_var_spatial =1, different spatial window size window

### 200kb ~ 1mb

```{r fig.width=16, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_2_minseg4_maxseg20.rdata")


plot_randeff_windows_smashgen(dat)
```

### 500kb ~ 1mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_2_minseg10_maxseg20.rdata")



plot_randeff_windows_smashgen(dat)
```

### 1mb ~ 2mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_2_minseg20_maxseg40.rdata")



plot_randeff_windows_smashgen(dat)
```


### 2mb ~ 4mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_2_minseg40_maxseg80.rdata")



plot_randeff_windows_smashgen(dat)
```

### 4mb ~ 6mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_2_minseg80_maxseg120.rdata")



plot_randeff_windows_smashgen(dat)
```

### 6mb ~ 8mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_2_minseg120_maxseg160.rdata")



plot_randeff_windows_smashgen(dat)
```

### 8mb ~ 10mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_2_minseg160_maxseg200.rdata")


plot_randeff_windows_smashgen(dat)
```




# Gamma alpha = 3

## Fixed normal_var_spatial =0.5, different spatial window size window

### 200kb ~ 1mb

```{r fig.width=16, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_3_minseg4_maxseg20.rdata")


plot_randeff_windows_smashgen(dat)
```

### 500kb ~ 1mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_3_minseg10_maxseg20.rdata")



plot_randeff_windows_smashgen(dat)
```

### 1mb ~ 2mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_3_minseg20_maxseg40.rdata")



plot_randeff_windows_smashgen(dat)
```


### 2mb ~ 4mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_3_minseg40_maxseg80.rdata")



plot_randeff_windows_smashgen(dat)
```

### 4mb ~ 6mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_3_minseg80_maxseg120.rdata")



plot_randeff_windows_smashgen(dat)
```

### 6mb ~ 8mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_3_minseg120_maxseg160.rdata")



plot_randeff_windows_smashgen(dat)
```

### 8mb ~ 10mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_3_minseg160_maxseg200.rdata")



plot_randeff_windows_smashgen(dat)
```






## Fixed normal_var_spatial =1, different spatial window size window

### 200kb ~ 1mb

```{r fig.width=16, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_3_minseg4_maxseg20.rdata")


plot_randeff_windows_smashgen(dat)
```

### 500kb ~ 1mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_3_minseg10_maxseg20.rdata")



plot_randeff_windows_smashgen(dat)
```

### 1mb ~ 2mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_3_minseg20_maxseg40.rdata")



plot_randeff_windows_smashgen(dat)
```


### 2mb ~ 4mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_3_minseg40_maxseg80.rdata")



plot_randeff_windows_smashgen(dat)
```

### 4mb ~ 6mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_3_minseg80_maxseg120.rdata")



plot_randeff_windows_smashgen(dat)
```

### 6mb ~ 8mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_3_minseg120_maxseg160.rdata")



plot_randeff_windows_smashgen(dat)
```

### 8mb ~ 10mb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal_smashgenonly/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_3_minseg160_maxseg200.rdata")


plot_randeff_windows_smashgen(dat)
```










