---
title: "Simulation"
author: "XSun"
date: "2023-07-28"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r echo=F,message=FALSE}
library(ggplot2)
```

# Introduction

1.  We set random effects $\tilde{\theta_i}$ for each window $i$, the mean value of $\tilde{\theta_i}$ is 1.

Below is a random example for the random effects

```{r echo=F,message=FALSE, fig.width=12, fig.height=5}
generate_random_genome_vector <- function(num_windows_all, num_windows_per_seg_min, num_windows_per_seg_max, randeff_max, num_seg) {
  # Generate random segment lengths
  segment_lengths <- sample(num_windows_per_seg_min:num_windows_per_seg_max, num_seg, replace = TRUE)
  
  # Calculate the total length covered by segments
  total_segment_length <- sum(segment_lengths)
  
  # Check if the total segment length is greater than the genome length
  # If so, reduce the last segment length to fit within the genome
  if (total_segment_length > num_windows_all) {
    segment_lengths[num_seg] <- segment_lengths[num_seg] - (total_segment_length - num_windows_all)
  }
  
  # Calculate the number of zeros needed
  num_zeros <- num_windows_all - total_segment_length
  
  # Randomly select positions for segment starts
  segment_starts <- sample(1:num_zeros, num_seg, replace = FALSE)
  
  # Create the vector
  randeff <- numeric(num_windows_all)
  
  # Assign random values to segments
  for (i in 1:num_seg) {
    start_pos <- segment_starts[i]
    end_pos <- start_pos + segment_lengths[i] - 1
    randeff[start_pos:end_pos] <- runif(1, 0, randeff_max)
  }
  
  # Adjust values to meet the mean of 1
  mean_adjustment <- 1 - mean(randeff)
  randeff <- randeff + mean_adjustment
  
  # Clip values between 0 and randeff_max
  randeff[randeff < 0] <- 0
  randeff[randeff > randeff_max] <- randeff_max
  
  # Return the resulting genome vector
  return(randeff)
}

num_windows_all <- 1000
num_windows_per_seg_min <- 10
num_windows_per_seg_max <- 100  
randeff_max <- 5
num_seg <- 10

randeff <-generate_random_genome_vector(num_windows_all, num_windows_per_seg_min, num_windows_per_seg_max, randeff_max, num_seg)

plot(randeff, type="l")

```

2. We sample the $\tilde{y_i}$ from

$$\tilde{y_i} \sim pois(\tilde{\theta_i}*\mu_i)$$
where $\mu_i$ is the expected mutations for each window $i$ from Roulette. 

3. We fit the $\tilde{y_i}$ and $\mu_i$ using smashgen and plot the $\tilde{\theta_i}$ and smoothed $\tilde{\theta}_i{_{smoothed}}$


# smashgen, SSC (Cell 2022, chr1)

## 1kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_1kb_maxiter50_maxrandeff2.rdata")

print("maximum random effect = 2")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_1kb_maxiter50_maxrandeff5.rdata")

print("maximum random effect = 5")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_1kb_maxiter50_maxrandeff10.rdata")

print("maximum random effect = 10")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_1kb_maxiter50_maxrandeff15.rdata")

print("maximum random effect = 15")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

## 10kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_10kb_maxiter50_maxrandeff2.rdata")

print("maximum random effect = 2")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_10kb_maxiter50_maxrandeff5.rdata")

print("maximum random effect = 5")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_10kb_maxiter50_maxrandeff10.rdata")

print("maximum random effect = 10")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_10kb_maxiter50_maxrandeff15.rdata")

print("maximum random effect = 15")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

## 30kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_30kb_maxiter50_maxrandeff2.rdata")

print("maximum random effect = 2")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_30kb_maxiter50_maxrandeff5.rdata")

print("maximum random effect = 5")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_30kb_maxiter50_maxrandeff10.rdata")

print("maximum random effect = 10")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_30kb_maxiter50_maxrandeff15.rdata")

print("maximum random effect = 15")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

## 50kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_50kb_maxiter50_maxrandeff2.rdata")

print("maximum random effect = 2")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_50kb_maxiter50_maxrandeff5.rdata")

print("maximum random effect = 5")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_50kb_maxiter50_maxrandeff10.rdata")

print("maximum random effect = 10")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_50kb_maxiter50_maxrandeff15.rdata")

print("maximum random effect = 15")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

## 100kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_100kb_maxiter50_maxrandeff2.rdata")

print("maximum random effect = 2")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_100kb_maxiter50_maxrandeff5.rdata")

print("maximum random effect = 5")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_100kb_maxiter50_maxrandeff10.rdata")

print("maximum random effect = 10")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_100kb_maxiter50_maxrandeff15.rdata")

print("maximum random effect = 15")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```


# smash.poiss, SSC (Cell 2022, chr1)

## 1kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_1kb_maxrandeff2_smash.rdata")

print("maximum random effect = 2")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_1kb_maxrandeff5_smash.rdata")

print("maximum random effect = 5")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_1kb_maxrandeff10_smash.rdata")

print("maximum random effect = 10")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_1kb_maxrandeff15_smash.rdata")

print("maximum random effect = 15")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

## 10kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_10kb_maxrandeff2_smash.rdata")

print("maximum random effect = 2")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_10kb_maxrandeff5_smash.rdata")

print("maximum random effect = 5")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_10kb_maxrandeff10_smash.rdata")

print("maximum random effect = 10")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_10kb_maxrandeff15_smash.rdata")

print("maximum random effect = 15")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

## 30kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_30kb_maxrandeff2_smash.rdata")

print("maximum random effect = 2")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_30kb_maxrandeff5_smash.rdata")

print("maximum random effect = 5")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_30kb_maxrandeff10_smash.rdata")

print("maximum random effect = 10")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_30kb_maxrandeff15_smash.rdata")

print("maximum random effect = 15")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

## 50kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_50kb_maxrandeff2_smash.rdata")

print("maximum random effect = 2")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_50kb_maxrandeff5_smash.rdata")

print("maximum random effect = 5")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_50kb_maxrandeff10_smash.rdata")

print("maximum random effect = 10")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_50kb_maxrandeff15_smash.rdata")

print("maximum random effect = 15")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

## 100kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_100kb_maxrandeff2_smash.rdata")

print("maximum random effect = 2")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_100kb_maxrandeff5_smash.rdata")

print("maximum random effect = 5")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_100kb_maxrandeff10_smash.rdata")

print("maximum random effect = 10")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_100kb_maxrandeff15_smash.rdata")

print("maximum random effect = 15")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

```
