---
title: "Simulation for spatial random effects"
author: "XSun"
date: "2024-04-22"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
---

# Introduction

## Simulation setup

1. The window sizes for the local random effects range from hundreds kb to 1mb.

2. The random effects $\tilde{\theta}_i$ for the hotspots were sampled from uniform distribution $$\tilde{\theta}_i \sim \text{U}[0,\text{maxeff}]$$, The prior $\text{maxeff} \in \{3,5\}$. For others, random effects $\tilde{\theta}_i$ were set to 1.

3. Then the random effects across all windows were ajusted, making the $\text{mean}(\tilde{\theta}_i)=1$.

Below is an example for the random effect

```{r collapse=TRUE, eval=T ,message=FALSE, fig.width=12, fig.height=5}
generate_random_genome_vector <- function(num_windows_all, num_windows_per_seg_min, num_windows_per_seg_max, randeff_max, num_seg) {
  # Generate random segment lengths
  
  set.seed(1)
  segment_lengths <- ceiling(runif(num_seg, min = num_windows_per_seg_min, max = num_windows_per_seg_max))
  
  # Calculate the total length covered by segments
  total_segment_length <- sum(segment_lengths)
  
  # Check if the total segment length is greater than the genome length
  # If so, reduce the last segment length to fit within the genome
  if (total_segment_length > num_windows_all) {
    segment_lengths[num_seg] <- segment_lengths[num_seg] - (total_segment_length - num_windows_all)
  }
  
  # Calculate the number of zeros needed
  num_zeros <- num_windows_all - total_segment_length
  
  # Randomly select positions for segment starts
  segment_starts <- sample(1:num_zeros, num_seg, replace = FALSE)
  
  # Create the vector
  randeff <- numeric(num_windows_all)
  
  # Assign random values to segments
  for (i in 1:num_seg) {
    start_pos <- segment_starts[i]
    end_pos <- start_pos + segment_lengths[i] - 1
    randeff[start_pos:end_pos] <- runif(1, 0, randeff_max)
  }
  
  # Adjust values to meet the mean of 1
  mean_adjustment <- 1 - mean(randeff)
  randeff <- randeff + mean_adjustment
  
  # Clip values between 0 and randeff_max
  randeff[randeff < 0] <- 0
  randeff[randeff > randeff_max] <- randeff_max
  
  # Return the resulting genome vector
  return(randeff)
}

num_windows_all <- 1000
num_windows_per_seg_min <- 10
num_windows_per_seg_max <- 30  
randeff_max <- 5
num_seg <- 10

set.seed(2)
randeff <-generate_random_genome_vector(num_windows_all, num_windows_per_seg_min, num_windows_per_seg_max, randeff_max, num_seg)

plot(randeff, type="l", ylab = "random effect")

```


4. We sampled the $\tilde{y_i}$ from 

$$\tilde{y_i} \sim \text{pois}(\tilde{\theta_i}\mu_i)$$
where $\mu_i$ is the expected mutations for each window $i$ from Roulette. 

5. The $\tilde{y_i}$ and $\mu_i$ were used as inputs for  smashgen and the $\hat{\theta_i}$ was the output.

## Functions used

```{r} 
library(ggplot2)
plot_randeff <- function(data) {
  
  ggplot(data, aes(x = x)) +
    geom_line(aes(y = randeff_simulated, color = "Simulated"), alpha = 0.5) +
    geom_line(aes(y = randeff_est, color = "Estimated"), alpha = 0.7) +
    labs(x = "Window index",
         y = "random effects",
         color = "Type") +
    scale_color_manual(values = c("Simulated" = "blue", "Estimated" = "red")) +
    theme_minimal()
}

scatter_plot_randeff <- function(data) {
  # Fit the linear model
  fit <- lm(randeff_est ~ randeff_simulated, data = data)
  adj_rsq <- summary(fit)$adj.r.squared  # Extract the adjusted R-squared
  
  # Generate the plot
  ggplot(data, aes(x = randeff_simulated, y = randeff_est)) +
    geom_point(color = "black") +
    geom_abline(intercept = 0, slope = 1, color = "red") +
    geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
              x = Inf, y = -Inf, hjust = 1, vjust = 0,
              color = "blue") +
    labs(x = "True random effect theta_i",
         y = "Estimated random effect") +
    theme_minimal()
}

plot_dnm <- function(data, windowsize) {
   ggplot(data, aes(x = x)) +
    geom_point(aes(y = y_sample), color = "grey") +
    geom_line(aes(y = randeff_simulated), color = "blue") +
    geom_line(aes(y = randeff_est), color = "red", alpha = 0.5) +
    labs(x = "Window index",
         y = paste0("% of DNM/window (windowsize = ", as.numeric(windowsize) / 1000, "kb)")) +
    theme_minimal()
}


```

# Results

The estimated mutation rates are from Roulette chr1 

## 10kb

### maximum random effect = 3

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial/simu_spatial_chr1_10kb_maxiter50_maxrandeff3.rdata")
plot_randeff(data)
plot_dnm(data,windowsize)
scatter_plot_randeff(data)

```

### maximum random effect = 5

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial/simu_spatial_chr1_10kb_maxiter50_maxrandeff5.rdata")
plot_randeff(data)
plot_dnm(data,windowsize)
scatter_plot_randeff(data)

```




## 30kb

### maximum random effect = 3

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial/simu_spatial_chr1_30kb_maxiter50_maxrandeff3.rdata")
plot_randeff(data)
plot_dnm(data,windowsize)
scatter_plot_randeff(data)

```

### maximum random effect = 5

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial/simu_spatial_chr1_30kb_maxiter50_maxrandeff5.rdata")
plot_randeff(data)
plot_dnm(data,windowsize)
scatter_plot_randeff(data)

```



## 50kb

### maximum random effect = 3

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial/simu_spatial_chr1_50kb_maxiter50_maxrandeff3.rdata")
plot_randeff(data)
plot_dnm(data,windowsize)
scatter_plot_randeff(data)

```

### maximum random effect = 5

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial/simu_spatial_chr1_50kb_maxiter50_maxrandeff5.rdata")
plot_randeff(data)
plot_dnm(data,windowsize)
scatter_plot_randeff(data)

```



## 100kb

### maximum random effect = 3

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial/simu_spatial_chr1_100kb_maxiter50_maxrandeff3.rdata")
plot_randeff(data)
plot_dnm(data,windowsize)
scatter_plot_randeff(data)

```

### maximum random effect = 5

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial/simu_spatial_chr1_100kb_maxiter50_maxrandeff5.rdata")
plot_randeff(data)
plot_dnm(data,windowsize)
scatter_plot_randeff(data)

```

