---
title: "Data cleaning"
author: "XSun"
date: "2024-05-20"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
---

# Introduction

We checked the gaps from Roulette estimation here and get an overview for the observed fold change $r_i$ computed from [SSC cell 2022](https://xsun1229.github.io/mutation_rate/denovo_data_overview.html#SSC_from_Cell_2022)

```{r, message=F}
library(ggplot2)
library(kableExtra)
library(dplyr)
library(data.table)

create_filtered_plot <- function(data) {
  # Convert to data.table
  setDT(data)
  
  # Create group based on rleid
  data[, group := rleid(pass_threshold)]
  
  # Filter rows where pass_threshold is TRUE and the length of the group is greater than 1
  filtered_data <- data[pass_threshold == TRUE, .SD[.N > 1], by = group]
  
  # Remove the group column if not needed
  filtered_data[, group := NULL]
  
  # Create the plot
  plot <- ggplot(filtered_data, aes(x = Window_Start / 1000000, y = obs_sum / rl_rescaled_sum)) +
    geom_point() +
    labs(x = "Genomic position (mb)", y = "fold change r_i") +
    theme_minimal()
  
  return(plot)
}



```

# Gaps in Roulette estimation

The gaps in the roulette estimation are due to:

- **Long gaps:** Centromeres

- **Small gaps:**
  1. Low density of polymorphism in GNOMAD at 100 nucleotide scale
  2. Poor mappability issues and repetitive sequences (UCSC self chain track)

## chr1 

```{r}
load("/project/xinhe/xsun/mutation_rate/10.denovo_real/data_cleaning/ssc_trost_gap_summary_chr1.rdata")

ggplot(gap_summary, aes(x = Start/1000000, y = Length)) +
  geom_point() +
  labs(title = "Gap Length and Position",
       x = "Start Position (mb)",
       y = "Gap Length") +
  theme_minimal()

print("summary for the gap length")
print(summary(gap_summary$Length))

load("/project/xinhe/xsun/mutation_rate/10.denovo_real/data_cleaning/ssc_trost_per_window_chr1_100kb_frac4.rdata")
pct <- sum(df_per_window_data$Total_Gap_Length)/(as.numeric(df_per_window_data$Window_End[nrow(df_per_window_data)] - as.numeric(df_per_window_data$Window_Start[1])))
print(sprintf("The percentage of positions with missing Roulette estimations is %.5f%%", pct*10))

gap_table <- table(gap_summary$Length)
gap_df <- as.data.frame(gap_table)
colnames(gap_df) <- c("Length", "Frequency")

# Convert Length to numeric
gap_df$Length <- as.numeric(as.character(gap_df$Length))

# Bin the data by 1000
gap_df <- gap_df %>%
  mutate(Bin = cut(Length, breaks = seq(0, max(Length) + 1000, by = 1000), right = FALSE)) %>%
  group_by(Bin) %>%
  summarise(Frequency = sum(Frequency)) %>%
  ungroup()

# Display the table using kable
kable(gap_df, format = "markdown", caption = "Summary of Gap Lengths Binned by 1000") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)

index <- order(gap_summary$Length, decreasing = TRUE)[1:2]
gap_summary_topremoved <- gap_summary[-index,]
ggplot(gap_summary_topremoved, aes(x = Start/1000000, y = Length)) +
  geom_point() +
  labs(title = "Gap Length and Position, top 2 gaps removed",
       x = "Start Position (mb)",
       y = "Gap Length") +
  theme_minimal()
```


# Removing the gaps and computing the fold changes 

The data cleaning is performed as follows:

1. Partition Chromosome into Windows:

The chromosome is divided into fixed windows of various sizes (10kb, 30kb, 50kb, 100kb).

2. Compute Gaps within Windows:

For each window, the positions with missing roulette estimation (gaps) are computed.

3. Filter Windows Based on Gaps:

Windows with less than 25% gaps are retained for further analysis.

4. Summarize Data within Windows:

For each window, the sum of roulette estimation and the observed de novo mutation count from SSC cell 2022 are calculated.

5. Split Data for Cross-Validation:

Each window's positions are divided into a training set (first 90%) and a testing set (remaining 10%) for cross-validation.

## An example for chr1, 100kb window size

```{r}

load("/project/xinhe/xsun/mutation_rate/10.denovo_real/data_cleaning/ssc_trost_per_window_chr1_100kb_frac4.rdata")

DT::datatable(df_per_window_data,caption = htmltools::tags$caption(style = 'caption-side: left; text-align: left; color:black; font-size:150%;',''),options = list(pageLength = 10))

```

Other window sizes also have similar patterns. 

## fold change plots for chr1

We just select rows where the "TRUE" value in the pass_threshold column is continuous for more than one row.

### 10kb

```{r fig.width=10, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/10.denovo_real/data_cleaning/ssc_trost_per_window_chr1_10kb_frac4.rdata")

# Example usage
plot <- create_filtered_plot(df_per_window_data)
print(plot)
```


### 30kb

```{r fig.width=10, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/10.denovo_real/data_cleaning/ssc_trost_per_window_chr1_30kb_frac4.rdata")

# Example usage
plot <- create_filtered_plot(df_per_window_data)
print(plot)
```


### 50kb

```{r fig.width=10, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/10.denovo_real/data_cleaning/ssc_trost_per_window_chr1_50kb_frac4.rdata")

# Example usage
plot <- create_filtered_plot(df_per_window_data)
print(plot)
```


### 100kb

```{r fig.width=10, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/10.denovo_real/data_cleaning/ssc_trost_per_window_chr1_100kb_frac4.rdata")

# Example usage
plot <- create_filtered_plot(df_per_window_data)
print(plot)
```






## fold change plots for chr22

We just select rows where the "TRUE" value in the pass_threshold column is continuous for more than one row.

### 10kb

```{r fig.width=10, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/10.denovo_real/data_cleaning/ssc_trost_per_window_chr22_10kb_frac4.rdata")

# Example usage
plot <- create_filtered_plot(df_per_window_data)
print(plot)
```


### 30kb

```{r fig.width=10, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/10.denovo_real/data_cleaning/ssc_trost_per_window_chr22_30kb_frac4.rdata")

# Example usage
plot <- create_filtered_plot(df_per_window_data)
print(plot)
```


### 50kb

```{r fig.width=10, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/10.denovo_real/data_cleaning/ssc_trost_per_window_chr22_50kb_frac4.rdata")

# Example usage
plot <- create_filtered_plot(df_per_window_data)
print(plot)
```


### 100kb

```{r fig.width=10, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/10.denovo_real/data_cleaning/ssc_trost_per_window_chr22_100kb_frac4.rdata")

# Example usage
plot <- create_filtered_plot(df_per_window_data)
print(plot)
```




