---
title: "Simulation -- smashgen"
author: "XSun"
date: "2023-08-04"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r echo=F,message=FALSE}
library(ggplot2)
library(gridExtra)
library(data.table)
```

# Introduction

1.  We set fold change $\tilde{\theta_i}$ from uniform distribution for each window $i$, the mean value of $\tilde{\theta_i}$ is 1.

The window sizes for the local fold changes range from hundreds kb to 1mb.

Below is an example for the fold change

```{r collapse=TRUE, eval=T ,message=FALSE, fig.width=12, fig.height=5}
generate_random_genome_vector <- function(num_windows_all, num_windows_per_seg_min, num_windows_per_seg_max, randeff_max, num_seg) {
  # Generate random segment lengths
  
  set.seed(1)
  segment_lengths <- ceiling(runif(num_seg, min = num_windows_per_seg_min, max = num_windows_per_seg_max))
  
  # Calculate the total length covered by segments
  total_segment_length <- sum(segment_lengths)
  
  # Check if the total segment length is greater than the genome length
  # If so, reduce the last segment length to fit within the genome
  if (total_segment_length > num_windows_all) {
    segment_lengths[num_seg] <- segment_lengths[num_seg] - (total_segment_length - num_windows_all)
  }
  
  # Calculate the number of zeros needed
  num_zeros <- num_windows_all - total_segment_length
  
  # Randomly select positions for segment starts
  segment_starts <- sample(1:num_zeros, num_seg, replace = FALSE)
  
  # Create the vector
  randeff <- numeric(num_windows_all)
  
  # Assign random values to segments
  for (i in 1:num_seg) {
    start_pos <- segment_starts[i]
    end_pos <- start_pos + segment_lengths[i] - 1
    randeff[start_pos:end_pos] <- runif(1, 0, randeff_max)
  }
  
  # Adjust values to meet the mean of 1
  mean_adjustment <- 1 - mean(randeff)
  randeff <- randeff + mean_adjustment
  
  # Clip values between 0 and randeff_max
  randeff[randeff < 0] <- 0
  randeff[randeff > randeff_max] <- randeff_max
  
  # Return the resulting genome vector
  return(randeff)
}

num_windows_all <- 1000
num_windows_per_seg_min <- 10
num_windows_per_seg_max <- 100  
randeff_max <- 5
num_seg <- 10

set.seed(1)
randeff <-generate_random_genome_vector(num_windows_all, num_windows_per_seg_min, num_windows_per_seg_max, randeff_max, num_seg)

plot(randeff, type="l", ylab = "fold change")

```

2. We sample the $\tilde{y_i}$ from 

$$\tilde{y_i} \sim pois(\tilde{\theta_i}*\mu_i)$$
where $\mu_i$ is the expected mutations for each window $i$ from Roulette. 

```{r echo=T, eval=T ,message=FALSE, fig.width=12, fig.height=5}
sample_poisson_for_each_lambda <- function(lambda_vector) {
  return(rpois(length(lambda_vector), lambda_vector))
}
```

3. We fit the $\tilde{y_i}$ and $\mu_i$ using smashgen and plot the $\tilde{\theta_i}$ and smoothed $\tilde{\theta}_i{_{smoothed}}$


# smashgen, SSC (Cell 2022, chr1) -- sample the random effects from uniform distribution


## 10kb



```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_10kb_maxiter50_maxrandeff3_round2.rdata")

print("maximum fold change = 3")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

num_windows_all <- 10000
dat_rl <- fread("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr1/rl_ssc_trost_combine_chr1_10000.txt")
colnames(dat_rl) <- c("chr", "start", "end","obs","exp")
dat_rl <- dat_rl[1:num_windows_all,]

data <- cbind(data,dat_rl)

fit1 <- lm(randeff_smooth ~ randeff_simulated + 0, data=data)
adj_rsq1 <- summary(fit1)$adj.r.squared
p1 <- ggplot(data, aes(x = randeff_simulated, y = randeff_smooth)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq1, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()

fit2 <- lm(y_sample/exp ~ randeff_simulated + 0, data=data)
adj_rsq2 <- summary(fit2)$adj.r.squared
p2 <- ggplot(data, aes(x = randeff_simulated, y = y_sample/exp)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq2, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "fold change (#y_sample/#exp)") +
  theme_minimal()
grid.arrange(p1, p2, ncol = 2)
```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_10kb_maxiter50_maxrandeff5_round2.rdata")

print("maximum fold change = 5")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()


num_windows_all <- 10000
dat_rl <- fread("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr1/rl_ssc_trost_combine_chr1_10000.txt")
colnames(dat_rl) <- c("chr", "start", "end","obs","exp")
dat_rl <- dat_rl[1:num_windows_all,]

data <- cbind(data,dat_rl)

fit1 <- lm(randeff_smooth ~ randeff_simulated + 0, data=data)
adj_rsq1 <- summary(fit1)$adj.r.squared
p1 <- ggplot(data, aes(x = randeff_simulated, y = randeff_smooth)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq1, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()

fit2 <- lm(y_sample/exp ~ randeff_simulated + 0, data=data)
adj_rsq2 <- summary(fit2)$adj.r.squared
p2 <- ggplot(data, aes(x = randeff_simulated, y = y_sample/exp)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq2, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "fold change (#y_sample/#exp)") +
  theme_minimal()
grid.arrange(p1, p2, ncol = 2)
```



## 30kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_30kb_maxiter50_maxrandeff3_round2.rdata")

print("maximum fold change = 3")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

num_windows_all <- 5000
dat_rl <- fread("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr1/rl_ssc_trost_combine_chr1_30000.txt")
colnames(dat_rl) <- c("chr", "start", "end","obs","exp")
dat_rl <- dat_rl[1:num_windows_all,]

data <- cbind(data,dat_rl)

fit1 <- lm(randeff_smooth ~ randeff_simulated + 0, data=data)
adj_rsq1 <- summary(fit1)$adj.r.squared
p1 <- ggplot(data, aes(x = randeff_simulated, y = randeff_smooth)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq1, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()

fit2 <- lm(y_sample/exp ~ randeff_simulated + 0, data=data)
adj_rsq2 <- summary(fit2)$adj.r.squared
p2 <- ggplot(data, aes(x = randeff_simulated, y = y_sample/exp)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq2, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "fold change (#y_sample/#exp)") +
  theme_minimal()
grid.arrange(p1, p2, ncol = 2)
```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_30kb_maxiter50_maxrandeff5_round2.rdata")

print("maximum fold change = 5")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()


num_windows_all <- 5000
dat_rl <- fread("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr1/rl_ssc_trost_combine_chr1_30000.txt")
colnames(dat_rl) <- c("chr", "start", "end","obs","exp")
dat_rl <- dat_rl[1:num_windows_all,]

data <- cbind(data,dat_rl)

fit1 <- lm(randeff_smooth ~ randeff_simulated + 0, data=data)
adj_rsq1 <- summary(fit1)$adj.r.squared
p1 <- ggplot(data, aes(x = randeff_simulated, y = randeff_smooth)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq1, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()

fit2 <- lm(y_sample/exp ~ randeff_simulated + 0, data=data)
adj_rsq2 <- summary(fit2)$adj.r.squared
p2 <- ggplot(data, aes(x = randeff_simulated, y = y_sample/exp)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq2, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "fold change (#y_sample/#exp)") +
  theme_minimal()
grid.arrange(p1, p2, ncol = 2)
```



## 50kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_50kb_maxiter50_maxrandeff3_round2.rdata")

print("maximum fold change = 3")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()


num_windows_all <- 3000
dat_rl <- fread("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr1/rl_ssc_trost_combine_chr1_50000.txt")
colnames(dat_rl) <- c("chr", "start", "end","obs","exp")
dat_rl <- dat_rl[1:num_windows_all,]

data <- cbind(data,dat_rl)

fit1 <- lm(randeff_smooth ~ randeff_simulated + 0, data=data)
adj_rsq1 <- summary(fit1)$adj.r.squared
p1 <- ggplot(data, aes(x = randeff_simulated, y = randeff_smooth)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq1, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()

fit2 <- lm(y_sample/exp ~ randeff_simulated + 0, data=data)
adj_rsq2 <- summary(fit2)$adj.r.squared
p2 <- ggplot(data, aes(x = randeff_simulated, y = y_sample/exp)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq2, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "fold change (#y_sample/#exp)") +
  theme_minimal()
grid.arrange(p1, p2, ncol = 2)
```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_50kb_maxiter50_maxrandeff5_round2.rdata")

print("maximum fold change = 5")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

num_windows_all <- 3000
dat_rl <- fread("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr1/rl_ssc_trost_combine_chr1_50000.txt")
colnames(dat_rl) <- c("chr", "start", "end","obs","exp")
dat_rl <- dat_rl[1:num_windows_all,]

data <- cbind(data,dat_rl)

fit1 <- lm(randeff_smooth ~ randeff_simulated + 0, data=data)
adj_rsq1 <- summary(fit1)$adj.r.squared
p1 <- ggplot(data, aes(x = randeff_simulated, y = randeff_smooth)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq1, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()

fit2 <- lm(y_sample/exp ~ randeff_simulated + 0, data=data)
adj_rsq2 <- summary(fit2)$adj.r.squared
p2 <- ggplot(data, aes(x = randeff_simulated, y = y_sample/exp)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq2, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "fold change (#y_sample/#exp)") +
  theme_minimal()
grid.arrange(p1, p2, ncol = 2)
```



## 100kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_100kb_maxiter50_maxrandeff3_round2.rdata")

print("maximum fold change = 3")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

num_windows_all <- 1000
dat_rl <- fread("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr1/rl_ssc_trost_combine_chr1_100000.txt")
colnames(dat_rl) <- c("chr", "start", "end","obs","exp")
dat_rl <- dat_rl[1:num_windows_all,]

data <- cbind(data,dat_rl)

fit1 <- lm(randeff_smooth ~ randeff_simulated + 0, data=data)
adj_rsq1 <- summary(fit1)$adj.r.squared
p1 <- ggplot(data, aes(x = randeff_simulated, y = randeff_smooth)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq1, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()

fit2 <- lm(y_sample/exp ~ randeff_simulated + 0, data=data)
adj_rsq2 <- summary(fit2)$adj.r.squared
p2 <- ggplot(data, aes(x = randeff_simulated, y = y_sample/exp)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq2, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "fold change (#y_sample/#exp)") +
  theme_minimal()
grid.arrange(p1, p2, ncol = 2)

```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_100kb_maxiter50_maxrandeff5_round2.rdata")

print("maximum fold change = 5")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

num_windows_all <- 1000
dat_rl <- fread("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr1/rl_ssc_trost_combine_chr1_100000.txt")
colnames(dat_rl) <- c("chr", "start", "end","obs","exp")
dat_rl <- dat_rl[1:num_windows_all,]

data <- cbind(data,dat_rl)

fit1 <- lm(randeff_smooth ~ randeff_simulated + 0, data=data)
adj_rsq1 <- summary(fit1)$adj.r.squared
p1 <- ggplot(data, aes(x = randeff_simulated, y = randeff_smooth)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq1, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()

fit2 <- lm(y_sample/exp ~ randeff_simulated + 0, data=data)
adj_rsq2 <- summary(fit2)$adj.r.squared
p2 <- ggplot(data, aes(x = randeff_simulated, y = y_sample/exp)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq2, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "fold change (#y_sample/#exp)") +
  theme_minimal()
grid.arrange(p1, p2, ncol = 2)
```







# Checking if smashgen works on non-spatial data, SSC (Cell 2022, chr1)

## 10kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_10kb_maxiter50_maxrandeff3_nonspatial.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

fit <- lm(randeff_smooth ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_smooth)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_10kb_maxiter50_maxrandeff5_nonspatial.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

fit <- lm(randeff_smooth ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_smooth)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```



## 30kb


```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_30kb_maxiter50_maxrandeff3_nonspatial.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

fit <- lm(randeff_smooth ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_smooth)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_30kb_maxiter50_maxrandeff5_nonspatial.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

fit <- lm(randeff_smooth ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_smooth)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```



## 50kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_50kb_maxiter50_maxrandeff3_nonspatial.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

fit <- lm(randeff_smooth ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_smooth)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_50kb_maxiter50_maxrandeff5_nonspatial.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

fit <- lm(randeff_smooth ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_smooth)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```



## 100kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_100kb_maxiter50_maxrandeff3_nonspatial.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

fit <- lm(randeff_smooth ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_smooth)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_100kb_maxiter50_maxrandeff5_nonspatial.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

fit <- lm(randeff_smooth ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_smooth)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```






# smashgen, SSC (Cell 2022, chr1) -- use the random effects from real data

## 10kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_10kb_maxiter50_realrandeff.rdata")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_real), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

num_windows_all <- 10000
dat_rl <- fread("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr1/rl_ssc_trost_combine_chr1_10000.txt")
colnames(dat_rl) <- c("chr", "start", "end","obs","exp")
dat_rl <- dat_rl[1:num_windows_all,]

data <- cbind(data,dat_rl)

fit1 <- lm(randeff_smooth ~ randeff_real + 0, data=data)
adj_rsq1 <- summary(fit1)$adj.r.squared
p1 <- ggplot(data, aes(x = randeff_real, y = randeff_smooth)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq1, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()

fit2 <- lm(y_sample/exp ~ randeff_real + 0, data=data)
adj_rsq2 <- summary(fit2)$adj.r.squared
p2 <- ggplot(data, aes(x = randeff_real, y = y_sample/exp)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq2, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "fold change (#y_sample/#exp)") +
  theme_minimal()
grid.arrange(p1, p2, ncol = 2)
```





## 30kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_30kb_maxiter50_realrandeff.rdata")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_real), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

num_windows_all <- 5000
dat_rl <- fread("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr1/rl_ssc_trost_combine_chr1_30000.txt")
colnames(dat_rl) <- c("chr", "start", "end","obs","exp")
dat_rl <- dat_rl[1:num_windows_all,]

data <- cbind(data,dat_rl)

fit1 <- lm(randeff_smooth ~ randeff_real + 0, data=data)
adj_rsq1 <- summary(fit1)$adj.r.squared
p1 <- ggplot(data, aes(x = randeff_real, y = randeff_smooth)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq1, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()

fit2 <- lm(y_sample/exp ~ randeff_real + 0, data=data)
adj_rsq2 <- summary(fit2)$adj.r.squared
p2 <- ggplot(data, aes(x = randeff_real, y = y_sample/exp)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq2, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "fold change (#y_sample/#exp)") +
  theme_minimal()
grid.arrange(p1, p2, ncol = 2)
```





## 50kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_50kb_maxiter50_realrandeff.rdata")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_real), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

num_windows_all <- 3000
dat_rl <- fread("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr1/rl_ssc_trost_combine_chr1_50000.txt")
colnames(dat_rl) <- c("chr", "start", "end","obs","exp")
dat_rl <- dat_rl[1:num_windows_all,]

data <- cbind(data,dat_rl)

fit1 <- lm(randeff_smooth ~ randeff_real + 0, data=data)
adj_rsq1 <- summary(fit1)$adj.r.squared
p1 <- ggplot(data, aes(x = randeff_real, y = randeff_smooth)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq1, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()

fit2 <- lm(y_sample/exp ~ randeff_real + 0, data=data)
adj_rsq2 <- summary(fit2)$adj.r.squared
p2 <- ggplot(data, aes(x = randeff_real, y = y_sample/exp)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq2, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "fold change (#y_sample/#exp)") +
  theme_minimal()
grid.arrange(p1, p2, ncol = 2)
```





## 100kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_100kb_maxiter50_realrandeff.rdata")

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_real), color = "blue") +
  geom_line(aes(y = randeff_smooth), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb)")) +
  theme_minimal()

num_windows_all <-1000
dat_rl <- fread("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr1/rl_ssc_trost_combine_chr1_100000.txt")
colnames(dat_rl) <- c("chr", "start", "end","obs","exp")
dat_rl <- dat_rl[1:num_windows_all,]

data <- cbind(data,dat_rl)

fit1 <- lm(randeff_smooth ~ randeff_real + 0, data=data)
adj_rsq1 <- summary(fit1)$adj.r.squared
p1 <- ggplot(data, aes(x = randeff_real, y = randeff_smooth)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq1, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()

fit2 <- lm(y_sample/exp ~ randeff_real + 0, data=data)
adj_rsq2 <- summary(fit2)$adj.r.squared
p2 <- ggplot(data, aes(x = randeff_real, y = y_sample/exp)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq2, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "fold change (#y_sample/#exp)") +
  theme_minimal()
grid.arrange(p1, p2, ncol = 2)
```



