---
title: "Simulation -- non-spatial, gamma distribution"
author: "XSun"
date: "2023-08-24"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r echo=F,message=FALSE}
library(ggplot2)
library(gridExtra)
library(data.table)
```

# Introduction

1.  We set fold change $\tilde{\theta_i}$ from uniform distribution for each window $i$, the mean value of $\tilde{\theta_i}$ is 1.

The window sizes for the local fold changes range from hundreds kb to 1mb.
 
2.  We sample the $\tilde{y_i}$ from 

$$\tilde{y_i} \sim pois(\tilde{\theta_i}*\mu_i)$$
where $\mu_i$ is the expected mutations for each window $i$ from Roulette. 

3. We can use gamma distribution to model $\theta_i$ :

$$\theta_i \sim Gamma(\alpha, \beta)$$ Since the $\mu_i$ is calibrated (the total mutation rate is given), we just have single overdispersion parameter

so, the simple non-spatial random effect model is

$$\theta_i \sim Gamma(\alpha, \alpha)$$ So the likelihood is :

$$\begin{aligned}
Pr(y_{i} |\mu_{i},\alpha) &= \int_0^{+\infty}
Pr(y_{i}|\mu_{i},\theta_{i})Pr(\theta_{i}|\alpha)\,d\theta_{i} \\
\end{aligned}$$

which can be optimized by MLE with "BFGS" method.

```{r echo=T, message=FALSE, warning=FALSE}

# Input is a dataframe of observed rates and expected rates
est.alpha <- function(dat) {
  mu_pos <- dat$exp > 0
  y_pos <- dat$obs > 0
  dat.mu_pos <- dat[(!y_pos)&mu_pos,]
  dat.y_pos <- dat[y_pos,]
  loglik <- function(alpha) {
    loglik.mu_pos <- alpha*(log(alpha) - log(dat.mu_pos$exp+alpha))
    loglik.y_pos <- log(gamma(dat.y_pos$obs+alpha)) - log(gamma(alpha)) + alpha*log(alpha) - (dat.y_pos$obs+alpha)*log(dat.y_pos$exp+alpha)
    -sum(loglik.mu_pos) - sum(loglik.y_pos)
  }
  res <- optim(par=1, fn=loglik, method="BFGS") # May need to try different initiations and methods
  res$par # MLE of alpha
}

```

The posterior distribution of $ \theta_i $ is :

$$ \theta_{i} | y_i,\mu_i,\alpha \sim Gamma( y_i+\alpha, \mu_i+\alpha)$$ 
So the posterior mean of $ \theta_i $ is:

$$E(\theta_{i} | y_i,\mu_i,\alpha) = \dfrac{y_i+\alpha}{\mu_i+\alpha} $$

# Simulation - more local random effects

## 1kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_1kb_maxrandeff3_nonspatial_gamma_more.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))
print(sprintf("alpha = %s", alpha))
ggplot(data, aes(x = x)) +
  geom_line(aes(y = randeff_simulated), color = "blue",alpha=0.5) +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.7) +
  labs(x = "Window index",
       y = paste0("random effects")) +
  theme_minimal()

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb) & random effects")) +
  theme_minimal()

fit <- lm(randeff_est ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_est)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_1kb_maxrandeff5_nonspatial_gamma_more.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))
print(sprintf("alpha = %s", alpha))
ggplot(data, aes(x = x)) +
  geom_line(aes(y = randeff_simulated), color = "blue",alpha=0.5) +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.7) +
  labs(x = "Window index",
       y = paste0("random effects")) +
  theme_minimal()

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb) & random effects")) +
  theme_minimal()

fit <- lm(randeff_est ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_est)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_1kb_maxrandeff10_nonspatial_gamma_more.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))
print(sprintf("alpha = %s", alpha))
ggplot(data, aes(x = x)) +
  geom_line(aes(y = randeff_simulated), color = "blue",alpha=0.5) +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.7) +
  labs(x = "Window index",
       y = paste0("random effects")) +
  theme_minimal()

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb) & random effects")) +
  theme_minimal()

fit <- lm(randeff_est ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_est)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```





## 10kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_10kb_maxrandeff3_nonspatial_gamma_more.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))
print(sprintf("alpha = %s", alpha))
ggplot(data, aes(x = x)) +
  geom_line(aes(y = randeff_simulated), color = "blue",alpha=0.5) +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.7) +
  labs(x = "Window index",
       y = paste0("random effects")) +
  theme_minimal()

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb) & random effects")) +
  theme_minimal()

fit <- lm(randeff_est ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_est)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_10kb_maxrandeff5_nonspatial_gamma_more.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))
print(sprintf("alpha = %s", alpha))
ggplot(data, aes(x = x)) +
  geom_line(aes(y = randeff_simulated), color = "blue",alpha=0.5) +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.7) +
  labs(x = "Window index",
       y = paste0("random effects")) +
  theme_minimal()

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb) & random effects")) +
  theme_minimal()

fit <- lm(randeff_est ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_est)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_10kb_maxrandeff10_nonspatial_gamma_more.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))
print(sprintf("alpha = %s", alpha))
ggplot(data, aes(x = x)) +
  geom_line(aes(y = randeff_simulated), color = "blue",alpha=0.5) +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.7) +
  labs(x = "Window index",
       y = paste0("random effects")) +
  theme_minimal()

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb) & random effects")) +
  theme_minimal()

fit <- lm(randeff_est ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_est)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```




## 30kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_30kb_maxrandeff3_nonspatial_gamma_more.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))
print(sprintf("alpha = %s", alpha))
ggplot(data, aes(x = x)) +
  geom_line(aes(y = randeff_simulated), color = "blue",alpha=0.5) +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.7) +
  labs(x = "Window index",
       y = paste0("random effects")) +
  theme_minimal()

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb) & random effects")) +
  theme_minimal()

fit <- lm(randeff_est ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_est)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_30kb_maxrandeff5_nonspatial_gamma_more.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))
print(sprintf("alpha = %s", alpha))
ggplot(data, aes(x = x)) +
  geom_line(aes(y = randeff_simulated), color = "blue",alpha=0.5) +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.7) +
  labs(x = "Window index",
       y = paste0("random effects")) +
  theme_minimal()

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb) & random effects")) +
  theme_minimal()

fit <- lm(randeff_est ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_est)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_30kb_maxrandeff10_nonspatial_gamma_more.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))
print(sprintf("alpha = %s", alpha))
ggplot(data, aes(x = x)) +
  geom_line(aes(y = randeff_simulated), color = "blue",alpha=0.5) +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.7) +
  labs(x = "Window index",
       y = paste0("random effects")) +
  theme_minimal()

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb) & random effects")) +
  theme_minimal()

fit <- lm(randeff_est ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_est)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```




## 50kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_50kb_maxrandeff3_nonspatial_gamma_more.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))
print(sprintf("alpha = %s", alpha))
ggplot(data, aes(x = x)) +
  geom_line(aes(y = randeff_simulated), color = "blue",alpha=0.5) +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.7) +
  labs(x = "Window index",
       y = paste0("random effects")) +
  theme_minimal()

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb) & random effects")) +
  theme_minimal()

fit <- lm(randeff_est ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_est)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_50kb_maxrandeff5_nonspatial_gamma_more.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))
print(sprintf("alpha = %s", alpha))
ggplot(data, aes(x = x)) +
  geom_line(aes(y = randeff_simulated), color = "blue",alpha=0.5) +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.7) +
  labs(x = "Window index",
       y = paste0("random effects")) +
  theme_minimal()

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb) & random effects")) +
  theme_minimal()

fit <- lm(randeff_est ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_est)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```


```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_50kb_maxrandeff10_nonspatial_gamma_more.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))
print(sprintf("alpha = %s", alpha))
ggplot(data, aes(x = x)) +
  geom_line(aes(y = randeff_simulated), color = "blue",alpha=0.5) +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.7) +
  labs(x = "Window index",
       y = paste0("random effects")) +
  theme_minimal()

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb) & random effects")) +
  theme_minimal()

fit <- lm(randeff_est ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_est)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```



## 100kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_100kb_maxrandeff3_nonspatial_gamma_more.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))
print(sprintf("alpha = %s", alpha))
ggplot(data, aes(x = x)) +
  geom_line(aes(y = randeff_simulated), color = "blue",alpha=0.5) +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.7) +
  labs(x = "Window index",
       y = paste0("random effects")) +
  theme_minimal()

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb) & random effects")) +
  theme_minimal()

fit <- lm(randeff_est ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_est)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_100kb_maxrandeff5_nonspatial_gamma_more.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))
print(sprintf("alpha = %s", alpha))
ggplot(data, aes(x = x)) +
  geom_line(aes(y = randeff_simulated), color = "blue",alpha=0.5) +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.7) +
  labs(x = "Window index",
       y = paste0("random effects")) +
  theme_minimal()

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb) & random effects")) +
  theme_minimal()

fit <- lm(randeff_est ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_est)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_100kb_maxrandeff10_nonspatial_gamma_more.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))
print(sprintf("alpha = %s", alpha))
ggplot(data, aes(x = x)) +
  geom_line(aes(y = randeff_simulated), color = "blue",alpha=0.5) +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.7) +
  labs(x = "Window index",
       y = paste0("random effects")) +
  theme_minimal()

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb) & random effects")) +
  theme_minimal()

fit <- lm(randeff_est ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_est)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```



# Simulation

## 1kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_1kb_maxrandeff3_nonspatial_gamma.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))
print(sprintf("alpha = %s", alpha))
ggplot(data, aes(x = x)) +
  geom_line(aes(y = randeff_simulated), color = "blue",alpha=0.5) +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.7) +
  labs(x = "Window index",
       y = paste0("random effects")) +
  theme_minimal()

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb) & random effects")) +
  theme_minimal()

fit <- lm(randeff_est ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_est)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_1kb_maxrandeff5_nonspatial_gamma.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))
print(sprintf("alpha = %s", alpha))
ggplot(data, aes(x = x)) +
  geom_line(aes(y = randeff_simulated), color = "blue",alpha=0.5) +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.7) +
  labs(x = "Window index",
       y = paste0("random effects")) +
  theme_minimal()

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb) & random effects")) +
  theme_minimal()

fit <- lm(randeff_est ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_est)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_1kb_maxrandeff10_nonspatial_gamma.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))
print(sprintf("alpha = %s", alpha))
ggplot(data, aes(x = x)) +
  geom_line(aes(y = randeff_simulated), color = "blue",alpha=0.5) +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.7) +
  labs(x = "Window index",
       y = paste0("random effects")) +
  theme_minimal()

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb) & random effects")) +
  theme_minimal()

fit <- lm(randeff_est ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_est)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```





## 10kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_10kb_maxrandeff3_nonspatial_gamma.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))
print(sprintf("alpha = %s", alpha))
ggplot(data, aes(x = x)) +
  geom_line(aes(y = randeff_simulated), color = "blue",alpha=0.5) +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.7) +
  labs(x = "Window index",
       y = paste0("random effects")) +
  theme_minimal()

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb) & random effects")) +
  theme_minimal()

fit <- lm(randeff_est ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_est)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_10kb_maxrandeff5_nonspatial_gamma.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))
print(sprintf("alpha = %s", alpha))
ggplot(data, aes(x = x)) +
  geom_line(aes(y = randeff_simulated), color = "blue",alpha=0.5) +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.7) +
  labs(x = "Window index",
       y = paste0("random effects")) +
  theme_minimal()

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb) & random effects")) +
  theme_minimal()

fit <- lm(randeff_est ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_est)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_10kb_maxrandeff10_nonspatial_gamma.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))
print(sprintf("alpha = %s", alpha))
ggplot(data, aes(x = x)) +
  geom_line(aes(y = randeff_simulated), color = "blue",alpha=0.5) +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.7) +
  labs(x = "Window index",
       y = paste0("random effects")) +
  theme_minimal()

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb) & random effects")) +
  theme_minimal()

fit <- lm(randeff_est ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_est)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```




## 30kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_30kb_maxrandeff3_nonspatial_gamma.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))
print(sprintf("alpha = %s", alpha))
ggplot(data, aes(x = x)) +
  geom_line(aes(y = randeff_simulated), color = "blue",alpha=0.5) +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.7) +
  labs(x = "Window index",
       y = paste0("random effects")) +
  theme_minimal()

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb) & random effects")) +
  theme_minimal()

fit <- lm(randeff_est ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_est)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_30kb_maxrandeff5_nonspatial_gamma.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))
print(sprintf("alpha = %s", alpha))
ggplot(data, aes(x = x)) +
  geom_line(aes(y = randeff_simulated), color = "blue",alpha=0.5) +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.7) +
  labs(x = "Window index",
       y = paste0("random effects")) +
  theme_minimal()

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb) & random effects")) +
  theme_minimal()

fit <- lm(randeff_est ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_est)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_30kb_maxrandeff10_nonspatial_gamma.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))
print(sprintf("alpha = %s", alpha))
ggplot(data, aes(x = x)) +
  geom_line(aes(y = randeff_simulated), color = "blue",alpha=0.5) +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.7) +
  labs(x = "Window index",
       y = paste0("random effects")) +
  theme_minimal()

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb) & random effects")) +
  theme_minimal()

fit <- lm(randeff_est ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_est)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```




## 50kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_50kb_maxrandeff3_nonspatial_gamma.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))
print(sprintf("alpha = %s", alpha))
ggplot(data, aes(x = x)) +
  geom_line(aes(y = randeff_simulated), color = "blue",alpha=0.5) +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.7) +
  labs(x = "Window index",
       y = paste0("random effects")) +
  theme_minimal()

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb) & random effects")) +
  theme_minimal()

fit <- lm(randeff_est ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_est)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_50kb_maxrandeff5_nonspatial_gamma.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))
print(sprintf("alpha = %s", alpha))
ggplot(data, aes(x = x)) +
  geom_line(aes(y = randeff_simulated), color = "blue",alpha=0.5) +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.7) +
  labs(x = "Window index",
       y = paste0("random effects")) +
  theme_minimal()

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb) & random effects")) +
  theme_minimal()

fit <- lm(randeff_est ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_est)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```


```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_50kb_maxrandeff10_nonspatial_gamma.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))
print(sprintf("alpha = %s", alpha))
ggplot(data, aes(x = x)) +
  geom_line(aes(y = randeff_simulated), color = "blue",alpha=0.5) +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.7) +
  labs(x = "Window index",
       y = paste0("random effects")) +
  theme_minimal()

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb) & random effects")) +
  theme_minimal()

fit <- lm(randeff_est ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_est)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```



## 100kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_100kb_maxrandeff3_nonspatial_gamma.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))
print(sprintf("alpha = %s", alpha))
ggplot(data, aes(x = x)) +
  geom_line(aes(y = randeff_simulated), color = "blue",alpha=0.5) +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.7) +
  labs(x = "Window index",
       y = paste0("random effects")) +
  theme_minimal()

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb) & random effects")) +
  theme_minimal()

fit <- lm(randeff_est ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_est)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_100kb_maxrandeff5_nonspatial_gamma.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))
print(sprintf("alpha = %s", alpha))
ggplot(data, aes(x = x)) +
  geom_line(aes(y = randeff_simulated), color = "blue",alpha=0.5) +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.7) +
  labs(x = "Window index",
       y = paste0("random effects")) +
  theme_minimal()

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb) & random effects")) +
  theme_minimal()

fit <- lm(randeff_est ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_est)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_100kb_maxrandeff10_nonspatial_gamma.rdata")

print(paste0("maximum fold change =" ,max(data$randeff_simulated)))
print(sprintf("alpha = %s", alpha))
ggplot(data, aes(x = x)) +
  geom_line(aes(y = randeff_simulated), color = "blue",alpha=0.5) +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.7) +
  labs(x = "Window index",
       y = paste0("random effects")) +
  theme_minimal()

ggplot(data, aes(x = x)) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_est), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = paste0("# of DNM/window (windowsize =",windowsize/1000,"kb) & random effects")) +
  theme_minimal()

fit <- lm(randeff_est ~ randeff_simulated + 0, data=data)
adj_rsq <- summary(fit)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_est)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "smoothed random effect") +
  theme_minimal()
```




# Checking if the algorithm recovers the correct alpha 

Use Gamma prior to sample $\theta_i$ in the simulation. Alpha was varied from 5 to 50. 

## 10kb



```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_10kb_alpha_init1.rdata")

data <- data.frame(
  alpha_true_all = alpha_true_all,
  alpha_lb_all = alpha_lb_all,
  alpha_cg_all = alpha_cg_all,
  alpha_bf_all = alpha_bf_all
)

# Create the plot
p1 <- ggplot(data, aes(x = alpha_true_all)) +
  geom_line(aes(y = alpha_lb_all, color = "L-BFGS-B")) +
  geom_line(aes(y = alpha_cg_all, color = "CG")) +
  geom_line(aes(y = alpha_bf_all, color = "BFGS")) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray") + 
  labs(x = "True Alpha", y = "Alpha estimated by different methods", title = "initial alpha = 1") +
  scale_color_manual(values = c("L-BFGS-B" = "red", "CG" = "green", "BFGS" = "blue")) +
  theme_minimal()

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_10kb_alpha_init3.rdata")

data <- data.frame(
  alpha_true_all = alpha_true_all,
  alpha_lb_all = alpha_lb_all,
  alpha_cg_all = alpha_cg_all,
  alpha_bf_all = alpha_bf_all
)

# Create the plot
p2 <- ggplot(data, aes(x = alpha_true_all)) +
  geom_line(aes(y = alpha_lb_all, color = "L-BFGS-B")) +
  geom_line(aes(y = alpha_cg_all, color = "CG")) +
  geom_line(aes(y = alpha_bf_all, color = "BFGS")) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray") + 
  labs(x = "True Alpha", y = "Alpha estimated by different methods", title = "initial alpha = 3") +
  scale_color_manual(values = c("L-BFGS-B" = "red", "CG" = "green", "BFGS" = "blue")) +
  theme_minimal()
  

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_10kb_alpha_init5.rdata")

data <- data.frame(
  alpha_true_all = alpha_true_all,
  alpha_lb_all = alpha_lb_all,
  alpha_cg_all = alpha_cg_all,
  alpha_bf_all = alpha_bf_all
)

# Create the plot
p3 <- ggplot(data, aes(x = alpha_true_all)) +
  geom_line(aes(y = alpha_lb_all, color = "L-BFGS-B")) +
  geom_line(aes(y = alpha_cg_all, color = "CG")) +
  geom_line(aes(y = alpha_bf_all, color = "BFGS")) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray") + 
  labs(x = "True Alpha", y = "Alpha estimated by different methods", title = "initial alpha = 5") +
  scale_color_manual(values = c("L-BFGS-B" = "red", "CG" = "green", "BFGS" = "blue")) +
  theme_minimal()
grid.arrange(p1, p2,p3,ncol = 3)

```



## 30kb



```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_30kb_alpha_init1.rdata")

data <- data.frame(
  alpha_true_all = alpha_true_all,
  alpha_lb_all = alpha_lb_all,
  alpha_cg_all = alpha_cg_all,
  alpha_bf_all = alpha_bf_all
)

# Create the plot
p1 <- ggplot(data, aes(x = alpha_true_all)) +
  geom_line(aes(y = alpha_lb_all, color = "L-BFGS-B")) +
  geom_line(aes(y = alpha_cg_all, color = "CG")) +
  geom_line(aes(y = alpha_bf_all, color = "BFGS")) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray") + 
  labs(x = "True Alpha", y = "Alpha estimated by different methods", title = "initial alpha = 1") +
  scale_color_manual(values = c("L-BFGS-B" = "red", "CG" = "green", "BFGS" = "blue")) +
  theme_minimal()

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_30kb_alpha_init3.rdata")

data <- data.frame(
  alpha_true_all = alpha_true_all,
  alpha_lb_all = alpha_lb_all,
  alpha_cg_all = alpha_cg_all,
  alpha_bf_all = alpha_bf_all
)

# Create the plot
p2 <- ggplot(data, aes(x = alpha_true_all)) +
  geom_line(aes(y = alpha_lb_all, color = "L-BFGS-B")) +
  geom_line(aes(y = alpha_cg_all, color = "CG")) +
  geom_line(aes(y = alpha_bf_all, color = "BFGS")) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray") + 
  labs(x = "True Alpha", y = "Alpha estimated by different methods", title = "initial alpha = 3") +
  scale_color_manual(values = c("L-BFGS-B" = "red", "CG" = "green", "BFGS" = "blue")) +
  theme_minimal()
  

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_30kb_alpha_init5.rdata")

data <- data.frame(
  alpha_true_all = alpha_true_all,
  alpha_lb_all = alpha_lb_all,
  alpha_cg_all = alpha_cg_all,
  alpha_bf_all = alpha_bf_all
)

# Create the plot
p3 <- ggplot(data, aes(x = alpha_true_all)) +
  geom_line(aes(y = alpha_lb_all, color = "L-BFGS-B")) +
  geom_line(aes(y = alpha_cg_all, color = "CG")) +
  geom_line(aes(y = alpha_bf_all, color = "BFGS")) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray") + 
  labs(x = "True Alpha", y = "Alpha estimated by different methods", title = "initial alpha = 5") +
  scale_color_manual(values = c("L-BFGS-B" = "red", "CG" = "green", "BFGS" = "blue")) +
  theme_minimal()
grid.arrange(p1, p2,p3,ncol = 3)

```

## 50kb



```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_50kb_alpha_init1.rdata")

data <- data.frame(
  alpha_true_all = alpha_true_all,
  alpha_lb_all = alpha_lb_all,
  alpha_cg_all = alpha_cg_all,
  alpha_bf_all = alpha_bf_all
)

# Create the plot
p1 <- ggplot(data, aes(x = alpha_true_all)) +
  geom_line(aes(y = alpha_lb_all, color = "L-BFGS-B")) +
  geom_line(aes(y = alpha_cg_all, color = "CG")) +
  geom_line(aes(y = alpha_bf_all, color = "BFGS")) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray") + 
  labs(x = "True Alpha", y = "Alpha estimated by different methods", title = "initial alpha = 1") +
  scale_color_manual(values = c("L-BFGS-B" = "red", "CG" = "green", "BFGS" = "blue")) +
  theme_minimal()

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_50kb_alpha_init3.rdata")

data <- data.frame(
  alpha_true_all = alpha_true_all,
  alpha_lb_all = alpha_lb_all,
  alpha_cg_all = alpha_cg_all,
  alpha_bf_all = alpha_bf_all
)

# Create the plot
p2 <- ggplot(data, aes(x = alpha_true_all)) +
  geom_line(aes(y = alpha_lb_all, color = "L-BFGS-B")) +
  geom_line(aes(y = alpha_cg_all, color = "CG")) +
  geom_line(aes(y = alpha_bf_all, color = "BFGS")) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray") + 
  labs(x = "True Alpha", y = "Alpha estimated by different methods", title = "initial alpha = 3") +
  scale_color_manual(values = c("L-BFGS-B" = "red", "CG" = "green", "BFGS" = "blue")) +
  theme_minimal()
  

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_50kb_alpha_init5.rdata")

data <- data.frame(
  alpha_true_all = alpha_true_all,
  alpha_lb_all = alpha_lb_all,
  alpha_cg_all = alpha_cg_all,
  alpha_bf_all = alpha_bf_all
)

# Create the plot
p3 <- ggplot(data, aes(x = alpha_true_all)) +
  geom_line(aes(y = alpha_lb_all, color = "L-BFGS-B")) +
  geom_line(aes(y = alpha_cg_all, color = "CG")) +
  geom_line(aes(y = alpha_bf_all, color = "BFGS")) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray") + 
  labs(x = "True Alpha", y = "Alpha estimated by different methods", title = "initial alpha = 5") +
  scale_color_manual(values = c("L-BFGS-B" = "red", "CG" = "green", "BFGS" = "blue")) +
  theme_minimal()
grid.arrange(p1, p2,p3,ncol = 3)

```



## 100kb



```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_100kb_alpha_init1.rdata")

data <- data.frame(
  alpha_true_all = alpha_true_all,
  alpha_lb_all = alpha_lb_all,
  alpha_cg_all = alpha_cg_all,
  alpha_bf_all = alpha_bf_all
)

# Create the plot
p1 <- ggplot(data, aes(x = alpha_true_all)) +
  geom_line(aes(y = alpha_lb_all, color = "L-BFGS-B")) +
  geom_line(aes(y = alpha_cg_all, color = "CG")) +
  geom_line(aes(y = alpha_bf_all, color = "BFGS")) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray") + 
  labs(x = "True Alpha", y = "Alpha estimated by different methods", title = "initial alpha = 1") +
  scale_color_manual(values = c("L-BFGS-B" = "red", "CG" = "green", "BFGS" = "blue")) +
  theme_minimal()

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_100kb_alpha_init3.rdata")

data <- data.frame(
  alpha_true_all = alpha_true_all,
  alpha_lb_all = alpha_lb_all,
  alpha_cg_all = alpha_cg_all,
  alpha_bf_all = alpha_bf_all
)

# Create the plot
p2 <- ggplot(data, aes(x = alpha_true_all)) +
  geom_line(aes(y = alpha_lb_all, color = "L-BFGS-B")) +
  geom_line(aes(y = alpha_cg_all, color = "CG")) +
  geom_line(aes(y = alpha_bf_all, color = "BFGS")) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray") + 
  labs(x = "True Alpha", y = "Alpha estimated by different methods", title = "initial alpha = 3") +
  scale_color_manual(values = c("L-BFGS-B" = "red", "CG" = "green", "BFGS" = "blue")) +
  theme_minimal()
  

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_chr1_100kb_alpha_init5.rdata")

data <- data.frame(
  alpha_true_all = alpha_true_all,
  alpha_lb_all = alpha_lb_all,
  alpha_cg_all = alpha_cg_all,
  alpha_bf_all = alpha_bf_all
)

# Create the plot
p3 <- ggplot(data, aes(x = alpha_true_all)) +
  geom_line(aes(y = alpha_lb_all, color = "L-BFGS-B")) +
  geom_line(aes(y = alpha_cg_all, color = "CG")) +
  geom_line(aes(y = alpha_bf_all, color = "BFGS")) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray") + 
  labs(x = "True Alpha", y = "Alpha estimated by different methods", title = "initial alpha = 5") +
  scale_color_manual(values = c("L-BFGS-B" = "red", "CG" = "green", "BFGS" = "blue")) +
  theme_minimal()
grid.arrange(p1, p2,p3,ncol = 3)

```

