---
title: "Non-spatial random effect"
author: "XSun"
date: "2023-07-17"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

# Introduction

```{r echo=F, message=FALSE, warning=FALSE}
library(ggplot2)
```

We divide the whole genome into bins, in each window $i$, we have:

$$ y_i \sim  pois(\mu_i*\theta _i) $$ where $y_i$ is the observed de novo mutation counts. $\mu_i$ is expected mutation counts, $\theta_i$ is the non-spatial random effect in this window.

We can use gamma distribution to model $\theta_i$ :

$$\theta_i \sim Gamma(\alpha, \beta)$$ Since the $\mu_i$ is calibrated (the total mutation rate is given), we just have single overdispersion parameter

so, the simple non-spatial random effect model is

$$\theta_i \sim Gamma(\alpha, \alpha)$$
So the likelihood is :

$$\begin{aligned}
Pr(y_{i} |\mu_{i},\alpha) &= \int_0^{+\infty}
Pr(y_{i}|\mu_{i},\theta_{i})Pr(\theta_{i}|\alpha)\,d\theta_{i} \\
\end{aligned}$$

which can be optimized by MLE with “BFGS” method.

```{r echo=T, message=FALSE, warning=FALSE}

# Input is a dataframe of observed rates and expected rates
est.alpha <- function(dat) {
  mu_pos <- dat$exp > 0
  y_pos <- dat$obs > 0
  dat.mu_pos <- dat[(!y_pos)&mu_pos,]
  dat.y_pos <- dat[y_pos,]
  loglik <- function(alpha) {
    loglik.mu_pos <- alpha*(log(alpha) - log(dat.mu_pos$exp+alpha))
    loglik.y_pos <- log(gamma(dat.y_pos$obs+alpha)) - log(gamma(alpha)) + alpha*log(alpha) - (dat.y_pos$obs+alpha)*log(dat.y_pos$exp+alpha)
    -sum(loglik.mu_pos) - sum(loglik.y_pos)
  }
  res <- optim(par=1, fn=loglik, method="BFGS") # May need to try different initiations and methods
  res$par # MLE of alpha
}

```


The posterior distribution of $ \theta_i $ is :

$$ \theta_{i} | y_i,\mu_i,\alpha \sim Gamma( y_i+\alpha, \mu_i+\alpha)$$
So the posterior mean of $ \theta_i $ is: 

$$E(\theta_{i} | y_i,\mu_i,\alpha) = \dfrac{y_i+\alpha}{\mu_i+\alpha} $$

# Estimation of $\theta$ with a single $\alpha$ across all windows

## SSC (Cell 2022)

### chr1, windowsize =10000

```{r echo=F, message=FALSE, warning=FALSE}

dat <- read.table("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr1/rl_ssc_trost_combine_chr1_10000.txt",fill = T)
dat <- dat[complete.cases(dat$V5),]
  
colnames(dat) <- c("chr","start","end","obs","exp")
alpha <- est.alpha(dat)
print(sprintf("alpha = %s", alpha))
  
theta <- (dat$obs + alpha) / (dat$exp + alpha)
ggplot(data = data.frame(theta), aes(x = theta)) +
  geom_histogram(binwidth = 0.5, color = "black", fill = "steelblue") +
  labs(title = "Histogram of Theta",
       x = "Theta",
       y = "Frequency")

ggplot(data = data.frame(log(theta)), aes(x = log(theta))) +
  geom_histogram(binwidth = 0.5, color = "black", fill = "steelblue") +
  labs(title = "Histogram of log(theta)",
       x = "log(theta)",
       y = "Frequency")

```







### chr1, windowsize =50000

```{r echo=F, message=FALSE, warning=FALSE}

dat <- read.table("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr1/rl_ssc_trost_combine_chr1_50000.txt",fill = T)
dat <- dat[complete.cases(dat$V5),]
  
colnames(dat) <- c("chr","start","end","obs","exp")
alpha <- est.alpha(dat)
print(sprintf("alpha = %s", alpha))
  
theta <- (dat$obs + alpha) / (dat$exp + alpha)
ggplot(data = data.frame(theta), aes(x = theta)) +
  geom_histogram(binwidth = 0.5, color = "black", fill = "steelblue") +
  labs(title = "Histogram of Theta",
       x = "Theta",
       y = "Frequency")

ggplot(data = data.frame(log(theta)), aes(x = log(theta))) +
  geom_histogram(binwidth = 0.5, color = "black", fill = "steelblue") +
  labs(title = "Histogram of log(theta)",
       x = "log(theta)",
       y = "Frequency")

```






### chr22, windowsize =10000

```{r echo=F, message=FALSE, warning=FALSE}

dat <- read.table("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr22/rl_ssc_trost_combine_chr22_10000.txt",fill = T)
dat <- dat[complete.cases(dat$V5),]
  
colnames(dat) <- c("chr","start","end","obs","exp")
alpha <- est.alpha(dat)
print(sprintf("alpha = %s", alpha))
  
theta <- (dat$obs + alpha) / (dat$exp + alpha)
ggplot(data = data.frame(theta), aes(x = theta)) +
  geom_histogram(bins =50, color = "black", fill = "steelblue") +
  labs(title = "Histogram of Theta",
       x = "Theta",
       y = "Frequency")

ggplot(data = data.frame(log(theta)), aes(x = log(theta))) +
  geom_histogram(bins =50, color = "black", fill = "steelblue") +
  labs(title = "Histogram of log(theta)",
       x = "log(theta)",
       y = "Frequency")

```

### chr22, windowsize =50000

```{r echo=F, message=FALSE, warning=FALSE}

dat <- read.table("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr22/rl_ssc_trost_combine_chr22_50000.txt",fill = T)
dat <- dat[complete.cases(dat$V5),]
  
colnames(dat) <- c("chr","start","end","obs","exp")
alpha <- est.alpha(dat)
print(sprintf("alpha = %s", alpha))
  
theta <- (dat$obs + alpha) / (dat$exp + alpha)
ggplot(data = data.frame(theta), aes(x = theta)) +
  geom_histogram(bins =50, color = "black", fill = "steelblue") +
  labs(title = "Histogram of Theta",
       x = "Theta",
       y = "Frequency")

ggplot(data = data.frame(log(theta)), aes(x = log(theta))) +
  geom_histogram(bins =50, color = "black", fill = "steelblue") +
  labs(title = "Histogram of log(theta)",
       x = "log(theta)",
       y = "Frequency")

```










## MSSNG (Cell 2022)

### chr1, windowsize =10000

```{r echo=F, message=FALSE, warning=FALSE}

dat <- read.table("/project/xinhe/xsun/mutation_rate/5.de_novo_trost/data/per_chr_processed/chr1/rl_trost_combine_chr1_10000_mssng.txt",fill = T)
dat <- dat[complete.cases(dat$V5),]
  
colnames(dat) <- c("chr","start","end","obs","exp")
alpha <- est.alpha(dat)
print(sprintf("alpha = %s", alpha))
  
theta <- (dat$obs + alpha) / (dat$exp + alpha)
ggplot(data = data.frame(theta), aes(x = theta)) +
  geom_histogram(binwidth = 0.5, color = "black", fill = "steelblue") +
  labs(title = "Histogram of Theta",
       x = "Theta",
       y = "Frequency")

ggplot(data = data.frame(log(theta)), aes(x = log(theta))) +
  geom_histogram(binwidth = 0.5, color = "black", fill = "steelblue") +
  labs(title = "Histogram of log(theta)",
       x = "log(theta)",
       y = "Frequency")

```







### chr1, windowsize =50000

```{r echo=F, message=FALSE, warning=FALSE}

dat <- read.table("/project/xinhe/xsun/mutation_rate/5.de_novo_trost/data/per_chr_processed/chr1/rl_trost_combine_chr1_50000_mssng.txt",fill = T)
dat <- dat[complete.cases(dat$V5),]
  
colnames(dat) <- c("chr","start","end","obs","exp")
alpha <- est.alpha(dat)
print(sprintf("alpha = %s", alpha))
  
theta <- (dat$obs + alpha) / (dat$exp + alpha)
ggplot(data = data.frame(theta), aes(x = theta)) +
  geom_histogram(binwidth = 0.5, color = "black", fill = "steelblue") +
  labs(title = "Histogram of Theta",
       x = "Theta",
       y = "Frequency")

ggplot(data = data.frame(log(theta)), aes(x = log(theta))) +
  geom_histogram(binwidth = 0.5, color = "black", fill = "steelblue") +
  labs(title = "Histogram of log(theta)",
       x = "log(theta)",
       y = "Frequency")

```






### chr22, windowsize =10000

```{r echo=F, message=FALSE, warning=FALSE}

dat <- read.table("/project/xinhe/xsun/mutation_rate/5.de_novo_trost/data/per_chr_processed/chr22/rl_trost_combine_chr22_10000_mssng.txt",fill = T)
dat <- dat[complete.cases(dat$V5),]
  
colnames(dat) <- c("chr","start","end","obs","exp")
alpha <- est.alpha(dat)
print(sprintf("alpha = %s", alpha))
  
theta <- (dat$obs + alpha) / (dat$exp + alpha)
ggplot(data = data.frame(theta), aes(x = theta)) +
  geom_histogram(bins =50, color = "black", fill = "steelblue") +
  labs(title = "Histogram of Theta",
       x = "Theta",
       y = "Frequency")

ggplot(data = data.frame(log(theta)), aes(x = log(theta))) +
  geom_histogram(bins =50, color = "black", fill = "steelblue") +
  labs(title = "Histogram of log(theta)",
       x = "log(theta)",
       y = "Frequency")

```

### chr22, windowsize =50000

```{r echo=F, message=FALSE, warning=FALSE}

dat <- read.table("/project/xinhe/xsun/mutation_rate/5.de_novo_trost/data/per_chr_processed/chr22/rl_trost_combine_chr22_50000_mssng.txt",fill = T)
dat <- dat[complete.cases(dat$V5),]
  
colnames(dat) <- c("chr","start","end","obs","exp")
alpha <- est.alpha(dat)
print(sprintf("alpha = %s", alpha))
  
theta <- (dat$obs + alpha) / (dat$exp + alpha)
ggplot(data = data.frame(theta), aes(x = theta)) +
  geom_histogram(bins =50, color = "black", fill = "steelblue") +
  labs(title = "Histogram of Theta",
       x = "Theta",
       y = "Frequency")

ggplot(data = data.frame(log(theta)), aes(x = log(theta))) +
  geom_histogram(bins =50, color = "black", fill = "steelblue") +
  labs(title = "Histogram of log(theta)",
       x = "log(theta)",
       y = "Frequency")

```










# Cross validation 