---
title: "Simulation"
author: "XSun"
date: "2023-07-19"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r echo=F, message=FALSE, warning=FALSE}
library(ggplot2)
library(data.table)
library(gridExtra)
```

# Introduction

We cut the genome into bins with different sizes (10kb, 30kb, 50kb,100kb) and computed the overdispersion $\alpha$. Then we computed the non-spatial random effects $\theta_i$ from the $\alpha$, and simulated the observed $y_i$ with $y_i \sim Pois(\mu_i*\theta_i)$, where $\mu_i$ is the Roulette exprected mutation rates.

The simulation was performed 1000 times. 

For each simulation, we have $y_{i_{simulated}}$, we set $y_{i_{simulated}} = 0.05$ for all 0s, then computed  $\theta_{i_{simulated}}$ for each window. After that, we smoothed the $log(\theta_{i_{simulated}})$ with smash.gaus and computed the mean value of the smoothed data for all 1000 times simulation. 


# SSC (Cell 2022)

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}
folder <- "/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/results/"
```


## chr1,10kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load(paste0(folder,"simulation_chr1_10kb.rdata"))

dat$Mb <- dat$start/1000000

mean_values <- apply(smoothed_matrix, 1, mean)
standard_errors <- apply(smoothed_matrix, 2, function(x) sd(x) / sqrt(length(x)))

# Create a data frame for plotting
plot_data <- data.frame(
  x = 1:length(theta),
  mean_value = mean_values,
  lower_bound = mean_values - standard_errors,
  upper_bound = mean_values + standard_errors,
  theta = theta,
  mb = dat$Mb
)


ggplot(plot_data, aes(x = mb)) +
  geom_ribbon(aes(ymin = lower_bound, ymax = upper_bound), fill = "gray70", alpha = 0.5) +
  geom_line(aes(y = mean_value), color = "black") +
  geom_line(aes(y = theta), color = "red", alpha = 0.5) +
  labs(x = "Genomic Positions (Mb)", y = "theta") +
  theme_minimal() +
  annotate("text", x = Inf, y = Inf, hjust = 1, vjust = 1, label = "true theta (from non-spatial estimation)", color = "red") +
  annotate("text", x = Inf, y = Inf, hjust = 1, vjust = 2, label = "mean value of smoothed theta(from simulation)", color = "black") +
  theme(
    text = element_text(size = 12),  # Change this value to adjust the size of all text
    axis.title = element_text(size = 14),  # Change this value to adjust axis label size
    axis.text = element_text(size = 12),   # Change this value to adjust axis tick label size
    plot.title = element_text(size = 16)   # Change this value to adjust the plot title size
  )



```





<!-- ## chr1,30kb -->

<!-- ```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5} -->

<!-- load(paste0(folder,"simulation_chr1_30kb.rdata")) -->

<!-- dat$Mb <- dat$start/1000000 -->

<!-- mean_values <- apply(smoothed_matrix, 1, mean) -->
<!-- standard_errors <- apply(smoothed_matrix, 2, function(x) sd(x) / sqrt(length(x))) -->

<!-- # Create a data frame for plotting -->
<!-- plot_data <- data.frame( -->
<!--   x = 1:length(theta), -->
<!--   mean_value = mean_values, -->
<!--   lower_bound = mean_values - standard_errors, -->
<!--   upper_bound = mean_values + standard_errors, -->
<!--   theta = theta, -->
<!--   mb = dat$Mb -->
<!-- ) -->


<!-- ggplot(plot_data, aes(x = mb)) + -->
<!--   geom_ribbon(aes(ymin = lower_bound, ymax = upper_bound), fill = "gray70", alpha = 0.5) + -->
<!--   geom_line(aes(y = mean_value), color = "black") + -->
<!--   geom_line(aes(y = theta), color = "red", alpha = 0.5) + -->
<!--   labs(x = "Genomic Positions (Mb)", y = "theta") + -->
<!--   theme_minimal() + -->
<!--   annotate("text", x = Inf, y = Inf, hjust = 1, vjust = 1, label = "true theta (from non-spatial estimation)", color = "red") + -->
<!--   annotate("text", x = Inf, y = Inf, hjust = 1, vjust = 2, label = "mean value of smoothed theta(from simulation)", color = "black") + -->
<!--   theme( -->
<!--     text = element_text(size = 12),  # Change this value to adjust the size of all text -->
<!--     axis.title = element_text(size = 14),  # Change this value to adjust axis label size -->
<!--     axis.text = element_text(size = 12),   # Change this value to adjust axis tick label size -->
<!--     plot.title = element_text(size = 16)   # Change this value to adjust the plot title size -->
<!--   ) -->



<!-- ``` -->






## chr1,50kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load(paste0(folder,"simulation_chr1_50kb.rdata"))

dat$Mb <- dat$start/1000000

mean_values <- apply(smoothed_matrix, 1, mean)
standard_errors <- apply(smoothed_matrix, 2, function(x) sd(x) / sqrt(length(x)))

# Create a data frame for plotting
plot_data <- data.frame(
  x = 1:length(theta),
  mean_value = mean_values,
  lower_bound = mean_values - standard_errors,
  upper_bound = mean_values + standard_errors,
  theta = theta,
  mb = dat$Mb
)


ggplot(plot_data, aes(x = mb)) +
  geom_ribbon(aes(ymin = lower_bound, ymax = upper_bound), fill = "gray70", alpha = 0.5) +
  geom_line(aes(y = mean_value), color = "black") +
  geom_line(aes(y = theta), color = "red", alpha = 0.5) +
  labs(x = "Genomic Positions (Mb)", y = "theta") +
  theme_minimal() +
  annotate("text", x = Inf, y = Inf, hjust = 1, vjust = 1, label = "true theta (from non-spatial estimation)", color = "red") +
  annotate("text", x = Inf, y = Inf, hjust = 1, vjust = 2, label = "mean value of smoothed theta(from simulation)", color = "black") +
  theme(
    text = element_text(size = 12),  # Change this value to adjust the size of all text
    axis.title = element_text(size = 14),  # Change this value to adjust axis label size
    axis.text = element_text(size = 12),   # Change this value to adjust axis tick label size
    plot.title = element_text(size = 16)   # Change this value to adjust the plot title size
  )



```



## chr1,100kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load(paste0(folder,"simulation_chr1_100kb.rdata"))

dat$Mb <- dat$start/1000000

mean_values <- apply(smoothed_matrix, 1, mean)
standard_errors <- apply(smoothed_matrix, 2, function(x) sd(x) / sqrt(length(x)))

# Create a data frame for plotting
plot_data <- data.frame(
  x = 1:length(theta),
  mean_value = mean_values,
  lower_bound = mean_values - standard_errors,
  upper_bound = mean_values + standard_errors,
  theta = theta,
  mb = dat$Mb
)


ggplot(plot_data, aes(x = mb)) +
  geom_ribbon(aes(ymin = lower_bound, ymax = upper_bound), fill = "gray70", alpha = 0.5) +
  geom_line(aes(y = mean_value), color = "black") +
  geom_line(aes(y = theta), color = "red", alpha = 0.5) +
  labs(x = "Genomic Positions (Mb)", y = "theta") +
  theme_minimal() +
  annotate("text", x = Inf, y = Inf, hjust = 1, vjust = 1, label = "true theta (from non-spatial estimation)", color = "red") +
  annotate("text", x = Inf, y = Inf, hjust = 1, vjust = 2, label = "mean value of smoothed theta(from simulation)", color = "black") +
  theme(
    text = element_text(size = 12),  # Change this value to adjust the size of all text
    axis.title = element_text(size = 14),  # Change this value to adjust axis label size
    axis.text = element_text(size = 12),   # Change this value to adjust axis tick label size
    plot.title = element_text(size = 16)   # Change this value to adjust the plot title size
  )


```






# MSSNG (Cell 2022)

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}
folder <- "/project/xinhe/xsun/mutation_rate/5.de_novo_trost/results/"
```

## chr1,10kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load(paste0(folder,"simulation_chr1_10kb.rdata"))

dat$Mb <- dat$start/1000000

mean_values <- apply(smoothed_matrix, 1, mean)
standard_errors <- apply(smoothed_matrix, 2, function(x) sd(x) / sqrt(length(x)))

# Create a data frame for plotting
plot_data <- data.frame(
  x = 1:length(theta),
  mean_value = mean_values,
  lower_bound = mean_values - standard_errors,
  upper_bound = mean_values + standard_errors,
  theta = theta,
  mb = dat$Mb
)


ggplot(plot_data, aes(x = mb)) +
  geom_ribbon(aes(ymin = lower_bound, ymax = upper_bound), fill = "gray70", alpha = 0.5) +
  geom_line(aes(y = mean_value), color = "black") +
  geom_line(aes(y = theta), color = "red", alpha = 0.5) +
  labs(x = "Genomic Positions (Mb)", y = "theta") +
  theme_minimal() +
  annotate("text", x = Inf, y = Inf, hjust = 1, vjust = 1, label = "true theta (from non-spatial estimation)", color = "red") +
  annotate("text", x = Inf, y = Inf, hjust = 1, vjust = 2, label = "mean value of smoothed theta(from simulation)", color = "black") +
  theme(
    text = element_text(size = 12),  # Change this value to adjust the size of all text
    axis.title = element_text(size = 14),  # Change this value to adjust axis label size
    axis.text = element_text(size = 12),   # Change this value to adjust axis tick label size
    plot.title = element_text(size = 16)   # Change this value to adjust the plot title size
  )



```





## chr1,30kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load(paste0(folder,"simulation_chr1_30kb.rdata"))

dat$Mb <- dat$start/1000000

mean_values <- apply(smoothed_matrix, 1, mean)
standard_errors <- apply(smoothed_matrix, 2, function(x) sd(x) / sqrt(length(x)))

# Create a data frame for plotting
plot_data <- data.frame(
  x = 1:length(theta),
  mean_value = mean_values,
  lower_bound = mean_values - standard_errors,
  upper_bound = mean_values + standard_errors,
  theta = theta,
  mb = dat$Mb
)


ggplot(plot_data, aes(x = mb)) +
  geom_ribbon(aes(ymin = lower_bound, ymax = upper_bound), fill = "gray70", alpha = 0.5) +
  geom_line(aes(y = mean_value), color = "black") +
  geom_line(aes(y = theta), color = "red", alpha = 0.5) +
  labs(x = "Genomic Positions (Mb)", y = "theta") +
  theme_minimal() +
  annotate("text", x = Inf, y = Inf, hjust = 1, vjust = 1, label = "true theta (from non-spatial estimation)", color = "red") +
  annotate("text", x = Inf, y = Inf, hjust = 1, vjust = 2, label = "mean value of smoothed theta(from simulation)", color = "black") +
  theme(
    text = element_text(size = 12),  # Change this value to adjust the size of all text
    axis.title = element_text(size = 14),  # Change this value to adjust axis label size
    axis.text = element_text(size = 12),   # Change this value to adjust axis tick label size
    plot.title = element_text(size = 16)   # Change this value to adjust the plot title size
  )



```




## chr1,50kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load(paste0(folder,"simulation_chr1_50kb.rdata"))

dat$Mb <- dat$start/1000000

mean_values <- apply(smoothed_matrix, 1, mean)
standard_errors <- apply(smoothed_matrix, 2, function(x) sd(x) / sqrt(length(x)))

# Create a data frame for plotting
plot_data <- data.frame(
  x = 1:length(theta),
  mean_value = mean_values,
  lower_bound = mean_values - standard_errors,
  upper_bound = mean_values + standard_errors,
  theta = theta,
  mb = dat$Mb
)


ggplot(plot_data, aes(x = mb)) +
  geom_ribbon(aes(ymin = lower_bound, ymax = upper_bound), fill = "gray70", alpha = 0.5) +
  geom_line(aes(y = mean_value), color = "black") +
  geom_line(aes(y = theta), color = "red", alpha = 0.5) +
  labs(x = "Genomic Positions (Mb)", y = "theta") +
  theme_minimal() +
  annotate("text", x = Inf, y = Inf, hjust = 1, vjust = 1, label = "true theta (from non-spatial estimation)", color = "red") +
  annotate("text", x = Inf, y = Inf, hjust = 1, vjust = 2, label = "mean value of smoothed theta(from simulation)", color = "black") +
  theme(
    text = element_text(size = 12),  # Change this value to adjust the size of all text
    axis.title = element_text(size = 14),  # Change this value to adjust axis label size
    axis.text = element_text(size = 12),   # Change this value to adjust axis tick label size
    plot.title = element_text(size = 16)   # Change this value to adjust the plot title size
  )



```



## chr1,100kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load(paste0(folder,"simulation_chr1_100kb.rdata"))

dat$Mb <- dat$start/1000000

mean_values <- apply(smoothed_matrix, 1, mean)
standard_errors <- apply(smoothed_matrix, 2, function(x) sd(x) / sqrt(length(x)))

# Create a data frame for plotting
plot_data <- data.frame(
  x = 1:length(theta),
  mean_value = mean_values,
  lower_bound = mean_values - standard_errors,
  upper_bound = mean_values + standard_errors,
  theta = theta,
  mb = dat$Mb
)


ggplot(plot_data, aes(x = mb)) +
  geom_ribbon(aes(ymin = lower_bound, ymax = upper_bound), fill = "gray70", alpha = 0.5) +
  geom_line(aes(y = mean_value), color = "black") +
  geom_line(aes(y = theta), color = "red", alpha = 0.5) +
  labs(x = "Genomic Positions (Mb)", y = "theta") +
  theme_minimal() +
  annotate("text", x = Inf, y = Inf, hjust = 1, vjust = 1, label = "true theta (from non-spatial estimation)", color = "red") +
  annotate("text", x = Inf, y = Inf, hjust = 1, vjust = 2, label = "mean value of smoothed theta(from simulation)", color = "black") +
  theme(
    text = element_text(size = 12),  # Change this value to adjust the size of all text
    axis.title = element_text(size = 14),  # Change this value to adjust axis label size
    axis.text = element_text(size = 12),   # Change this value to adjust axis tick label size
    plot.title = element_text(size = 16)   # Change this value to adjust the plot title size
  )


```



