---
title: "Data cleaning -- SSC (An et al, science 2018)"
author: "XSun"
date: "2024-06-19"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
---

# Introduction

We checked the gaps from Roulette estimation here and get an overview for the observed fold change $r_i$ computed from [SSC Science 2018](https://xsun1229.github.io/mutation_rate/denovo_data_overview.html), the paper is here https://pubmed.ncbi.nlm.nih.gov/30545852/


```{r, message=F}
library(ggplot2)
library(kableExtra)
library(dplyr)
library(data.table)

create_filtered_plot <- function(data) {
  # Convert to data.table
  setDT(data)
  
  # Create group based on rleid
  data[, group := rleid(pass_threshold)]
  
  # Filter rows where pass_threshold is TRUE and the length of the group is greater than 1
  filtered_data <- data[pass_threshold == TRUE, .SD[.N > 1], by = group]
  
  # Remove the group column if not needed
  filtered_data[, group := NULL]
  
  # Create the plot
  plot <- ggplot(filtered_data, aes(x = Window_Start / 1000000, y = foldchange)) +
  geom_point(data = filtered_data[filtered_data$overlap_poliii == FALSE, ], aes(color = "FALSE")) +
  geom_point(data = filtered_data[filtered_data$overlap_poliii == TRUE, ], aes(color = "TRUE")) +
  labs(x = "Genomic position (mb)", y = "fold change r_i", color = "Overlap with polIII genes") +
  scale_color_manual(values = c("FALSE" = "black", "TRUE" = "red")) +
  theme_minimal()
  
  return(plot)
}

data_filter <- function(data) {
  
  setDT(data)
  
  # Create group based on rleid
  data[, group := rleid(pass_threshold)]
  
  # Filter rows where pass_threshold is TRUE and the length of the group is greater than 1
  filtered_data <- data[pass_threshold == TRUE, .SD[.N > 1], by = group]
  
  # Remove the group column if not needed
  filtered_data[, group := NULL]
  
  return(filtered_data)
  
}



```

# Notes from Roulette 

## Hotspot regions

Genes transcribed by RNA polymerase III (Pol III), especially tRNA and small nuclear RNA (RNU) genes, show about a 10-fold increase in mutation rate compared to Roulette's predictions. 

tRNA genes: http://gtrnadb.ucsc.edu/genomes/bacteria/Vibr_para_UCM_V493/Vibr_para_UCM_V493-gene-list.html

## Gaps in Roulette estimation

The gaps in the roulette estimation are due to:

- **Long gaps:** Centromeres

- **Small gaps:**
  1. Low density of polymorphism in GNOMAD at 100 nucleotide scale
  2. Poor mappability issues and repetitive sequences (UCSC self chain track)


# Removing the gaps and computing the fold changes 

The data cleaning is performed as follows:

1. Partition Chromosome into Windows:

The chromosome is divided into fixed windows of various sizes (10kb, 30kb, 50kb, 100kb).

2. Compute Gaps within Windows:

For each window, the positions with missing roulette estimation (gaps) are computed.

3. Filter Windows Based on Gaps:

Windows with less than 25% gaps are retained for further analysis.

4. Summarize Data within Windows:

For each window, the sum of roulette estimation and the observed de novo mutation count from MSSNG cell 2022 are calculated.

5. Split Data for Cross-Validation:

Each window's positions are divided into a training set (first 90%) and a testing set (remaining 10%) for cross-validation.

## An example for chr1, 100kb window size

```{r}

load("/project/xinhe/xsun/mutation_rate/12.denovo_real_ssc_an/data_cleaning/ssc_an_per_window_chr1_100kb_frac4.rdata")

DT::datatable(df_per_window_data,caption = htmltools::tags$caption(style = 'caption-side: left; text-align: left; color:black; font-size:150%;',''),options = list(pageLength = 10))

```

Other window sizes also have similar patterns. 

## fold change plots for chr1

We just select rows where the "TRUE" value in the pass_threshold column is continuous for more than one row.

### 10kb

```{r fig.width=10, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/12.denovo_real_ssc_an/data_cleaning/ssc_an_per_window_chr1_10kb_frac4.rdata")

# Example usage
plot <- create_filtered_plot(df_per_window_data)
print(plot)

filted_data <- data_filter(df_per_window_data)
filted_data$fold_change <- filted_data$obs_sum / filted_data$rl_rescaled_sum
filted_data_sorted <- filted_data[order(filted_data$fold_change,decreasing = T),]

DT::datatable(filted_data_sorted[1:10],caption = htmltools::tags$caption(style = 'caption-side: left; text-align: left; color:black; font-size:150%;','the windows with largest fold change(filtered)'),options = list(pageLength = 5))

```


### 30kb

```{r fig.width=10, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/12.denovo_real_ssc_an/data_cleaning/ssc_an_per_window_chr1_30kb_frac4.rdata")

# Example usage
plot <- create_filtered_plot(df_per_window_data)
print(plot)
```


### 50kb

```{r fig.width=10, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/12.denovo_real_ssc_an/data_cleaning/ssc_an_per_window_chr1_50kb_frac4.rdata")

# Example usage
plot <- create_filtered_plot(df_per_window_data)
print(plot)
```


### 100kb

```{r fig.width=10, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/12.denovo_real_ssc_an/data_cleaning/ssc_an_per_window_chr1_100kb_frac4.rdata")

# Example usage
plot <- create_filtered_plot(df_per_window_data)
print(plot)
```


