---
title: "Simulation -- combining spatial and non-spatial models"
author: "XSun"
date: "2023-09-06"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---
```{r echo=F,message=FALSE}
library(ggplot2)
library(gridExtra)
library(data.table)
```

# Introduction

In this part, we combined the spatial and non-spatial models.

1. Simulating the random effects

```{r echo=FALSE, eval=T ,message=FALSE, fig.width=12, fig.height=5}
generate_randeff_nonspatial_gammaprior_spatial <-  function(num_windows_all, num_windows_per_seg_min, num_windows_per_seg_max, randeff_max, num_seg_spatial,num_seg_nonspatial,alpha) {
  
  # Create the vector
  randeff <- numeric(num_windows_all)
  
  #spatial
  segment_lengths_spatial <- ceiling(runif(num_seg_spatial, min = num_windows_per_seg_min, max = num_windows_per_seg_max))
  
  # Calculate the total length covered by spatial random effects
  total_segment_length <- sum(segment_lengths_spatial) 
  
  # Calculate the number of zeros needed
  num_zeros <- num_windows_all - total_segment_length
  
  # Randomly select positions for spatial segment starts 
  segment_starts_spatial <- sample(1:num_zeros, num_seg_spatial, replace = FALSE)
  segment_ends_spatial <- segment_starts_spatial + segment_lengths_spatial - 1
  
  # Assign random values to segments
  for (i in 1:(num_seg_spatial)) {
    start_pos <- segment_starts_spatial[i]
    end_pos <- start_pos + segment_lengths_spatial[i] - 1
    randeff[start_pos:end_pos] <- runif(1, 0, randeff_max)
  }
  
  
  ## non-spatial
  concatenated_ranges_spatial <- numeric(0)
  for (i in 1:length(segment_starts_spatial)) {
    concatenated_ranges_spatial <- c(concatenated_ranges_spatial, segment_starts_spatial[i]:segment_ends_spatial[i])
  }
  
  # Randomly select positions for segment starts -- non-spatial
  segment_starts_nonspatial <- sample(setdiff(1:num_windows_all,concatenated_ranges_spatial), num_seg_nonspatial, replace = FALSE)
  randeff_nonspatial <- rgamma(n = num_seg_nonspatial, shape = alpha, rate = alpha)
  
  randeff[segment_starts_nonspatial] <- randeff_nonspatial
  
  # Adjust values to meet the mean of 1
  mean_adjustment <- 1 - mean(randeff)
  randeff <- randeff + mean_adjustment
  # Clip values between 0 and randeff_max
  randeff[randeff < 0] <- 0
  randeff[randeff > randeff_max] <- randeff_max
  
  
  return(randeff)
}



num_windows_all <- 2000
num_windows_per_seg_min <- 4
num_windows_per_seg_max <- 20  
randeff_max <- 5
num_seg_nonspatial <- 100
num_seg_spatial <- 20
alpha <- 5

set.seed(1)
randeff <-generate_randeff_nonspatial_gammaprior_spatial(num_windows_all, num_windows_per_seg_min, num_windows_per_seg_max, randeff_max, num_seg_spatial,num_seg_nonspatial,alpha)

plot(randeff, type="l", ylab = "simulated random effects")

```

2. Estimating the random effect $\hat\theta_i^{NS}$ with non-spatial model

3. Using $\hat\theta_i^{NS} * \mu_i$ as $\mu_{i_{new}}$ in spatial model, using smashgen to smooth the random effect $\hat\theta_i^{S}$

4. The final estimated random effect $\hat\theta_i = \hat\theta_i^{S} * \hat\theta_i^{NS}$

# Simulation

## 50kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_snons_chr1_50kb_maxiter50_maxrandeff10.rdata")

print("maximum random effects = 10")

ggplot(data, aes(x = seq_along(y_sample))) +
  geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  #geom_line(aes(y = randeff_estimated_nonspatial), color = "red",alpha=0.5) +
  #geom_line(aes(y = randeff_smooth_spatial), color = "green") +
  labs(x = "Window index",
       y = "#of DNM/window && true random effects",) +
  theme_minimal()

ggplot(data, aes(x = seq_along(y_sample))) +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_estimated_nonspatial), color = "red",alpha=0.5) +
  #geom_line(aes(y = randeff_smooth_spatial), color = "green") +
  labs(x = "Window index",
       y = "random effects",
       title = "estimated by non-spatial model") +
  theme_minimal()


fit1 <- lm(randeff_estimated_nonspatial ~ randeff_simulated + 0, data=data)
adj_rsq1 <- summary(fit1)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_estimated_nonspatial)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq1, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "random effect - NS",
       title = "non-spatial model") +
  theme_minimal()

print("then we used dat$exp*theta_est_NS as new dat$exp, and used smashgen to smooth the estimation")

ggplot(data, aes(x = seq_along(y_sample))) +
  geom_line(aes(y = randeff_smooth_spatial), color = "red",alpha=0.5) +
  labs(x = "Window index",
       y = "smoothed random effects",
       title = "estimated by spatial model") +
  theme_minimal()

print("the final estimated random effects = random effects_NS * random effects_S ")

ggplot(data, aes(x = seq_along(y_sample))) +
  #geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_final), color = "red",alpha=0.5) +
  #geom_line(aes(y = randeff_smooth_spatial), color = "green") +
  labs(x = "Window index",
       y = "true random effects and final estimated random effects",
       title = "random effects_NS * random effects_S ") +
  theme_minimal()

fit2 <- lm(randeff_final ~ randeff_simulated + 0, data=data)
adj_rsq2 <- summary(fit2)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_final)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq2, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "random effect - NS * S",
       title = "non-spatial & spatial model") +
  theme_minimal()

```


# More investigation


It's weird that the smoothed random effects are constant. So we did more analysis to figure out what happened. 

## Fitting spatial model only

If we smooth the random effects using origin roulette expected mutation rate, we have: 

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_snons_debug.rdata")

ggplot(data, aes(x = seq_along(y_sample))) +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_smooth_org), color = "red",alpha=0.5) +
  #geom_line(aes(y = randeff_smooth_spatial), color = "green") +
  labs(x = "Window index",
       y = "random effects",) +
  theme_minimal()

```

the smoothing looks ok, the only difference is that the earlier expected rate was computed from  $\hat\theta_i^{NS} * \mu_i$, (where $\mu_i$ is the roulette expected mutation rate) and the current expected rate is $\mu_i$,

so we compare the $\mu_i$ and $\hat\theta_i^{NS} * \mu_i$: 

```{r echo=T, message=FALSE, warning=FALSE, fig.width=5, fig.height=5}

ggplot(data, aes(x = exp_nonspatial, y = exp_rl)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  labs(x = "exp_nonspatial (theta_i_NS * mu_i)",
       y = "exp_rl (mu_i)") +
  theme_minimal()

```

<!-- Aggregate all estimated random effect in one plot: -->

<!-- ```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5} -->


<!-- ggplot(data, aes(x = seq_along(y_sample))) + -->
<!--   geom_line(aes(y = randeff_simulated), color = "blue") + -->
<!--   geom_line(aes(y = randeff_smooth_org), color = "red",alpha=0.5) + -->
<!--   geom_line(aes(y = randeff_smooth_spatial), color = "green") + -->
<!--   labs(x = "Window index", -->
<!--        y = "random effects",) + -->
<!--   theme_minimal() -->

<!-- ``` -->


## Varying the strenth of non-spatial model

For the non-spatial model, we very the gamma prior alpha from 5 to 50 to check if the results look ok. The model behavior depends on the alpha parameter in the Gamma prior. As alpha increases (stronger prior), the non-spatial model basically shrinks the NS random effects to 1, then the model should reduce to spatial models

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_snons_debug_truealpha_5.rdata")

print("true alpha = 5")

ggplot(data, aes(x = seq_along(y_sample))) +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_estimated_nonspatial), color = "red",alpha=0.5) +
  geom_line(aes(y = randeff_smooth_spatial), color = "green") +
  labs(x = "Window index",
       y = "random effects",) +
  theme_minimal()


load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_snons_debug_truealpha_10.rdata")

print("true alpha = 10")

ggplot(data, aes(x = seq_along(y_sample))) +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_estimated_nonspatial), color = "red",alpha=0.5) +
  geom_line(aes(y = randeff_smooth_spatial), color = "green") +
  labs(x = "Window index",
       y = "random effects",) +
  theme_minimal()


load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_snons_debug_truealpha_30.rdata")

print("true alpha = 30")

ggplot(data, aes(x = seq_along(y_sample))) +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_estimated_nonspatial), color = "red",alpha=0.5) +
  geom_line(aes(y = randeff_smooth_spatial), color = "green") +
  labs(x = "Window index",
       y = "random effects",) +
  theme_minimal()


load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_snons_debug_truealpha_50.rdata")

print("true alpha = 50")

ggplot(data, aes(x = seq_along(y_sample))) +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_estimated_nonspatial), color = "red",alpha=0.5) +
  geom_line(aes(y = randeff_smooth_spatial), color = "green") +
  labs(x = "Window index",
       y = "random effects",) +
  theme_minimal()

```

## Varying the strenth of spatial model

We used larger hot-spots, but making the max.effects weaker. because of the weaker max. effects, the non-spatial model is unlikely to find all random effects, then the spatial model may do better.

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_snons_debug_stronger_spatial.rdata")

ggplot(data, aes(x = seq_along(y_sample))) +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_estimated_nonspatial), color = "red",alpha=0.5) +
  geom_line(aes(y = randeff_smooth_spatial), color = "green") +
  labs(x = "Window index",
       y = "random effects") +
  theme_minimal()

```


## Fitting spatial model first

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/7.denovo_simulation/results/simu_snons_debug_spatial_first.rdata")

print("true alpha = 5")

ggplot(data, aes(x = seq_along(y_sample))) +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_spatial), color = "red",alpha=0.5) +
  geom_line(aes(y = randeff_nonspatial), color = "green") +
  labs(x = "Window index",
       y = "random effects",) +
  theme_minimal()


fit1 <- lm(randeff_spatial ~ randeff_simulated + 0, data=data)
adj_rsq1 <- summary(fit1)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_spatial)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq1, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "random effect - Spatial",
       title = "spatial model") +
  theme_minimal()

print("the final estimated random effects = random effects_NS * random effects_S ")

data$randeff_final <- data$randeff_spatial * data$randeff_nonspatial

ggplot(data, aes(x = seq_along(y_sample))) +
  #geom_point(aes(y = y_sample), color = "grey") +
  geom_line(aes(y = randeff_simulated), color = "blue") +
  geom_line(aes(y = randeff_final), color = "red",alpha=0.5) +
  #geom_line(aes(y = randeff_smooth_spatial), color = "green") +
  labs(x = "Window index",
       y = "true random effects and final estimated random effects",
       title = "random effects_NS * random effects_S ") +
  theme_minimal()



fit2 <- lm(randeff_final ~ randeff_simulated + 0, data=data)
adj_rsq2 <- summary(fit2)$adj.r.squared
ggplot(data, aes(x = randeff_simulated, y = randeff_final)) +
  geom_point(color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq2, 3))),
            x = Inf, y = -Inf, hjust = 1, vjust = 0,
            color = "blue") +
  labs(x = "true random effect theta_i",
       y = "random effect - NS * S",
       title = "non-spatial & spatial model") +
  theme_minimal()

```
