---
title: "Cross Validation for non-spatial random effect"
author: "XSun"
date: "2023-07-18"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r echo=F, message=FALSE, warning=FALSE}
library(ggplot2)
library(data.table)
library(gridExtra)
```

# Introduction

We divided the whole genome into 100kb windows, used the first 90% as training set and the remaining 10% as testing set. Then we estimated $\alpha$ and $E(\theta_i|\alpha,\mu_i,y_i)$ on the training set for each window. Then we predicted $\hat{y}_{i}=\theta_{i} \cdot \mu_{i}$ for the testing set.

```{r echo=F, message=FALSE, warning=FALSE}

# Input is a dataframe of observed rates and expected rates
est.alpha <- function(dat) {
  mu_pos <- dat$exp > 0
  y_pos <- dat$obs > 0
  dat.mu_pos <- dat[(!y_pos)&mu_pos,]
  dat.y_pos <- dat[y_pos,]
  loglik <- function(alpha) {
    loglik.mu_pos <- alpha*(log(alpha) - log(dat.mu_pos$exp+alpha))
    loglik.y_pos <- log(gamma(dat.y_pos$obs+alpha)) - log(gamma(alpha)) + alpha*log(alpha) - (dat.y_pos$obs+alpha)*log(dat.y_pos$exp+alpha)
    -sum(loglik.mu_pos) - sum(loglik.y_pos)
  }
  res <- optim(par=1, fn=loglik, method="BFGS") # May need to try different initiations and methods
  res$par # MLE of alpha
}

est.alpha_n <- function(dat) {
  mu_pos <- dat$exp > 0
  y_pos <- dat$obs > 0
  dat.mu_pos <- dat[(!y_pos)&mu_pos,]
  dat.y_pos <- dat[y_pos,]
  loglik <- function(alpha) {
    loglik.mu_pos <- alpha*(log(alpha) - log(dat.mu_pos$exp+alpha))
    loglik.y_pos <- log(gamma(dat.y_pos$obs+alpha)) - log(gamma(alpha)) + alpha*log(alpha) - (dat.y_pos$obs+alpha)*log(dat.y_pos$exp+alpha)
    -sum(loglik.mu_pos) - sum(loglik.y_pos)
  }
  res <- optim(par=1, lower = 1, upper = 2,fn=loglik, method="Brent") # May need to try different initiations and methods
  res$par # MLE of alpha
}

```

# CV summary

| data set | chr | window size | alpha   | adjusted rsq -- random  | adjusted rsq -- roulette |
|----------|-----|-------------|---------|-------------------------|--------------------------|
| SSC      | 1   | 50kb        | 1.0786  | 0.2169                  | 0.2567                   |
|          |     | 100kb       | 0.1214  | 0.1843                  | 0.3267                   |
|          |     | 200kb       | 2.0000  | 0.2754                  | 0.4225                   |
|          | 22  | 50kb        | 31.9072 | 0.2945                  | 0.2893                   |
|          |     | 100kb       | 16.8044 | 0.5097                  | 0.5110                   |
|          |     | 200kb       | 77.4343 | 0.6881                  | 0.6873                   |
| MSSNG    | 1   | 50kb        | 0.4752  | 0.1976                  | 0.1528                   |
|          |     | 100kb       | 0.5307  | 0.0909                  | 0.0875                   |
|          |     | 200kb       | 0.5987  | 0.2897                  | 0.2370                   |
|          | 22  | 50kb        | 0.4633  | 0.1738                  | 0.1544                   |
|          |     | 100kb       | 0.5064  | 0.2593                  | 0.2522                   |
|          |     | 200kb       | 0.5260  | 0.4494                  | 0.3936                   |

# SSC (Cell 2022)

## chr1, windowsize =50kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

dat <- read.table("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr1/cv_combine_chr1_50kb_trainprop9.txt",header = T)

dat_train <- dat[,c("mr_rescale_train","count_train")]
colnames(dat_train) <- c("exp","obs")

alpha <- est.alpha(dat_train)
print(sprintf("alpha on training set = %s", alpha))
  
theta <- (dat_train$obs + alpha) / (dat_train$exp + alpha)
p1 <- ggplot(data = data.frame(theta), aes(x = theta)) +
  geom_histogram(bins =50, color = "black", fill = "steelblue") +
  labs(title = "Histogram of Theta on training set",
       x = "Theta",
       y = "Frequency") 
  

p2 <- ggplot(data = data.frame(log(theta)), aes(x = log(theta))) +
  geom_histogram(bins =50, color = "black", fill = "steelblue") +
  labs(title = "Histogram of log(theta) on training set",
       x = "log(theta)",
       y = "Frequency")
grid.arrange(p1, p2, nrow = 1)


y_test_pred <- theta * dat$mr_rescale_test 
data_plot <-  as.data.frame(cbind(y_test_pred,dat$mr_rescale_test,dat$count_test))
colnames(data_plot) <- c("exp_randeff","exp_rl","obs")

fit1 <- lm(data_plot$exp_randeff ~ data_plot$obs + 0 )

p1 <- ggplot(data = data_plot, aes(x = obs, y = exp_randeff)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  xlab("observed de novo SNVs") +
  ylab("predicted with random effects") +
  geom_text(aes(label = paste0("adjusted rsq =", round(summary(fit1)$adj.r.squared, digits = 4))),
            x = Inf, y = Inf, hjust = 1, vjust = 1, size = 5, color = "red")


#print("fitting: exp_randeff ~ obs + 0")
#print(summary(fit1))


fit2 <- lm(data_plot$exp_rl ~ data_plot$obs + 0 )
p2 <- ggplot(data = data_plot, aes(x = obs, y = exp_rl)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  xlab("observed de novo SNVs") +
  ylab("Roulette baseline") +
  geom_text(aes(label = paste0("adjusted rsq =", round(summary(fit2)$adj.r.squared, digits = 4))),
            x = Inf, y = Inf, hjust = 1, vjust = 1, size = 5, color = "red")

grid.arrange(p1, p2, nrow = 1)

#print("fitting: exp_rl ~ obs + 0")
#print(summary(fit2))
```

## chr1, windowsize =100kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

dat <- read.table("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr1/cv_combine_chr1_100kb_trainprop9.txt",header = T)

dat_train <- dat[,c("mr_rescale_train","count_train")]
colnames(dat_train) <- c("exp","obs")

alpha <- est.alpha(dat_train)
print(sprintf("alpha on training set = %s", alpha))
  
theta <- (dat_train$obs + alpha) / (dat_train$exp + alpha)
p1 <- ggplot(data = data.frame(theta), aes(x = theta)) +
  geom_histogram(bins =50, color = "black", fill = "steelblue") +
  labs(title = "Histogram of Theta on training set",
       x = "Theta",
       y = "Frequency") 
  

p2 <- ggplot(data = data.frame(log(theta)), aes(x = log(theta))) +
  geom_histogram(bins =50, color = "black", fill = "steelblue") +
  labs(title = "Histogram of log(theta) on training set",
       x = "log(theta)",
       y = "Frequency")
grid.arrange(p1, p2, nrow = 1)


y_test_pred <- theta * dat$mr_rescale_test 
data_plot <-  as.data.frame(cbind(y_test_pred,dat$mr_rescale_test,dat$count_test))
colnames(data_plot) <- c("exp_randeff","exp_rl","obs")

fit1 <- lm(data_plot$exp_randeff ~ data_plot$obs + 0 )

p1 <- ggplot(data = data_plot, aes(x = obs, y = exp_randeff)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  xlab("observed de novo SNVs") +
  ylab("predicted with random effects") +
  geom_text(aes(label = paste0("adjusted rsq =", round(summary(fit1)$adj.r.squared, digits = 4))),
            x = Inf, y = Inf, hjust = 1, vjust = 1, size = 5, color = "red")


#print("fitting: exp_randeff ~ obs + 0")
#print(summary(fit1))


fit2 <- lm(data_plot$exp_rl ~ data_plot$obs + 0 )
p2 <- ggplot(data = data_plot, aes(x = obs, y = exp_rl)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  xlab("observed de novo SNVs") +
  ylab("Roulette baseline") +
  geom_text(aes(label = paste0("adjusted rsq =", round(summary(fit2)$adj.r.squared, digits = 4))),
            x = Inf, y = Inf, hjust = 1, vjust = 1, size = 5, color = "red")

grid.arrange(p1, p2, nrow = 1)

#print("fitting: exp_rl ~ obs + 0")
#print(summary(fit2))
```


## chr1, windowsize =200kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

dat <- read.table("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr1/cv_combine_chr1_200kb_trainprop9.txt",header = T)

dat_train <- dat[,c("mr_rescale_train","count_train")]
colnames(dat_train) <- c("exp","obs")

alpha <- est.alpha_n(dat_train)
print(sprintf("alpha on training set = %s", alpha))
  
theta <- (dat_train$obs + alpha) / (dat_train$exp + alpha)
p1 <- ggplot(data = data.frame(theta), aes(x = theta)) +
  geom_histogram(bins =50, color = "black", fill = "steelblue") +
  labs(title = "Histogram of Theta on training set",
       x = "Theta",
       y = "Frequency") 
  

p2 <- ggplot(data = data.frame(log(theta)), aes(x = log(theta))) +
  geom_histogram(bins =50, color = "black", fill = "steelblue") +
  labs(title = "Histogram of log(theta) on training set",
       x = "log(theta)",
       y = "Frequency")
grid.arrange(p1, p2, nrow = 1)


y_test_pred <- theta * dat$mr_rescale_test 
data_plot <-  as.data.frame(cbind(y_test_pred,dat$mr_rescale_test,dat$count_test))
colnames(data_plot) <- c("exp_randeff","exp_rl","obs")

fit1 <- lm(data_plot$exp_randeff ~ data_plot$obs + 0 )

p1 <- ggplot(data = data_plot, aes(x = obs, y = exp_randeff)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  xlab("observed de novo SNVs") +
  ylab("predicted with random effects") +
  geom_text(aes(label = paste0("adjusted rsq =", round(summary(fit1)$adj.r.squared, digits = 4))),
            x = Inf, y = Inf, hjust = 1, vjust = 1, size = 5, color = "red")


#print("fitting: exp_randeff ~ obs + 0")
#print(summary(fit1))


fit2 <- lm(data_plot$exp_rl ~ data_plot$obs + 0 )
p2 <- ggplot(data = data_plot, aes(x = obs, y = exp_rl)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  xlab("observed de novo SNVs") +
  ylab("Roulette baseline") +
  geom_text(aes(label = paste0("adjusted rsq =", round(summary(fit2)$adj.r.squared, digits = 4))),
            x = Inf, y = Inf, hjust = 1, vjust = 1, size = 5, color = "red")

grid.arrange(p1, p2, nrow = 1)

#print("fitting: exp_rl ~ obs + 0")
#print(summary(fit2))
```

## chr22, windowsize =50kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

dat <- read.table("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr22/cv_combine_chr22_50kb_trainprop9.txt",header = T)

dat_train <- dat[,c("mr_rescale_train","count_train")]
colnames(dat_train) <- c("exp","obs")

alpha <- est.alpha(dat_train)
print(sprintf("alpha on training set = %s", alpha))
  
theta <- (dat_train$obs + alpha) / (dat_train$exp + alpha)
p1 <- ggplot(data = data.frame(theta), aes(x = theta)) +
  geom_histogram(bins =50, color = "black", fill = "steelblue") +
  labs(title = "Histogram of Theta on training set",
       x = "Theta",
       y = "Frequency") 
  

p2 <- ggplot(data = data.frame(log(theta)), aes(x = log(theta))) +
  geom_histogram(bins =50, color = "black", fill = "steelblue") +
  labs(title = "Histogram of log(theta) on training set",
       x = "log(theta)",
       y = "Frequency")
grid.arrange(p1, p2, nrow = 1)


y_test_pred <- theta * dat$mr_rescale_test 
data_plot <-  as.data.frame(cbind(y_test_pred,dat$mr_rescale_test,dat$count_test))
colnames(data_plot) <- c("exp_randeff","exp_rl","obs")

fit1 <- lm(data_plot$exp_randeff ~ data_plot$obs + 0 )

p1 <- ggplot(data = data_plot, aes(x = obs, y = exp_randeff)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  xlab("observed de novo SNVs") +
  ylab("predicted with random effects") +
  geom_text(aes(label = paste0("adjusted rsq =", round(summary(fit1)$adj.r.squared, digits = 4))),
            x = Inf, y = Inf, hjust = 1, vjust = 1, size = 5, color = "red")


#print("fitting: exp_randeff ~ obs + 0")
#print(summary(fit1))


fit2 <- lm(data_plot$exp_rl ~ data_plot$obs + 0 )
p2 <- ggplot(data = data_plot, aes(x = obs, y = exp_rl)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  xlab("observed de novo SNVs") +
  ylab("Roulette baseline") +
  geom_text(aes(label = paste0("adjusted rsq =", round(summary(fit2)$adj.r.squared, digits = 4))),
            x = Inf, y = Inf, hjust = 1, vjust = 1, size = 5, color = "red")

grid.arrange(p1, p2, nrow = 1)

#print("fitting: exp_rl ~ obs + 0")
#print(summary(fit2))
```


## chr22, windowsize =100kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

dat <- read.table("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr22/cv_combine_chr22_100kb_trainprop9.txt",header = T)

dat_train <- dat[,c("mr_rescale_train","count_train")]
colnames(dat_train) <- c("exp","obs")

alpha <- est.alpha(dat_train)
print(sprintf("alpha on training set = %s", alpha))
  
theta <- (dat_train$obs + alpha) / (dat_train$exp + alpha)
p1 <- ggplot(data = data.frame(theta), aes(x = theta)) +
  geom_histogram(bins =50, color = "black", fill = "steelblue") +
  labs(title = "Histogram of Theta on training set",
       x = "Theta",
       y = "Frequency") 
  

p2 <- ggplot(data = data.frame(log(theta)), aes(x = log(theta))) +
  geom_histogram(bins =50, color = "black", fill = "steelblue") +
  labs(title = "Histogram of log(theta) on training set",
       x = "log(theta)",
       y = "Frequency")
grid.arrange(p1, p2, nrow = 1)


y_test_pred <- theta * dat$mr_rescale_test 
data_plot <-  as.data.frame(cbind(y_test_pred,dat$mr_rescale_test,dat$count_test))
colnames(data_plot) <- c("exp_randeff","exp_rl","obs")

fit1 <- lm(data_plot$exp_randeff ~ data_plot$obs + 0 )

p1 <- ggplot(data = data_plot, aes(x = obs, y = exp_randeff)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  xlab("observed de novo SNVs") +
  ylab("predicted with random effects") +
  geom_text(aes(label = paste0("adjusted rsq =", round(summary(fit1)$adj.r.squared, digits = 4))),
            x = Inf, y = Inf, hjust = 1, vjust = 1, size = 5, color = "red")


#print("fitting: exp_randeff ~ obs + 0")
#print(summary(fit1))


fit2 <- lm(data_plot$exp_rl ~ data_plot$obs + 0 )
p2 <- ggplot(data = data_plot, aes(x = obs, y = exp_rl)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  xlab("observed de novo SNVs") +
  ylab("Roulette baseline") +
  geom_text(aes(label = paste0("adjusted rsq =", round(summary(fit2)$adj.r.squared, digits = 4))),
            x = Inf, y = Inf, hjust = 1, vjust = 1, size = 5, color = "red")

grid.arrange(p1, p2, nrow = 1)

#print("fitting: exp_rl ~ obs + 0")
#print(summary(fit2))
```



## chr22, windowsize =200kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

dat <- read.table("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr22/cv_combine_chr22_200kb_trainprop9.txt",header = T)

dat_train <- dat[,c("mr_rescale_train","count_train")]
colnames(dat_train) <- c("exp","obs")

alpha <- est.alpha(dat_train)
print(sprintf("alpha on training set = %s", alpha))
  
theta <- (dat_train$obs + alpha) / (dat_train$exp + alpha)
p1 <- ggplot(data = data.frame(theta), aes(x = theta)) +
  geom_histogram(bins =50, color = "black", fill = "steelblue") +
  labs(title = "Histogram of Theta on training set",
       x = "Theta",
       y = "Frequency") 
  

p2 <- ggplot(data = data.frame(log(theta)), aes(x = log(theta))) +
  geom_histogram(bins =50, color = "black", fill = "steelblue") +
  labs(title = "Histogram of log(theta) on training set",
       x = "log(theta)",
       y = "Frequency")
grid.arrange(p1, p2, nrow = 1)


y_test_pred <- theta * dat$mr_rescale_test 
data_plot <-  as.data.frame(cbind(y_test_pred,dat$mr_rescale_test,dat$count_test))
colnames(data_plot) <- c("exp_randeff","exp_rl","obs")

fit1 <- lm(data_plot$exp_randeff ~ data_plot$obs + 0 )

p1 <- ggplot(data = data_plot, aes(x = obs, y = exp_randeff)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  xlab("observed de novo SNVs") +
  ylab("predicted with random effects") +
  geom_text(aes(label = paste0("adjusted rsq =", round(summary(fit1)$adj.r.squared, digits = 4))),
            x = Inf, y = Inf, hjust = 1, vjust = 1, size = 5, color = "red")


#print("fitting: exp_randeff ~ obs + 0")
#print(summary(fit1))


fit2 <- lm(data_plot$exp_rl ~ data_plot$obs + 0 )
p2 <- ggplot(data = data_plot, aes(x = obs, y = exp_rl)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  xlab("observed de novo SNVs") +
  ylab("Roulette baseline") +
  geom_text(aes(label = paste0("adjusted rsq =", round(summary(fit2)$adj.r.squared, digits = 4))),
            x = Inf, y = Inf, hjust = 1, vjust = 1, size = 5, color = "red")

grid.arrange(p1, p2, nrow = 1)

#print("fitting: exp_rl ~ obs + 0")
#print(summary(fit2))
```

# MSSNG (Cell 2022)

## chr1, windowsize =50kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

dat <- read.table("/project/xinhe/xsun/mutation_rate/5.de_novo_trost/data/per_chr_processed/chr1/cv_combine_chr1_50kb_trainprop9.txt",header = T)

dat_train <- dat[,c("mr_rescale_train","count_train")]
colnames(dat_train) <- c("exp","obs")

alpha <- est.alpha(dat_train)
print(sprintf("alpha on training set = %s", alpha))
  
theta <- (dat_train$obs + alpha) / (dat_train$exp + alpha)
p1 <- ggplot(data = data.frame(theta), aes(x = theta)) +
  geom_histogram(bins =50, color = "black", fill = "steelblue") +
  labs(title = "Histogram of Theta on training set",
       x = "Theta",
       y = "Frequency") 
  

p2 <- ggplot(data = data.frame(log(theta)), aes(x = log(theta))) +
  geom_histogram(bins =50, color = "black", fill = "steelblue") +
  labs(title = "Histogram of log(theta) on training set",
       x = "log(theta)",
       y = "Frequency")
grid.arrange(p1, p2, nrow = 1)


y_test_pred <- theta * dat$mr_rescale_test 
data_plot <-  as.data.frame(cbind(y_test_pred,dat$mr_rescale_test,dat$count_test))
colnames(data_plot) <- c("exp_randeff","exp_rl","obs")

fit1 <- lm(data_plot$exp_randeff ~ data_plot$obs + 0 )

p1 <- ggplot(data = data_plot, aes(x = obs, y = exp_randeff)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  xlab("observed de novo SNVs") +
  ylab("predicted with random effects") +
  geom_text(aes(label = paste0("adjusted rsq =", round(summary(fit1)$adj.r.squared, digits = 4))),
            x = Inf, y = Inf, hjust = 1, vjust = 1, size = 5, color = "red")


#print("fitting: exp_randeff ~ obs + 0")
#print(summary(fit1))


fit2 <- lm(data_plot$exp_rl ~ data_plot$obs + 0 )
p2 <- ggplot(data = data_plot, aes(x = obs, y = exp_rl)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  xlab("observed de novo SNVs") +
  ylab("Roulette baseline") +
  geom_text(aes(label = paste0("adjusted rsq =", round(summary(fit2)$adj.r.squared, digits = 4))),
            x = Inf, y = Inf, hjust = 1, vjust = 1, size = 5, color = "red")

grid.arrange(p1, p2, nrow = 1)

##print("fitting: exp_rl ~ obs + 0")
##print(summary(fit2))
```




## chr1, windowsize =100kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

dat <- read.table("/project/xinhe/xsun/mutation_rate/5.de_novo_trost/data/per_chr_processed/chr1/cv_combine_chr1_100kb_trainprop9.txt",header = T)

dat_train <- dat[,c("mr_rescale_train","count_train")]
colnames(dat_train) <- c("exp","obs")

alpha <- est.alpha(dat_train)
print(sprintf("alpha on training set = %s", alpha))
  
theta <- (dat_train$obs + alpha) / (dat_train$exp + alpha)
p1 <- ggplot(data = data.frame(theta), aes(x = theta)) +
  geom_histogram(bins =50, color = "black", fill = "steelblue") +
  labs(title = "Histogram of Theta on training set",
       x = "Theta",
       y = "Frequency") 
  

p2 <- ggplot(data = data.frame(log(theta)), aes(x = log(theta))) +
  geom_histogram(bins =50, color = "black", fill = "steelblue") +
  labs(title = "Histogram of log(theta) on training set",
       x = "log(theta)",
       y = "Frequency")
grid.arrange(p1, p2, nrow = 1)


y_test_pred <- theta * dat$mr_rescale_test 
data_plot <-  as.data.frame(cbind(y_test_pred,dat$mr_rescale_test,dat$count_test))
colnames(data_plot) <- c("exp_randeff","exp_rl","obs")

fit1 <- lm(data_plot$exp_randeff ~ data_plot$obs + 0 )

p1 <- ggplot(data = data_plot, aes(x = obs, y = exp_randeff)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  xlab("observed de novo SNVs") +
  ylab("predicted with random effects") +
  geom_text(aes(label = paste0("adjusted rsq =", round(summary(fit1)$adj.r.squared, digits = 4))),
            x = Inf, y = Inf, hjust = 1, vjust = 1, size = 5, color = "red")


#print("fitting: exp_randeff ~ obs + 0")
#print(summary(fit1))


fit2 <- lm(data_plot$exp_rl ~ data_plot$obs + 0 )
p2 <- ggplot(data = data_plot, aes(x = obs, y = exp_rl)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  xlab("observed de novo SNVs") +
  ylab("Roulette baseline") +
  geom_text(aes(label = paste0("adjusted rsq =", round(summary(fit2)$adj.r.squared, digits = 4))),
            x = Inf, y = Inf, hjust = 1, vjust = 1, size = 5, color = "red")

grid.arrange(p1, p2, nrow = 1)

##print("fitting: exp_rl ~ obs + 0")
##print(summary(fit2))
```




## chr1, windowsize =200kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

dat <- read.table("/project/xinhe/xsun/mutation_rate/5.de_novo_trost/data/per_chr_processed/chr1/cv_combine_chr1_200kb_trainprop9.txt",header = T)

dat_train <- dat[,c("mr_rescale_train","count_train")]
colnames(dat_train) <- c("exp","obs")

alpha <- est.alpha(dat_train)
print(sprintf("alpha on training set = %s", alpha))
  
theta <- (dat_train$obs + alpha) / (dat_train$exp + alpha)
p1 <- ggplot(data = data.frame(theta), aes(x = theta)) +
  geom_histogram(bins =50, color = "black", fill = "steelblue") +
  labs(title = "Histogram of Theta on training set",
       x = "Theta",
       y = "Frequency") 
  

p2 <- ggplot(data = data.frame(log(theta)), aes(x = log(theta))) +
  geom_histogram(bins =50, color = "black", fill = "steelblue") +
  labs(title = "Histogram of log(theta) on training set",
       x = "log(theta)",
       y = "Frequency")
grid.arrange(p1, p2, nrow = 1)


y_test_pred <- theta * dat$mr_rescale_test 
data_plot <-  as.data.frame(cbind(y_test_pred,dat$mr_rescale_test,dat$count_test))
colnames(data_plot) <- c("exp_randeff","exp_rl","obs")

fit1 <- lm(data_plot$exp_randeff ~ data_plot$obs + 0 )

p1 <- ggplot(data = data_plot, aes(x = obs, y = exp_randeff)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  xlab("observed de novo SNVs") +
  ylab("predicted with random effects") +
  geom_text(aes(label = paste0("adjusted rsq =", round(summary(fit1)$adj.r.squared, digits = 4))),
            x = Inf, y = Inf, hjust = 1, vjust = 1, size = 5, color = "red")


#print("fitting: exp_randeff ~ obs + 0")
#print(summary(fit1))


fit2 <- lm(data_plot$exp_rl ~ data_plot$obs + 0 )
p2 <- ggplot(data = data_plot, aes(x = obs, y = exp_rl)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  xlab("observed de novo SNVs") +
  ylab("Roulette baseline") +
  geom_text(aes(label = paste0("adjusted rsq =", round(summary(fit2)$adj.r.squared, digits = 4))),
            x = Inf, y = Inf, hjust = 1, vjust = 1, size = 5, color = "red")

grid.arrange(p1, p2, nrow = 1)

##print("fitting: exp_rl ~ obs + 0")
##print(summary(fit2))
```

## chr22, windowsize =50kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

dat <- read.table("/project/xinhe/xsun/mutation_rate/5.de_novo_trost/data/per_chr_processed/chr22/cv_combine_chr22_50kb_trainprop9.txt",header = T)

dat_train <- dat[,c("mr_rescale_train","count_train")]
colnames(dat_train) <- c("exp","obs")

alpha <- est.alpha(dat_train)
print(sprintf("alpha on training set = %s", alpha))
  
theta <- (dat_train$obs + alpha) / (dat_train$exp + alpha)
p1 <- ggplot(data = data.frame(theta), aes(x = theta)) +
  geom_histogram(bins =50, color = "black", fill = "steelblue") +
  labs(title = "Histogram of Theta on training set",
       x = "Theta",
       y = "Frequency") 
  

p2 <- ggplot(data = data.frame(log(theta)), aes(x = log(theta))) +
  geom_histogram(bins =50, color = "black", fill = "steelblue") +
  labs(title = "Histogram of log(theta) on training set",
       x = "log(theta)",
       y = "Frequency")
grid.arrange(p1, p2, nrow = 1)


y_test_pred <- theta * dat$mr_rescale_test 
data_plot <-  as.data.frame(cbind(y_test_pred,dat$mr_rescale_test,dat$count_test))
colnames(data_plot) <- c("exp_randeff","exp_rl","obs")

fit1 <- lm(data_plot$exp_randeff ~ data_plot$obs + 0 )

p1 <- ggplot(data = data_plot, aes(x = obs, y = exp_randeff)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  xlab("observed de novo SNVs") +
  ylab("predicted with random effects") +
  geom_text(aes(label = paste0("adjusted rsq =", round(summary(fit1)$adj.r.squared, digits = 4))),
            x = Inf, y = Inf, hjust = 1, vjust = 1, size = 5, color = "red")


#print("fitting: exp_randeff ~ obs + 0")
#print(summary(fit1))


fit2 <- lm(data_plot$exp_rl ~ data_plot$obs + 0 )
p2 <- ggplot(data = data_plot, aes(x = obs, y = exp_rl)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  xlab("observed de novo SNVs") +
  ylab("Roulette baseline") +
  geom_text(aes(label = paste0("adjusted rsq =", round(summary(fit2)$adj.r.squared, digits = 4))),
            x = Inf, y = Inf, hjust = 1, vjust = 1, size = 5, color = "red")

grid.arrange(p1, p2, nrow = 1)

#print("fitting: exp_rl ~ obs + 0")
#print(summary(fit2))
```



## chr22, windowsize =100kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

dat <- read.table("/project/xinhe/xsun/mutation_rate/5.de_novo_trost/data/per_chr_processed/chr22/cv_combine_chr22_100kb_trainprop9.txt",header = T)

dat_train <- dat[,c("mr_rescale_train","count_train")]
colnames(dat_train) <- c("exp","obs")

alpha <- est.alpha(dat_train)
print(sprintf("alpha on training set = %s", alpha))
  
theta <- (dat_train$obs + alpha) / (dat_train$exp + alpha)
p1 <- ggplot(data = data.frame(theta), aes(x = theta)) +
  geom_histogram(bins =50, color = "black", fill = "steelblue") +
  labs(title = "Histogram of Theta on training set",
       x = "Theta",
       y = "Frequency") 
  

p2 <- ggplot(data = data.frame(log(theta)), aes(x = log(theta))) +
  geom_histogram(bins =50, color = "black", fill = "steelblue") +
  labs(title = "Histogram of log(theta) on training set",
       x = "log(theta)",
       y = "Frequency")
grid.arrange(p1, p2, nrow = 1)


y_test_pred <- theta * dat$mr_rescale_test 
data_plot <-  as.data.frame(cbind(y_test_pred,dat$mr_rescale_test,dat$count_test))
colnames(data_plot) <- c("exp_randeff","exp_rl","obs")

fit1 <- lm(data_plot$exp_randeff ~ data_plot$obs + 0 )

p1 <- ggplot(data = data_plot, aes(x = obs, y = exp_randeff)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  xlab("observed de novo SNVs") +
  ylab("predicted with random effects") +
  geom_text(aes(label = paste0("adjusted rsq =", round(summary(fit1)$adj.r.squared, digits = 4))),
            x = Inf, y = Inf, hjust = 1, vjust = 1, size = 5, color = "red")


#print("fitting: exp_randeff ~ obs + 0")
#print(summary(fit1))


fit2 <- lm(data_plot$exp_rl ~ data_plot$obs + 0 )
p2 <- ggplot(data = data_plot, aes(x = obs, y = exp_rl)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  xlab("observed de novo SNVs") +
  ylab("Roulette baseline") +
  geom_text(aes(label = paste0("adjusted rsq =", round(summary(fit2)$adj.r.squared, digits = 4))),
            x = Inf, y = Inf, hjust = 1, vjust = 1, size = 5, color = "red")

grid.arrange(p1, p2, nrow = 1)

#print("fitting: exp_rl ~ obs + 0")
#print(summary(fit2))
```



## chr22, windowsize =200kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

dat <- read.table("/project/xinhe/xsun/mutation_rate/5.de_novo_trost/data/per_chr_processed/chr22/cv_combine_chr22_200kb_trainprop9.txt",header = T)

dat_train <- dat[,c("mr_rescale_train","count_train")]
colnames(dat_train) <- c("exp","obs")

alpha <- est.alpha(dat_train)
print(sprintf("alpha on training set = %s", alpha))
  
theta <- (dat_train$obs + alpha) / (dat_train$exp + alpha)
p1 <- ggplot(data = data.frame(theta), aes(x = theta)) +
  geom_histogram(bins =50, color = "black", fill = "steelblue") +
  labs(title = "Histogram of Theta on training set",
       x = "Theta",
       y = "Frequency") 
  

p2 <- ggplot(data = data.frame(log(theta)), aes(x = log(theta))) +
  geom_histogram(bins =50, color = "black", fill = "steelblue") +
  labs(title = "Histogram of log(theta) on training set",
       x = "log(theta)",
       y = "Frequency")
grid.arrange(p1, p2, nrow = 1)


y_test_pred <- theta * dat$mr_rescale_test 
data_plot <-  as.data.frame(cbind(y_test_pred,dat$mr_rescale_test,dat$count_test))
colnames(data_plot) <- c("exp_randeff","exp_rl","obs")

fit1 <- lm(data_plot$exp_randeff ~ data_plot$obs + 0 )

p1 <- ggplot(data = data_plot, aes(x = obs, y = exp_randeff)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  xlab("observed de novo SNVs") +
  ylab("predicted with random effects") +
  geom_text(aes(label = paste0("adjusted rsq =", round(summary(fit1)$adj.r.squared, digits = 4))),
            x = Inf, y = Inf, hjust = 1, vjust = 1, size = 5, color = "red")


#print("fitting: exp_randeff ~ obs + 0")
#print(summary(fit1))


fit2 <- lm(data_plot$exp_rl ~ data_plot$obs + 0 )
p2 <- ggplot(data = data_plot, aes(x = obs, y = exp_rl)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  xlab("observed de novo SNVs") +
  ylab("Roulette baseline") +
  geom_text(aes(label = paste0("adjusted rsq =", round(summary(fit2)$adj.r.squared, digits = 4))),
            x = Inf, y = Inf, hjust = 1, vjust = 1, size = 5, color = "red")

grid.arrange(p1, p2, nrow = 1)

#print("fitting: exp_rl ~ obs + 0")
#print(summary(fit2))
```

