---
title: "Simulation summary for all settings"
author: "XSun"
date: "2025-02-12"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
---

# Introduction

We compared the \( R^2 \) values obtained from fitting simulated random effects and estimated random effects.  

- `_total_smashgen`: Total random effect estimated by Smashgen.  
- `_spatial_smashgen`: Spatial random effect estimated by Smashgen.  
- `_nonspatial_smashgen`: Nonspatial random effect estimated by Smashgen.  
- `_total_gamma`: Total random effect estimated by the gamma model.  

```{r }
library(ggplot2)
library(gridExtra)


plot_rsq_comparison_randeff <- function(data, color_var, color_label) {
  ggplot(data, aes(x = as.numeric(rsq_randeff_total_gamma), 
                   y = as.numeric(rsq_randeff_total_smashgen), 
                   color = .data[[color_var]])) +
    geom_point() +
    geom_abline(slope = 1, intercept = 0, col = "black") +
    labs(x = "rsq -- total randeff -- Gamma", 
         y = "rsq -- total randeff -- smashgen", 
         color = color_label) +
    theme_minimal()
}

plot_rsq_comparison_mutct <- function(data, color_var, color_label) {
  ggplot(data, aes(x = as.numeric(rsq_mutct_total_gamma), 
                   y = as.numeric(rsq_mutct_total_smashgen), 
                   color = .data[[color_var]])) +
    geom_point() +
    geom_abline(slope = 1, intercept = 0, col = "black") +
    labs(x = "rsq -- mutation count -- Gamma", 
         y = "rsq -- mutation count -- smashgen", 
         color = color_label) +
    theme_minimal()
}

plot_rsq_comparison_mutct_withrl <- function(data, color_var, color_label) {
  ggplot(data, aes(x = as.numeric(rsq_mutct_roulette), 
                   y = as.numeric(rsq_mutct_total_smashgen), 
                   color = .data[[color_var]])) +
    geom_point() +
    geom_abline(slope = 1, intercept = 0, col = "black") +
    labs(x = "rsq -- mutation count -- Gamma", 
         y = "rsq -- mutation count -- smashgen", 
         color = color_label) +
    theme_minimal()
}

```

# SSC 

## Window size = 50kb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/summary/winsize_50kb.rdata")

DT::datatable(summary,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Summary for different simulation settings'),options = list(pageLength = 10) )

summary_all <-c()
summary_all <- rbind(summary_all, summary)

p1 <- plot_rsq_comparison_randeff(summary, "non_spatial_alpha_true", "Non-spatial alpha -- true")
p2 <- plot_rsq_comparison_randeff(summary, "spatial_sigma", "spatial_sigma")
p3 <- plot_rsq_comparison_randeff(summary, "spatial_region_min", "spatial_region_min")
p4 <- plot_rsq_comparison_randeff(summary, "total_windows", "total_windows")

grid.arrange(p1, p2, p3, p4, nrow = 1)

p1 <- plot_rsq_comparison_mutct(summary, "non_spatial_alpha_true", "Non-spatial alpha -- true")
p2 <- plot_rsq_comparison_mutct(summary, "spatial_sigma", "spatial_sigma")
p3 <- plot_rsq_comparison_mutct(summary, "spatial_region_min", "spatial_region_min")
p4 <- plot_rsq_comparison_mutct(summary, "total_windows", "total_windows")

grid.arrange(p1, p2, p3, p4, nrow = 1)

p1 <- plot_rsq_comparison_mutct_withrl(summary, "non_spatial_alpha_true", "Non-spatial alpha -- true")
p2 <- plot_rsq_comparison_mutct_withrl(summary, "spatial_sigma", "spatial_sigma")
p3 <- plot_rsq_comparison_mutct_withrl(summary, "spatial_region_min", "spatial_region_min")
p4 <- plot_rsq_comparison_mutct_withrl(summary, "total_windows", "total_windows")

grid.arrange(p1, p2, p3, p4, nrow = 1)
```


## Window size = 100kb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/summary/winsize_100kb.rdata")

summary_all <- rbind(summary_all, summary)

DT::datatable(summary,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Summary for different simulation settings'),options = list(pageLength = 10) )

p1 <- plot_rsq_comparison_randeff(summary, "non_spatial_alpha_true", "Non-spatial alpha -- true")
p2 <- plot_rsq_comparison_randeff(summary, "spatial_sigma", "spatial_sigma")
p3 <- plot_rsq_comparison_randeff(summary, "spatial_region_min", "spatial_region_min")
p4 <- plot_rsq_comparison_randeff(summary, "total_windows", "total_windows")

grid.arrange(p1, p2, p3, p4, nrow = 1)

p1 <- plot_rsq_comparison_mutct(summary, "non_spatial_alpha_true", "Non-spatial alpha -- true")
p2 <- plot_rsq_comparison_mutct(summary, "spatial_sigma", "spatial_sigma")
p3 <- plot_rsq_comparison_mutct(summary, "spatial_region_min", "spatial_region_min")
p4 <- plot_rsq_comparison_mutct(summary, "total_windows", "total_windows")

grid.arrange(p1, p2, p3, p4, nrow = 1)
p1 <- plot_rsq_comparison_mutct_withrl(summary, "non_spatial_alpha_true", "Non-spatial alpha -- true")
p2 <- plot_rsq_comparison_mutct_withrl(summary, "spatial_sigma", "spatial_sigma")
p3 <- plot_rsq_comparison_mutct_withrl(summary, "spatial_region_min", "spatial_region_min")
p4 <- plot_rsq_comparison_mutct_withrl(summary, "total_windows", "total_windows")

grid.arrange(p1, p2, p3, p4, nrow = 1)
```

## Window size = 500kb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/summary/winsize_500kb.rdata")

summary_all <- rbind(summary_all, summary)

DT::datatable(summary,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Summary for different simulation settings'),options = list(pageLength = 10) )

p1 <- plot_rsq_comparison_randeff(summary, "non_spatial_alpha_true", "Non-spatial alpha -- true")
p2 <- plot_rsq_comparison_randeff(summary, "spatial_sigma", "spatial_sigma")
p3 <- plot_rsq_comparison_randeff(summary, "spatial_region_min", "spatial_region_min")
p4 <- plot_rsq_comparison_randeff(summary, "total_windows", "total_windows")

grid.arrange(p1, p2, p3, p4, nrow = 1)

p1 <- plot_rsq_comparison_mutct(summary, "non_spatial_alpha_true", "Non-spatial alpha -- true")
p2 <- plot_rsq_comparison_mutct(summary, "spatial_sigma", "spatial_sigma")
p3 <- plot_rsq_comparison_mutct(summary, "spatial_region_min", "spatial_region_min")
p4 <- plot_rsq_comparison_mutct(summary, "total_windows", "total_windows")

grid.arrange(p1, p2, p3, p4, nrow = 1)
p1 <- plot_rsq_comparison_mutct_withrl(summary, "non_spatial_alpha_true", "Non-spatial alpha -- true")
p2 <- plot_rsq_comparison_mutct_withrl(summary, "spatial_sigma", "spatial_sigma")
p3 <- plot_rsq_comparison_mutct_withrl(summary, "spatial_region_min", "spatial_region_min")
p4 <- plot_rsq_comparison_mutct_withrl(summary, "total_windows", "total_windows")

grid.arrange(p1, p2, p3, p4, nrow = 1)
```

## Pooling all window sizes

```{r fig.width=6, fig.height=5}

plot_rsq_comparison_randeff(summary_all, "windowsize", "windowsize")

plot_rsq_comparison_mutct(summary_all, "windowsize", "windowsize")

plot_rsq_comparison_mutct_withrl(summary_all, "windowsize", "windowsize")
```



# MSSNG

## Window size = 50kb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/summary/winsize_50kb_mssng.rdata")

DT::datatable(summary,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Summary for different simulation settings'),options = list(pageLength = 10) )

summary_all <-c()
summary_all <- rbind(summary_all, summary)

p1 <- plot_rsq_comparison_randeff(summary, "non_spatial_alpha_true", "Non-spatial alpha -- true")
p2 <- plot_rsq_comparison_randeff(summary, "spatial_sigma", "spatial_sigma")
p3 <- plot_rsq_comparison_randeff(summary, "spatial_region_min", "spatial_region_min")
p4 <- plot_rsq_comparison_randeff(summary, "total_windows", "total_windows")

grid.arrange(p1, p2, p3, p4, nrow = 1)

p1 <- plot_rsq_comparison_mutct(summary, "non_spatial_alpha_true", "Non-spatial alpha -- true")
p2 <- plot_rsq_comparison_mutct(summary, "spatial_sigma", "spatial_sigma")
p3 <- plot_rsq_comparison_mutct(summary, "spatial_region_min", "spatial_region_min")
p4 <- plot_rsq_comparison_mutct(summary, "total_windows", "total_windows")

grid.arrange(p1, p2, p3, p4, nrow = 1)

p1 <- plot_rsq_comparison_mutct_withrl(summary, "non_spatial_alpha_true", "Non-spatial alpha -- true")
p2 <- plot_rsq_comparison_mutct_withrl(summary, "spatial_sigma", "spatial_sigma")
p3 <- plot_rsq_comparison_mutct_withrl(summary, "spatial_region_min", "spatial_region_min")
p4 <- plot_rsq_comparison_mutct_withrl(summary, "total_windows", "total_windows")

grid.arrange(p1, p2, p3, p4, nrow = 1)
```


## Window size = 100kb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/summary/winsize_100kb_mssng.rdata")

summary_all <- rbind(summary_all, summary)

DT::datatable(summary,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Summary for different simulation settings'),options = list(pageLength = 10) )

p1 <- plot_rsq_comparison_randeff(summary, "non_spatial_alpha_true", "Non-spatial alpha -- true")
p2 <- plot_rsq_comparison_randeff(summary, "spatial_sigma", "spatial_sigma")
p3 <- plot_rsq_comparison_randeff(summary, "spatial_region_min", "spatial_region_min")
p4 <- plot_rsq_comparison_randeff(summary, "total_windows", "total_windows")

grid.arrange(p1, p2, p3, p4, nrow = 1)

p1 <- plot_rsq_comparison_mutct(summary, "non_spatial_alpha_true", "Non-spatial alpha -- true")
p2 <- plot_rsq_comparison_mutct(summary, "spatial_sigma", "spatial_sigma")
p3 <- plot_rsq_comparison_mutct(summary, "spatial_region_min", "spatial_region_min")
p4 <- plot_rsq_comparison_mutct(summary, "total_windows", "total_windows")

grid.arrange(p1, p2, p3, p4, nrow = 1)
p1 <- plot_rsq_comparison_mutct_withrl(summary, "non_spatial_alpha_true", "Non-spatial alpha -- true")
p2 <- plot_rsq_comparison_mutct_withrl(summary, "spatial_sigma", "spatial_sigma")
p3 <- plot_rsq_comparison_mutct_withrl(summary, "spatial_region_min", "spatial_region_min")
p4 <- plot_rsq_comparison_mutct_withrl(summary, "total_windows", "total_windows")

grid.arrange(p1, p2, p3, p4, nrow = 1)
```

## Window size = 500kb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/summary/winsize_500kb_mssng.rdata")

summary_all <- rbind(summary_all, summary)

DT::datatable(summary,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Summary for different simulation settings'),options = list(pageLength = 10) )

p1 <- plot_rsq_comparison_randeff(summary, "non_spatial_alpha_true", "Non-spatial alpha -- true")
p2 <- plot_rsq_comparison_randeff(summary, "spatial_sigma", "spatial_sigma")
p3 <- plot_rsq_comparison_randeff(summary, "spatial_region_min", "spatial_region_min")
p4 <- plot_rsq_comparison_randeff(summary, "total_windows", "total_windows")

grid.arrange(p1, p2, p3, p4, nrow = 1)

p1 <- plot_rsq_comparison_mutct(summary, "non_spatial_alpha_true", "Non-spatial alpha -- true")
p2 <- plot_rsq_comparison_mutct(summary, "spatial_sigma", "spatial_sigma")
p3 <- plot_rsq_comparison_mutct(summary, "spatial_region_min", "spatial_region_min")
p4 <- plot_rsq_comparison_mutct(summary, "total_windows", "total_windows")

grid.arrange(p1, p2, p3, p4, nrow = 1)
p1 <- plot_rsq_comparison_mutct_withrl(summary, "non_spatial_alpha_true", "Non-spatial alpha -- true")
p2 <- plot_rsq_comparison_mutct_withrl(summary, "spatial_sigma", "spatial_sigma")
p3 <- plot_rsq_comparison_mutct_withrl(summary, "spatial_region_min", "spatial_region_min")
p4 <- plot_rsq_comparison_mutct_withrl(summary, "total_windows", "total_windows")

grid.arrange(p1, p2, p3, p4, nrow = 1)
```

## Pooling all window sizes

```{r fig.width=6, fig.height=5}

plot_rsq_comparison_randeff(summary_all, "windowsize", "windowsize")

plot_rsq_comparison_mutct(summary_all, "windowsize", "windowsize")

plot_rsq_comparison_mutct_withrl(summary_all, "windowsize", "windowsize")
```

# SSC + deCODE

## Window size = 50kb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/summary/winsize_50kb_chen.rdata")

DT::datatable(summary,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Summary for different simulation settings'),options = list(pageLength = 10) )

summary_all <-c()
summary_all <- rbind(summary_all, summary)

p1 <- plot_rsq_comparison_randeff(summary, "non_spatial_alpha_true", "Non-spatial alpha -- true")
p2 <- plot_rsq_comparison_randeff(summary, "spatial_sigma", "spatial_sigma")
p3 <- plot_rsq_comparison_randeff(summary, "spatial_region_min", "spatial_region_min")
p4 <- plot_rsq_comparison_randeff(summary, "total_windows", "total_windows")

grid.arrange(p1, p2, p3, p4, nrow = 1)

p1 <- plot_rsq_comparison_mutct(summary, "non_spatial_alpha_true", "Non-spatial alpha -- true")
p2 <- plot_rsq_comparison_mutct(summary, "spatial_sigma", "spatial_sigma")
p3 <- plot_rsq_comparison_mutct(summary, "spatial_region_min", "spatial_region_min")
p4 <- plot_rsq_comparison_mutct(summary, "total_windows", "total_windows")

grid.arrange(p1, p2, p3, p4, nrow = 1)

p1 <- plot_rsq_comparison_mutct_withrl(summary, "non_spatial_alpha_true", "Non-spatial alpha -- true")
p2 <- plot_rsq_comparison_mutct_withrl(summary, "spatial_sigma", "spatial_sigma")
p3 <- plot_rsq_comparison_mutct_withrl(summary, "spatial_region_min", "spatial_region_min")
p4 <- plot_rsq_comparison_mutct_withrl(summary, "total_windows", "total_windows")

grid.arrange(p1, p2, p3, p4, nrow = 1)
```


## Window size = 100kb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/summary/winsize_100kb_chen.rdata")

summary_all <- rbind(summary_all, summary)

DT::datatable(summary,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Summary for different simulation settings'),options = list(pageLength = 10) )

p1 <- plot_rsq_comparison_randeff(summary, "non_spatial_alpha_true", "Non-spatial alpha -- true")
p2 <- plot_rsq_comparison_randeff(summary, "spatial_sigma", "spatial_sigma")
p3 <- plot_rsq_comparison_randeff(summary, "spatial_region_min", "spatial_region_min")
p4 <- plot_rsq_comparison_randeff(summary, "total_windows", "total_windows")

grid.arrange(p1, p2, p3, p4, nrow = 1)

p1 <- plot_rsq_comparison_mutct(summary, "non_spatial_alpha_true", "Non-spatial alpha -- true")
p2 <- plot_rsq_comparison_mutct(summary, "spatial_sigma", "spatial_sigma")
p3 <- plot_rsq_comparison_mutct(summary, "spatial_region_min", "spatial_region_min")
p4 <- plot_rsq_comparison_mutct(summary, "total_windows", "total_windows")

grid.arrange(p1, p2, p3, p4, nrow = 1)
p1 <- plot_rsq_comparison_mutct_withrl(summary, "non_spatial_alpha_true", "Non-spatial alpha -- true")
p2 <- plot_rsq_comparison_mutct_withrl(summary, "spatial_sigma", "spatial_sigma")
p3 <- plot_rsq_comparison_mutct_withrl(summary, "spatial_region_min", "spatial_region_min")
p4 <- plot_rsq_comparison_mutct_withrl(summary, "total_windows", "total_windows")

grid.arrange(p1, p2, p3, p4, nrow = 1)
```

## Window size = 500kb

```{r fig.width=20, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/summary/winsize_500kb_chen.rdata")

summary_all <- rbind(summary_all, summary)

DT::datatable(summary,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Summary for different simulation settings'),options = list(pageLength = 10) )

p1 <- plot_rsq_comparison_randeff(summary, "non_spatial_alpha_true", "Non-spatial alpha -- true")
p2 <- plot_rsq_comparison_randeff(summary, "spatial_sigma", "spatial_sigma")
p3 <- plot_rsq_comparison_randeff(summary, "spatial_region_min", "spatial_region_min")
p4 <- plot_rsq_comparison_randeff(summary, "total_windows", "total_windows")

grid.arrange(p1, p2, p3, p4, nrow = 1)

p1 <- plot_rsq_comparison_mutct(summary, "non_spatial_alpha_true", "Non-spatial alpha -- true")
p2 <- plot_rsq_comparison_mutct(summary, "spatial_sigma", "spatial_sigma")
p3 <- plot_rsq_comparison_mutct(summary, "spatial_region_min", "spatial_region_min")
p4 <- plot_rsq_comparison_mutct(summary, "total_windows", "total_windows")

grid.arrange(p1, p2, p3, p4, nrow = 1)
p1 <- plot_rsq_comparison_mutct_withrl(summary, "non_spatial_alpha_true", "Non-spatial alpha -- true")
p2 <- plot_rsq_comparison_mutct_withrl(summary, "spatial_sigma", "spatial_sigma")
p3 <- plot_rsq_comparison_mutct_withrl(summary, "spatial_region_min", "spatial_region_min")
p4 <- plot_rsq_comparison_mutct_withrl(summary, "total_windows", "total_windows")

grid.arrange(p1, p2, p3, p4, nrow = 1)
```

## Pooling all window sizes

```{r fig.width=6, fig.height=5}

plot_rsq_comparison_randeff(summary_all, "windowsize", "windowsize")

plot_rsq_comparison_mutct(summary_all, "windowsize", "windowsize")

plot_rsq_comparison_mutct_withrl(summary_all, "windowsize", "windowsize")
```

