---
title: "Cross Validation for spatial random effect -- without smoothing"
author: "XSun"
date: "2023-07-18"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(data.table)
library(ggplot2)
library(gridExtra)
```

# Introduction

We divided the whole genome into 100kb windows, used the first 90% as training set and the remaining 10% as testing set. Then we estimated $\theta_i$ for each window on the training sets and predicted $\hat{y}_{i}=\theta_{i} \cdot \mu_{i}$ for the testing set.


# CV summary




| data set | chr | window size | adjusted rsq -- random  | adjusted rsq -- roulette |
|----------|-----|-------------|-------------------------|--------------------------|
| SSC      | 1   | 50kb        | 0.2641                  | 0.2567                   |
|          |     | 100kb       | 0.2474                  | 0.3267                   |
|          |     | 200kb       | 0.0908                  | 0.4225                   |
|          | 22  | 50kb        | 0.2721                  | 0.2893                   |
|          |     | 100kb       | 0.4742                  | 0.5110                   |
|          |     | 200kb       | 0.6481                  | 0.6873                   |
| MSSNG    | 1   | 50kb        | 0.1859                  | 0.1528                   |
|          |     | 100kb       | 0.0874                  | 0.0875                   |
|          |     | 200kb       | 0.2859                  | 0.2370                   |
|          | 22  | 50kb        | 0.1677                  | 0.1544                   |
|          |     | 100kb       | 0.2526                  | 0.2522                   |
|          |     | 200kb       | 0.4455                  | 0.3936                   |

# SSC (Cell 2022)

## chr1, windowsize =50kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

dat <- read.table("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr1/cv_combine_chr1_50kb_trainprop9.txt",header = T)

dat$Mb <- dat$start/1000000
dat$randeff_train <- dat$count_train/dat$mr_rescale_train

print("number of de nove mutation in testing data")
print(table(dat$count_test))

ggplot(dat, aes(x = Mb, y = randeff_train )) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("Random effects") +
  ggtitle("Distribution of random effects on training set") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

dat$obs_pred_randeff <- dat$mr_rescale_test*dat$randeff_train


x_max <- max(dat$obs_pred_randeff, dat$mr_rescale_test,na.rm = T)
y_max <- max(dat$count_test)

fit1 <- lm(count_test ~ obs_pred_randeff + 0, data = dat)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

p1 <- ggplot(dat, aes(x = obs_pred_randeff, y = count_test)) +
  geom_point() +
  labs(title = "Adding random effects", x ="Predicted_random effects" , y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = Inf , y = Inf, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 4)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(count_test ~ mr_rescale_test + 0, data = dat)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

p2 <- ggplot(dat, aes(x = mr_rescale_test, y = count_test)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Predicted_Roulette", y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = Inf, y = Inf, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 4)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(p1, p2, ncol = 2)






```








## chr1, windowsize =100kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

dat <- read.table("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr1/cv_combine_chr1_100kb_trainprop9.txt",header = T)

dat$Mb <- dat$start/1000000
dat$randeff_train <- dat$count_train/dat$mr_rescale_train

print("number of de nove mutation in testing data")
print(table(dat$count_test))

ggplot(dat, aes(x = Mb, y = randeff_train )) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("Random effects") +
  ggtitle("Distribution of random effects on training set") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

dat$obs_pred_randeff <- dat$mr_rescale_test*dat$randeff_train


x_max <- max(dat$obs_pred_randeff, dat$mr_rescale_test,na.rm = T)
y_max <- max(dat$count_test)

fit1 <- lm(count_test ~ obs_pred_randeff + 0, data = dat)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

p1 <- ggplot(dat, aes(x = obs_pred_randeff, y = count_test)) +
  geom_point() +
  labs(title = "Adding random effects", x ="Predicted_random effects" , y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = Inf , y = Inf, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 4)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(count_test ~ mr_rescale_test + 0, data = dat)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

p2 <- ggplot(dat, aes(x = mr_rescale_test, y = count_test)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Predicted_Roulette", y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = Inf, y = Inf, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 4)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(p1, p2, ncol = 2)






```







## chr1, windowsize =200kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

dat <- read.table("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr1/cv_combine_chr1_200kb_trainprop9.txt",header = T)

dat$Mb <- dat$start/1000000
dat$randeff_train <- dat$count_train/dat$mr_rescale_train

print("number of de nove mutation in testing data")
print(table(dat$count_test))

ggplot(dat, aes(x = Mb, y = randeff_train )) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("Random effects") +
  ggtitle("Distribution of random effects on training set") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

dat$obs_pred_randeff <- dat$mr_rescale_test*dat$randeff_train


x_max <- max(dat$obs_pred_randeff, dat$mr_rescale_test,na.rm = T)
y_max <- max(dat$count_test)

fit1 <- lm(count_test ~ obs_pred_randeff + 0, data = dat)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

p1 <- ggplot(dat, aes(x = obs_pred_randeff, y = count_test)) +
  geom_point() +
  labs(title = "Adding random effects", x ="Predicted_random effects" , y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = Inf , y = Inf, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 4)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(count_test ~ mr_rescale_test + 0, data = dat)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

p2 <- ggplot(dat, aes(x = mr_rescale_test, y = count_test)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Predicted_Roulette", y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = Inf, y = Inf, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 4)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(p1, p2, ncol = 2)






```







## chr22, windowsize =50kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

dat <- read.table("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr22/cv_combine_chr22_50kb_trainprop9.txt",header = T)

dat$Mb <- dat$start/1000000
dat$randeff_train <- dat$count_train/dat$mr_rescale_train

print("number of de nove mutation in testing data") 
print(table(dat$count_test)) 

ggplot(dat, aes(x = Mb, y = randeff_train )) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("Random effects") +
  ggtitle("Distribution of random effects on training set") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

dat$obs_pred_randeff <- dat$mr_rescale_test*dat$randeff_train


x_max <- max(dat$obs_pred_randeff, dat$mr_rescale_test,na.rm = T)
y_max <- max(dat$count_test)

fit1 <- lm(count_test ~ obs_pred_randeff + 0, data = dat)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

p1 <- ggplot(dat, aes(x = obs_pred_randeff, y = count_test)) +
  geom_point() +
  labs(title = "Adding random effects", x ="Predicted_random effects" , y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = Inf , y = Inf, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 4)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(count_test ~ mr_rescale_test + 0, data = dat)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

p2 <- ggplot(dat, aes(x = mr_rescale_test, y = count_test)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Predicted_Roulette", y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = Inf, y = Inf, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 4)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(p1, p2, ncol = 2)

```








## chr22, windowsize =100kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

dat <- read.table("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr22/cv_combine_chr22_100kb_trainprop9.txt",header = T)

dat$Mb <- dat$start/1000000
dat$randeff_train <- dat$count_train/dat$mr_rescale_train

print("number of de nove mutation in testing data") 
print(table(dat$count_test)) 

ggplot(dat, aes(x = Mb, y = randeff_train )) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("Random effects") +
  ggtitle("Distribution of random effects on training set") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

dat$obs_pred_randeff <- dat$mr_rescale_test*dat$randeff_train


x_max <- max(dat$obs_pred_randeff, dat$mr_rescale_test,na.rm = T)
y_max <- max(dat$count_test)

fit1 <- lm(count_test ~ obs_pred_randeff + 0, data = dat)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

p1 <- ggplot(dat, aes(x = obs_pred_randeff, y = count_test)) +
  geom_point() +
  labs(title = "Adding random effects", x ="Predicted_random effects" , y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = Inf , y = Inf, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 4)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(count_test ~ mr_rescale_test + 0, data = dat)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

p2 <- ggplot(dat, aes(x = mr_rescale_test, y = count_test)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Predicted_Roulette", y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = Inf, y = Inf, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 4)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(p1, p2, ncol = 2)






```







## chr22, windowsize =200kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

dat <- read.table("/project/xinhe/xsun/mutation_rate/6.de_novo_ssc_trost/data/per_chr_processed/chr22/cv_combine_chr22_200kb_trainprop9.txt",header = T)

dat$Mb <- dat$start/1000000
dat$randeff_train <- dat$count_train/dat$mr_rescale_train

print("number of de nove mutation in testing data") 
print(table(dat$count_test)) 

ggplot(dat, aes(x = Mb, y = randeff_train )) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("Random effects") +
  ggtitle("Distribution of random effects on training set") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

dat$obs_pred_randeff <- dat$mr_rescale_test*dat$randeff_train


x_max <- max(dat$obs_pred_randeff, dat$mr_rescale_test,na.rm = T)
y_max <- max(dat$count_test)

fit1 <- lm(count_test ~ obs_pred_randeff + 0, data = dat)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

p1 <- ggplot(dat, aes(x = obs_pred_randeff, y = count_test)) +
  geom_point() +
  labs(title = "Adding random effects", x ="Predicted_random effects" , y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = Inf , y = Inf, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 4)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(count_test ~ mr_rescale_test + 0, data = dat)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

p2 <- ggplot(dat, aes(x = mr_rescale_test, y = count_test)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Predicted_Roulette", y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = Inf, y = Inf, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 4)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(p1, p2, ncol = 2)






```







# MSSNG (Cell 2022) 

## chr1, windowsize =50kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

dat <- read.table("/project/xinhe/xsun/mutation_rate/5.de_novo_trost/data/per_chr_processed/chr1/cv_combine_chr1_50kb_trainprop9.txt",header = T)

dat$Mb <- dat$start/1000000
dat$randeff_train <- dat$count_train/dat$mr_rescale_train

print("number of de nove mutation in testing data") 
print(table(dat$count_test)) 

ggplot(dat, aes(x = Mb, y = randeff_train )) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("Random effects") +
  ggtitle("Distribution of random effects on training set") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

dat$obs_pred_randeff <- dat$mr_rescale_test*dat$randeff_train


x_max <- max(dat$obs_pred_randeff, dat$mr_rescale_test,na.rm = T)
y_max <- max(dat$count_test)

fit1 <- lm(count_test ~ obs_pred_randeff + 0, data = dat)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

p1 <- ggplot(dat, aes(x = obs_pred_randeff, y = count_test)) +
  geom_point() +
  labs(title = "Adding random effects", x ="Predicted_random effects" , y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = Inf , y = Inf, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 4)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(count_test ~ mr_rescale_test + 0, data = dat)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

p2 <- ggplot(dat, aes(x = mr_rescale_test, y = count_test)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Predicted_Roulette", y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = Inf, y = Inf, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 4)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(p1, p2, ncol = 2)

```





## chr1, windowsize =100kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

dat <- read.table("/project/xinhe/xsun/mutation_rate/5.de_novo_trost/data/per_chr_processed/chr1/cv_combine_chr1_100kb_trainprop9.txt",header = T)

dat$Mb <- dat$start/1000000
dat$randeff_train <- dat$count_train/dat$mr_rescale_train

print("number of de nove mutation in testing data") 
print(table(dat$count_test)) 

ggplot(dat, aes(x = Mb, y = randeff_train )) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("Random effects") +
  ggtitle("Distribution of random effects on training set") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

dat$obs_pred_randeff <- dat$mr_rescale_test*dat$randeff_train


x_max <- max(dat$obs_pred_randeff, dat$mr_rescale_test,na.rm = T)
y_max <- max(dat$count_test)

fit1 <- lm(count_test ~ obs_pred_randeff + 0, data = dat)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

p1 <- ggplot(dat, aes(x = obs_pred_randeff, y = count_test)) +
  geom_point() +
  labs(title = "Adding random effects", x ="Predicted_random effects" , y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = Inf , y = Inf, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 4)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(count_test ~ mr_rescale_test + 0, data = dat)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

p2 <- ggplot(dat, aes(x = mr_rescale_test, y = count_test)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Predicted_Roulette", y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = Inf, y = Inf, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 4)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(p1, p2, ncol = 2)






```



## chr1, windowsize =200kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

dat <- read.table("/project/xinhe/xsun/mutation_rate/5.de_novo_trost/data/per_chr_processed/chr1/cv_combine_chr1_200kb_trainprop9.txt",header = T)

dat$Mb <- dat$start/1000000
dat$randeff_train <- dat$count_train/dat$mr_rescale_train

print("number of de nove mutation in testing data") 
print(table(dat$count_test)) 

ggplot(dat, aes(x = Mb, y = randeff_train )) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("Random effects") +
  ggtitle("Distribution of random effects on training set") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

dat$obs_pred_randeff <- dat$mr_rescale_test*dat$randeff_train


x_max <- max(dat$obs_pred_randeff, dat$mr_rescale_test,na.rm = T)
y_max <- max(dat$count_test)

fit1 <- lm(count_test ~ obs_pred_randeff + 0, data = dat)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

p1 <- ggplot(dat, aes(x = obs_pred_randeff, y = count_test)) +
  geom_point() +
  labs(title = "Adding random effects", x ="Predicted_random effects" , y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = Inf , y = Inf, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 4)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(count_test ~ mr_rescale_test + 0, data = dat)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

p2 <- ggplot(dat, aes(x = mr_rescale_test, y = count_test)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Predicted_Roulette", y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = Inf, y = Inf, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 4)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(p1, p2, ncol = 2)






```



## chr22, windowsize =50kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

dat <- read.table("/project/xinhe/xsun/mutation_rate/5.de_novo_trost/data/per_chr_processed/chr22/cv_combine_chr22_50kb_trainprop9.txt",header = T)

dat$Mb <- dat$start/1000000
dat$randeff_train <- dat$count_train/dat$mr_rescale_train

print("number of de nove mutation in testing data") 
print(table(dat$count_test)) 

ggplot(dat, aes(x = Mb, y = randeff_train )) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("Random effects") +
  ggtitle("Distribution of random effects on training set") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

dat$obs_pred_randeff <- dat$mr_rescale_test*dat$randeff_train


x_max <- max(dat$obs_pred_randeff, dat$mr_rescale_test,na.rm = T)
y_max <- max(dat$count_test)

fit1 <- lm(count_test ~ obs_pred_randeff + 0, data = dat)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

p1 <- ggplot(dat, aes(x = obs_pred_randeff, y = count_test)) +
  geom_point() +
  labs(title = "Adding random effects", x ="Predicted_random effects" , y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = Inf , y = Inf, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 4)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(count_test ~ mr_rescale_test + 0, data = dat)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

p2 <- ggplot(dat, aes(x = mr_rescale_test, y = count_test)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Predicted_Roulette", y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = Inf, y = Inf, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 4)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(p1, p2, ncol = 2)






```



## chr22, windowsize =100kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

dat <- read.table("/project/xinhe/xsun/mutation_rate/5.de_novo_trost/data/per_chr_processed/chr22/cv_combine_chr22_100kb_trainprop9.txt",header = T)

dat$Mb <- dat$start/1000000
dat$randeff_train <- dat$count_train/dat$mr_rescale_train

print("number of de nove mutation in testing data") 
print(table(dat$count_test)) 

ggplot(dat, aes(x = Mb, y = randeff_train )) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("Random effects") +
  ggtitle("Distribution of random effects on training set") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

dat$obs_pred_randeff <- dat$mr_rescale_test*dat$randeff_train


x_max <- max(dat$obs_pred_randeff, dat$mr_rescale_test,na.rm = T)
y_max <- max(dat$count_test)

fit1 <- lm(count_test ~ obs_pred_randeff + 0, data = dat)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

p1 <- ggplot(dat, aes(x = obs_pred_randeff, y = count_test)) +
  geom_point() +
  labs(title = "Adding random effects", x ="Predicted_random effects" , y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = Inf , y = Inf, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 4)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(count_test ~ mr_rescale_test + 0, data = dat)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

p2 <- ggplot(dat, aes(x = mr_rescale_test, y = count_test)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Predicted_Roulette", y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = Inf, y = Inf, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 4)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(p1, p2, ncol = 2)






```


## chr22, windowsize =200kb

```{r echo=F, message=FALSE, warning=FALSE, fig.width=12, fig.height=5}

dat <- read.table("/project/xinhe/xsun/mutation_rate/5.de_novo_trost/data/per_chr_processed/chr22/cv_combine_chr22_200kb_trainprop9.txt",header = T)

dat$Mb <- dat$start/1000000
dat$randeff_train <- dat$count_train/dat$mr_rescale_train

print("number of de nove mutation in testing data") 
print(table(dat$count_test)) 

ggplot(dat, aes(x = Mb, y = randeff_train )) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("Random effects") +
  ggtitle("Distribution of random effects on training set") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

dat$obs_pred_randeff <- dat$mr_rescale_test*dat$randeff_train


x_max <- max(dat$obs_pred_randeff, dat$mr_rescale_test,na.rm = T)
y_max <- max(dat$count_test)

fit1 <- lm(count_test ~ obs_pred_randeff + 0, data = dat)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

p1 <- ggplot(dat, aes(x = obs_pred_randeff, y = count_test)) +
  geom_point() +
  labs(title = "Adding random effects", x ="Predicted_random effects" , y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = Inf , y = Inf, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 4)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(count_test ~ mr_rescale_test + 0, data = dat)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

p2 <- ggplot(dat, aes(x = mr_rescale_test, y = count_test)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Predicted_Roulette", y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = Inf, y = Inf, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 4)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(p1, p2, ncol = 2)






```
