---
title: "Simulation -- combining spatial(normal prior) and non-spatial models"
author: "XSun"
date: "2024-05-06"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
---

# Introduction

## Simulation setup 

We combined the spatial and non-spatial models here

$y_i = \text{pois}(u_i\theta_i)$

where $u_i$ is expected rate from calibrated roulette. $y_i$ is the observed mutation rate. $\theta_i$ is the random effect.

1. Simulating the random effects

We simulated spatial random effects as we did in [spatial-only model](https://xsun1229.github.io/mutation_rate/denovo0422_simulation_spatial_norm.html) 
We simulated spatial random effects as we did in [non-spatial-only model](https://xsun1229.github.io/mutation_rate/denovo0422_simulation_nonspatial.html)

Then we summed up the spatial and non-spatial random effects, and normalized them, making $\text{mean}(\tilde\theta_i) =1$

```{r}

generate_randeff_nonspatial_gammaprior_spatial_normal <-  function(num_windows_all, num_windows_per_seg_min, num_windows_per_seg_max, randeff_max, num_seg_spatial,alpha,normal_sigma) {
  
  # Create the vector
  randeff <- rep(1,num_windows_all)
  
  #spatial
  segment_lengths_spatial <- ceiling(runif(num_seg_spatial, min = num_windows_per_seg_min, max = num_windows_per_seg_max))
  
  # Calculate the total length covered by spatial random effects
  total_segment_length <- sum(segment_lengths_spatial) 
  
  # Calculate the number of zeros needed
  num_zeros <- num_windows_all - total_segment_length
  
  # Randomly select positions for spatial segment starts 
  segment_starts_spatial <- sample(1:num_zeros, num_seg_spatial, replace = FALSE)
  segment_ends_spatial <- segment_starts_spatial + segment_lengths_spatial - 1
  
  # Assign random values to segments
  for (i in 1:(num_seg_spatial)) {
    start_pos <- segment_starts_spatial[i]
    end_pos <- start_pos + segment_lengths_spatial[i] - 1
    
    log_eff <- rnorm(1,mean = 0,sd=normal_sigma)
    eff <- exp(log_eff)
    randeff[start_pos:end_pos] <- eff
  }
  
  
  ## non-spatial
  alpha <- alpha
  randeff_nonspatial <- rgamma(n = num_windows_all, shape = alpha, rate = alpha)
  
  
  randeff <- randeff * randeff_nonspatial
  
  # Adjust values to meet the mean of 1
  mean_adjustment <- 1 - mean(randeff)
  randeff <- randeff + mean_adjustment
  # Clip values between 0 and randeff_max
  randeff[randeff < 0] <- 0
  #randeff[randeff > randeff_max] <- randeff_max
  
  
  return(randeff)
}

set.seed(2)
randeff <- generate_randeff_nonspatial_gammaprior_spatial_normal(num_windows_all = 3000,num_windows_per_seg_min = 20,num_windows_per_seg_max = 50,randeff_max = 3,num_seg_spatial = 10,alpha = 30,normal_sigma = 1)

plot(randeff,type = "l",ylab = "simulated random effect")
```

2. Then, we estimated $\hat\theta_i$ under different settings

  - Spatial only
  - Non-spatial only 
  - Spatial first and non-spatial the next: 
    Using $\hat\theta_i^{S} * \mu_i$ as $\mu_{i_{new}}$ in spatial model, estimating the random effect $\hat\theta_i^{NS}$ with non-spatial model
  - Non-spatial first and spatial the next: 
   Using $\hat\theta_i^{NS} * \mu_i$ as $\mu_{i_{new}}$ in spatial model, estimating the random effect $\hat\theta_i^{S}$ with spatial model


3. If estimated using the combined model, the final estimated random effect $\hat\theta_i = \hat\theta_i^{S} * \hat\theta_i^{NS}$



## Functions used

```{r}
library(ggplot2)
library(gridExtra)
plot_randeff_sp_nsp_joint <- function(dat) {
  ggplot(dat, aes(x = seq_along(y_sample))) +
  geom_line(aes(y = randeff_simulated, color = "Simulated"), alpha = 0.5) +
  geom_line(aes(y = randeff_est_sp_only, color = "Spatial-only")) +
  geom_line(aes(y = randeff_est_nsf_sp_final, color = "spatial&non-spatial"), alpha = 0.5) +
  geom_line(aes(y = randeff_est_nsp_only, color = "Non-spatial only"), alpha = 0.6) +
  labs(x = "Window index",
       y = "estimated random effects") +
  scale_color_manual(name = "Group", 
                     values = c("Simulated" = "blue", "Spatial-only" = "red", 
                                "spatial&non-spatial" = "brown", "Non-spatial only" = "green"), 
                     labels = c("Simulated", "Spatial-only", "spatial&non-spatial", "Non-spatial only")) +
  theme_minimal()
}

plot_randeff_sp_after_nsp <- function(dat) {
  ggplot(dat, aes(x = seq_along(y_sample))) +
  geom_line(aes(y = randeff_simulated, color = "Simulated"), alpha = 0.5) +
  #geom_line(aes(y = randeff_est_sp_only, color = "Spatial-only")) +
  geom_line(aes(y = randeff_est_nsf_sp, color = "spatial after nonspatial fitted")) +
  geom_line(aes(y = randeff_est_nsp_only, color = "Non-spatial only"), alpha = 0.6) +
  labs(x = "Window index",
       y = "estimated random effects") +
  scale_color_manual(name = "Group", 
                     values = c("Simulated" = "blue",
                                "spatial after nonspatial fitted" = "red", "Non-spatial only" = "green"), 
                     labels = c("Simulated","spatial after nonspatial fitted (without combining", "Non-spatial only")) +
  theme_minimal()
}

plot_randeff_nsp_after_sp <- function(dat) {
  ggplot(dat, aes(x = seq_along(y_sample))) +
  geom_line(aes(y = randeff_simulated, color = "Simulated"), alpha = 0.5) +
  #geom_line(aes(y = randeff_est_sp_only, color = "Spatial-only")) +
  geom_line(aes(y = randeff_est_sf_ns, color = "nonspatial after spacial fitted"),alpha= 0.5) +
  geom_line(aes(y = randeff_est_sp_only, color = "spatial only"), alpha = 0.6) +
  labs(x = "Window index",
       y = "estimated random effects") +
  scale_color_manual(name = "Group", 
                     values = c("Simulated" = "blue",
                                "nonspatial after spacial fitted" = "red", "spatial only" = "green"), 
                     labels = c("Simulated","nonspatial after spacial fitted (without combining", "spatial only")) +
  theme_minimal()
}

plot_model <- function(data, estimate_column, title, alpha_est = NULL) {
  print(title)
  
  # Optional printing of alpha, if provided
  if (!is.null(alpha_est)) {
    print(sprintf("alpha = %s", alpha_est))
  }
  
  # Fit linear model
  formula <- as.formula(paste(estimate_column, "~ randeff_simulated +0"))
  fit <- lm(formula, data = data)
  adj_rsq <- summary(fit)$adj.r.squared
  
  # Create the plot
  p <- ggplot(data, aes(x = randeff_simulated, y = get(estimate_column))) +
    geom_point(color = "black") +
    geom_abline(intercept = 0, slope = 1, color = "red") +
    geom_text(aes(label = paste("Adj R-sq =", round(adj_rsq, 3))),
              x = Inf, y = -Inf, hjust = 1, vjust = 0, color = "blue") +
    labs(x = paste("estimated random effect -", title),
         y = "true random effect theta_i",
         title = title) +
    theme_minimal()
  
  print(p)
}

scatter_plot_mutct <- function(data, xcol, ycol) {
  # Fit the linear model using user-defined columns
  formula <- as.formula(paste(ycol, "~", xcol,"+0"))
  fit <- lm(formula, data = data)
  adj_rsq <- summary(fit)$adj.r.squared  # Extract the adjusted R-squared
  
  # Generate the plot with user-defined columns
  ggplot(data, aes_string(x = xcol, y = ycol)) +
    geom_point(color = "black") +
    geom_abline(intercept = 0, slope = 1, color = "red") +
    geom_text(aes(label = paste0("Adj R-sq = ", round(adj_rsq, 3))),
              x = Inf, y = Inf, hjust = 1.1, vjust = 1.1,
              color = "blue", parse = FALSE) +  # Ensure label is treated as text
    labs(x = paste("Predicted y:", xcol),
         y = paste("True y (sample):", ycol)) +
    theme_minimal()
}


```

# 50kb 

## Deciding spatial model first or non-spatial model first 

### spatial window = 1mb ~ 2mb

If we estimate the random effect using non-spatial model first and then fit the spatial model, we lost the signal for non-spatial model 

```{r fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_5_minseg20_maxseg40.rdata")

plot_randeff_sp_after_nsp(dat)

```

But if we fit the spatial model first, we can keep the signal for both

```{r fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_5_minseg20_maxseg40.rdata")

plot_randeff_nsp_after_sp(dat)

```

So, in the following analysis, the joint model fits spatial effects first and the non-spatial effects the next.


## normar_var=0.5, alpha = 5, spatial window = 200kb ~ 1mb

```{r fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_5_minseg4_maxseg20.rdata")

plot_randeff_sp_nsp_joint(dat)

plot_model(data = dat, estimate_column = "randeff_est_sp_only", title = "spatial only")
plot_model(data = dat, estimate_column = "randeff_est_nsp_only", title = "non-spatial only", alpha_est = alpha_est_nsp_only)
plot_model(data = dat, estimate_column = "randeff_est_sf_ns_final", title = "spatial & non-spatial", alpha_est = alpha_est_sf_ns)

p1 <- scatter_plot_mutct(data = dat, xcol = "exp_nsf_sp_final", ycol = "y_sample")
p2 <- scatter_plot_mutct(data = dat, xcol ="roulette" , ycol = "y_sample")
grid.arrange(p1, p2, ncol = 2)

```


## normar_var=1, alpha = 5, spatial window = 200kb ~ 1mb

```{r fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_5_minseg4_maxseg20.rdata")

plot_randeff_sp_nsp_joint(dat)

plot_model(data = dat, estimate_column = "randeff_est_sp_only", title = "spatial only")
plot_model(data = dat, estimate_column = "randeff_est_nsp_only", title = "non-spatial only", alpha_est = alpha_est_nsp_only)
plot_model(data = dat, estimate_column = "randeff_est_sf_ns_final", title = "spatial & non-spatial", alpha_est = alpha_est_sf_ns)

p1 <- scatter_plot_mutct(data = dat, xcol = "exp_nsf_sp_final", ycol = "y_sample")
p2 <- scatter_plot_mutct(data = dat, xcol ="roulette" , ycol = "y_sample")
grid.arrange(p1, p2, ncol = 2)
```

## normar_var=0.5, alpha = 5, spatial window = 500kb ~ 1mb

```{r fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_5_minseg10_maxseg20.rdata")

plot_randeff_sp_nsp_joint(dat)

plot_model(data = dat, estimate_column = "randeff_est_sp_only", title = "spatial only")
plot_model(data = dat, estimate_column = "randeff_est_nsp_only", title = "non-spatial only", alpha_est = alpha_est_nsp_only)
plot_model(data = dat, estimate_column = "randeff_est_sf_ns_final", title = "spatial & non-spatial", alpha_est = alpha_est_sf_ns)

p1 <- scatter_plot_mutct(data = dat, xcol = "exp_nsf_sp_final", ycol = "y_sample")
p2 <- scatter_plot_mutct(data = dat, xcol ="roulette" , ycol = "y_sample")
grid.arrange(p1, p2, ncol = 2)
```


## normar_var=1, alpha = 5, spatial window = 500kb ~ 1mb

```{r fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_5_minseg10_maxseg20.rdata")

plot_randeff_sp_nsp_joint(dat)

plot_model(data = dat, estimate_column = "randeff_est_sp_only", title = "spatial only")
plot_model(data = dat, estimate_column = "randeff_est_nsp_only", title = "non-spatial only", alpha_est = alpha_est_nsp_only)
plot_model(data = dat, estimate_column = "randeff_est_sf_ns_final", title = "spatial & non-spatial", alpha_est = alpha_est_sf_ns)

p1 <- scatter_plot_mutct(data = dat, xcol = "exp_nsf_sp_final", ycol = "y_sample")
p2 <- scatter_plot_mutct(data = dat, xcol ="roulette" , ycol = "y_sample")
grid.arrange(p1, p2, ncol = 2)
```

## normar_var=0.5, alpha = 5, spatial window = 1mb ~ 2mb

```{r fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_5_minseg20_maxseg40.rdata")

plot_randeff_sp_nsp_joint(dat)

plot_model(data = dat, estimate_column = "randeff_est_sp_only", title = "spatial only")
plot_model(data = dat, estimate_column = "randeff_est_nsp_only", title = "non-spatial only", alpha_est = alpha_est_nsp_only)
plot_model(data = dat, estimate_column = "randeff_est_sf_ns_final", title = "spatial & non-spatial", alpha_est = alpha_est_sf_ns)

p1 <- scatter_plot_mutct(data = dat, xcol = "exp_nsf_sp_final", ycol = "y_sample")
p2 <- scatter_plot_mutct(data = dat, xcol ="roulette" , ycol = "y_sample")
grid.arrange(p1, p2, ncol = 2)
```


## normar_var=1, alpha = 5, spatial window = 1mb ~ 2mb

```{r fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_5_minseg20_maxseg40.rdata")

plot_randeff_sp_nsp_joint(dat)

plot_model(data = dat, estimate_column = "randeff_est_sp_only", title = "spatial only")
plot_model(data = dat, estimate_column = "randeff_est_nsp_only", title = "non-spatial only", alpha_est = alpha_est_nsp_only)
plot_model(data = dat, estimate_column = "randeff_est_sf_ns_final", title = "spatial & non-spatial", alpha_est = alpha_est_sf_ns)

p1 <- scatter_plot_mutct(data = dat, xcol = "exp_nsf_sp_final", ycol = "y_sample")
p2 <- scatter_plot_mutct(data = dat, xcol ="roulette" , ycol = "y_sample")
grid.arrange(p1, p2, ncol = 2)
```


## normar_var=0.5, alpha = 20, spatial window = 200kb ~ 1mb

```{r fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_20_minseg4_maxseg20.rdata")

plot_randeff_sp_nsp_joint(dat)

plot_model(data = dat, estimate_column = "randeff_est_sp_only", title = "spatial only")
plot_model(data = dat, estimate_column = "randeff_est_nsp_only", title = "non-spatial only", alpha_est = alpha_est_nsp_only)
plot_model(data = dat, estimate_column = "randeff_est_sf_ns_final", title = "spatial & non-spatial", alpha_est = alpha_est_sf_ns)

p1 <- scatter_plot_mutct(data = dat, xcol = "exp_nsf_sp_final", ycol = "y_sample")
p2 <- scatter_plot_mutct(data = dat, xcol ="roulette" , ycol = "y_sample")
grid.arrange(p1, p2, ncol = 2)
```


## normar_var=1, alpha = 20, spatial window = 200kb ~ 1mb

```{r fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_20_minseg4_maxseg20.rdata")

plot_randeff_sp_nsp_joint(dat)

plot_model(data = dat, estimate_column = "randeff_est_sp_only", title = "spatial only")
plot_model(data = dat, estimate_column = "randeff_est_nsp_only", title = "non-spatial only", alpha_est = alpha_est_nsp_only)
plot_model(data = dat, estimate_column = "randeff_est_sf_ns_final", title = "spatial & non-spatial", alpha_est = alpha_est_sf_ns)

p1 <- scatter_plot_mutct(data = dat, xcol = "exp_nsf_sp_final", ycol = "y_sample")
p2 <- scatter_plot_mutct(data = dat, xcol ="roulette" , ycol = "y_sample")
grid.arrange(p1, p2, ncol = 2)
```

## normar_var=0.5, alpha = 20, spatial window = 500kb ~ 1mb

```{r fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_20_minseg10_maxseg20.rdata")

plot_randeff_sp_nsp_joint(dat)

plot_model(data = dat, estimate_column = "randeff_est_sp_only", title = "spatial only")
plot_model(data = dat, estimate_column = "randeff_est_nsp_only", title = "non-spatial only", alpha_est = alpha_est_nsp_only)
plot_model(data = dat, estimate_column = "randeff_est_sf_ns_final", title = "spatial & non-spatial", alpha_est = alpha_est_sf_ns)

p1 <- scatter_plot_mutct(data = dat, xcol = "exp_nsf_sp_final", ycol = "y_sample")
p2 <- scatter_plot_mutct(data = dat, xcol ="roulette" , ycol = "y_sample")
grid.arrange(p1, p2, ncol = 2)
```


## normar_var=1, alpha = 20, spatial window = 500kb ~ 1mb

```{r fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_20_minseg10_maxseg20.rdata")

plot_randeff_sp_nsp_joint(dat)

plot_model(data = dat, estimate_column = "randeff_est_sp_only", title = "spatial only")
plot_model(data = dat, estimate_column = "randeff_est_nsp_only", title = "non-spatial only", alpha_est = alpha_est_nsp_only)
plot_model(data = dat, estimate_column = "randeff_est_sf_ns_final", title = "spatial & non-spatial", alpha_est = alpha_est_sf_ns)

p1 <- scatter_plot_mutct(data = dat, xcol = "exp_nsf_sp_final", ycol = "y_sample")
p2 <- scatter_plot_mutct(data = dat, xcol ="roulette" , ycol = "y_sample")
grid.arrange(p1, p2, ncol = 2)
```

## normar_var=0.5, alpha = 20, spatial window = 1mb ~ 2mb

```{r fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma0.5_alphansp_20_minseg20_maxseg40.rdata")

plot_randeff_sp_nsp_joint(dat)

plot_model(data = dat, estimate_column = "randeff_est_sp_only", title = "spatial only")
plot_model(data = dat, estimate_column = "randeff_est_nsp_only", title = "non-spatial only", alpha_est = alpha_est_nsp_only)
plot_model(data = dat, estimate_column = "randeff_est_sf_ns_final", title = "spatial & non-spatial", alpha_est = alpha_est_sf_ns)

p1 <- scatter_plot_mutct(data = dat, xcol = "exp_nsf_sp_final", ycol = "y_sample")
p2 <- scatter_plot_mutct(data = dat, xcol ="roulette" , ycol = "y_sample")
grid.arrange(p1, p2, ncol = 2)
```


## normar_var=1, alpha = 20, spatial window = 1mb ~ 2mb

```{r fig.width=12, fig.height=5}

load("/project/xinhe/xsun/mutation_rate/9.denovo_simulation_0422/results/spatial_nonspatial_sp_normal/simu_sp_nsp_chr1_50kb_maxiter50_normal_sigma1_alphansp_20_minseg20_maxseg40.rdata")

plot_randeff_sp_nsp_joint(dat)

plot_model(data = dat, estimate_column = "randeff_est_sp_only", title = "spatial only")
plot_model(data = dat, estimate_column = "randeff_est_nsp_only", title = "non-spatial only", alpha_est = alpha_est_nsp_only)
plot_model(data = dat, estimate_column = "randeff_est_sf_ns_final", title = "spatial & non-spatial", alpha_est = alpha_est_sf_ns)

p1 <- scatter_plot_mutct(data = dat, xcol = "exp_nsf_sp_final", ycol = "y_sample")
p2 <- scatter_plot_mutct(data = dat, xcol ="roulette" , ycol = "y_sample")
grid.arrange(p1, p2, ncol = 2)
```