---
title: "Cross Validation -- SSC"
author: "XSun"
date: "2024-06-12"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
---

# Introduction

We cleaned the real data [here](https://xsun1229.github.io/mutation_rate/denovo0422_realdata_data_cleaning.html). 

```{r warning=F, message=F}
library(ggplot2)
library(gridExtra)

scatter_plot_mutct <- function(data, xcol, ycol, title=NULL) {
  # Fit the linear model using user-defined columns
  formula <- as.formula(paste(ycol, "~", xcol, "+0"))
  fit <- lm(formula, data = data)
  adj_rsq <- summary(fit)$adj.r.squared  # Extract the adjusted R-squared

  
  ggplot(data) +
  geom_point(data = data, aes_string(x = xcol, y = ycol),color = "black") +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  annotate("text", x = Inf, y = Inf, label = paste0("Adj R-sq = ", round(adj_rsq, 3)),
           hjust = 1.1, vjust = 1.1, color = "blue", parse = FALSE) +  # Adjust text positioning and color
  labs(x = paste("Predicted mutation count on testing set:", xcol),
       y = paste("Observed mutation count on testing set:", ycol),
       title = title) +
  theme_minimal()
  
}

plot_scatter_randeff <- function(data) {
  ggplot(data = df_compare) +aes(x=seg_est,y=chr_est) +
  geom_point() +
  labs(x = "nugget effects estimated from partitioned segments",
       y= "nugget effects estimated from whole chromosome") +
  geom_abline(slope = 1, intercept = 0, col="red") +
  theme_minimal()
}

plot_randeff_genome <- function(data) {
  ggplot(df_compare, aes(x =  window_start/1000000)) +
  geom_line(aes(y = seg_est, color = "estimated from partitioned segments"), alpha = 0.4) +
  geom_line(aes(y = chr_est, color = "estimated from whole chr"), alpha = 0.3) +
  geom_line(aes(y = fold_change, color = "observed fold change"), alpha = 0.1) +
  labs(x = "gemonic position (mb)",
       y = "estimated random effects") +
  scale_color_manual(name = "Group",
                     values = c("estimated from partitioned segments" = "blue",
                                "estimated from whole chr" = "green", "observed fold change" = "red"),
                     labels = c("estimated from partitioned segments","estimated from whole chr", "observed fold change")) +
  theme_minimal()
}

```



# Setting one: partition the gaps based on large gaps 

* For large gap: treat as different segments. So for chr1, there will be 3 segments. 
* For small gaps like 50kb, treat them as continuous ones. 

## The whole window prediction

* Run smashgen on the whole window, to get an estimation for the local nugget effects 

`fit = ebps(df_seg_non0$obs_sum,df_seg_non0$rl_rescaled_sum,smooth_control = list(wave_trans='ndwt',ndwt_method='smash'), general_control = list(verbose=T,printevery=1, maxiter=50))`

* Predict the mutation rate on the whole window, based on the rescaled Roulette estimation.

`df_per_window_data$sum_pred <- df_per_window_data$randeff_est*df_per_window_data$rl_rescaled_sum`

### 10kb

```{r}

load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_10kb_fitted_testpred.rdata")

df_per_window_data$sum_pred <- df_per_window_data$randeff_est*df_per_window_data$rl_rescaled_sum
data <- df_per_window_data[complete.cases(df_per_window_data),]

p1 <- scatter_plot_mutct(data = data, xcol = "sum_pred", ycol = "obs_sum",title = "smashgen prediction")
p2 <- scatter_plot_mutct(data = data, xcol = "rl_rescaled_sum", ycol = "obs_sum",title = "Roulette baseline")

grid.arrange(p1, p2, ncol = 2)
```

### 30kb

```{r}

load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_30kb_fitted_testpred.rdata")

df_per_window_data$sum_pred <- df_per_window_data$randeff_est*df_per_window_data$rl_rescaled_sum
data <- df_per_window_data[complete.cases(df_per_window_data),]

p1 <- scatter_plot_mutct(data = data, xcol = "sum_pred", ycol = "obs_sum",title = "smashgen prediction")
p2 <- scatter_plot_mutct(data = data, xcol = "rl_rescaled_sum", ycol = "obs_sum",title = "Roulette baseline")

grid.arrange(p1, p2, ncol = 2)
```

### 50kb

```{r}

load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_50kb_fitted_testpred.rdata")

df_per_window_data$sum_pred <- df_per_window_data$randeff_est*df_per_window_data$rl_rescaled_sum
data <- df_per_window_data[complete.cases(df_per_window_data),]

p1 <- scatter_plot_mutct(data = data, xcol = "sum_pred", ycol = "obs_sum",title = "smashgen prediction")
p2 <- scatter_plot_mutct(data = data, xcol = "rl_rescaled_sum", ycol = "obs_sum",title = "Roulette baseline")

grid.arrange(p1, p2, ncol = 2)
```


### 100kb

```{r}

load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_100kb_fitted_testpred.rdata")

df_per_window_data$sum_pred <- df_per_window_data$randeff_est*df_per_window_data$rl_rescaled_sum
data <- df_per_window_data[complete.cases(df_per_window_data),]

p1 <- scatter_plot_mutct(data = data, xcol = "sum_pred", ycol = "obs_sum",title = "smashgen prediction")
p2 <- scatter_plot_mutct(data = data, xcol = "rl_rescaled_sum", ycol = "obs_sum",title = "Roulette baseline")

grid.arrange(p1, p2, ncol = 2)
```

## Cross validation

For each window, the first 90% base pairs are treated as training data, the remaining 10% are testing data

* Run smashgen on the training data, to get an estimation for the local nugget effects 

`fit = ebps(df_seg_non0$obs_train,df_seg_non0$rl_rescaled_train,smooth_control = list(wave_trans='ndwt',ndwt_method='smash'), general_control = list(verbose=T,printevery=1, maxiter=50))`

* Predict the mutation rate on testing data, based on the rescaled Roulette estimation.

`df_per_window_data$test_pred <- df_per_window_data$randeff_est_train*df_per_window_data$rl_rescaled_test`



### 10kb

```{r}

load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_10kb_fitted_testpred.rdata")

data <- df_per_window_data[complete.cases(df_per_window_data),]

p1 <- scatter_plot_mutct(data = data, xcol = "test_pred", ycol = "obs_test",title = "smashgen prediction")
p2 <- scatter_plot_mutct(data = data, xcol = "rl_rescaled_test", ycol = "obs_test",title = "Roulette baseline")

grid.arrange(p1, p2, ncol = 2)
```

### 30kb

```{r}

load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_30kb_fitted_testpred.rdata")

data <- df_per_window_data[complete.cases(df_per_window_data),]

p1 <- scatter_plot_mutct(data = data, xcol = "test_pred", ycol = "obs_test",title = "smashgen prediction")
p2 <- scatter_plot_mutct(data = data, xcol = "rl_rescaled_test", ycol = "obs_test",title = "Roulette baseline")

grid.arrange(p1, p2, ncol = 2)
```


### 50kb

```{r}

load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_50kb_fitted_testpred.rdata")

data <- df_per_window_data[complete.cases(df_per_window_data),]

p1 <- scatter_plot_mutct(data = data, xcol = "test_pred", ycol = "obs_test",title = "smashgen prediction")
p2 <- scatter_plot_mutct(data = data, xcol = "rl_rescaled_test", ycol = "obs_test",title = "Roulette baseline")

grid.arrange(p1, p2, ncol = 2)
```


### 100kb

```{r}

load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_100kb_fitted_testpred.rdata")

data <- df_per_window_data[complete.cases(df_per_window_data),]

p1 <- scatter_plot_mutct(data = data, xcol = "test_pred", ycol = "obs_test",title = "smashgen prediction")
p2 <- scatter_plot_mutct(data = data, xcol = "rl_rescaled_test", ycol = "obs_test",title = "Roulette baseline")

grid.arrange(p1, p2, ncol = 2)
```

# Setting two: treat the whole genome as one segment

## The whole window prediction

### 10kb

```{r}

load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_10kb_fitted_testpred_chr.rdata")

df_per_window_data$sum_pred <- df_per_window_data$randeff_est*df_per_window_data$rl_rescaled_sum
data <- df_per_window_data[complete.cases(df_per_window_data),]

p1 <- scatter_plot_mutct(data = data, xcol = "sum_pred", ycol = "obs_sum",title = "smashgen prediction")
p2 <- scatter_plot_mutct(data = data, xcol = "rl_rescaled_sum", ycol = "obs_sum",title = "Roulette baseline")

grid.arrange(p1, p2, ncol = 2)
```

### 30kb

```{r}

load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_30kb_fitted_testpred_chr.rdata")

df_per_window_data$sum_pred <- df_per_window_data$randeff_est*df_per_window_data$rl_rescaled_sum
data <- df_per_window_data[complete.cases(df_per_window_data),]

p1 <- scatter_plot_mutct(data = data, xcol = "sum_pred", ycol = "obs_sum",title = "smashgen prediction")
p2 <- scatter_plot_mutct(data = data, xcol = "rl_rescaled_sum", ycol = "obs_sum",title = "Roulette baseline")

grid.arrange(p1, p2, ncol = 2)
```

### 50kb

```{r}

load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_50kb_fitted_testpred_chr.rdata")

df_per_window_data$sum_pred <- df_per_window_data$randeff_est*df_per_window_data$rl_rescaled_sum
data <- df_per_window_data[complete.cases(df_per_window_data),]

p1 <- scatter_plot_mutct(data = data, xcol = "sum_pred", ycol = "obs_sum",title = "smashgen prediction")
p2 <- scatter_plot_mutct(data = data, xcol = "rl_rescaled_sum", ycol = "obs_sum",title = "Roulette baseline")

grid.arrange(p1, p2, ncol = 2)
```


### 100kb

```{r}

load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_100kb_fitted_testpred_chr.rdata")

df_per_window_data$sum_pred <- df_per_window_data$randeff_est*df_per_window_data$rl_rescaled_sum
data <- df_per_window_data[complete.cases(df_per_window_data),]

p1 <- scatter_plot_mutct(data = data, xcol = "sum_pred", ycol = "obs_sum",title = "smashgen prediction")
p2 <- scatter_plot_mutct(data = data, xcol = "rl_rescaled_sum", ycol = "obs_sum",title = "Roulette baseline")

grid.arrange(p1, p2, ncol = 2)
```

## Cross validation

For each window, the first 90% base pairs are treated as training data, the remaining 10% are testing data

* Run smashgen on the training data, to get an estimation for the local nugget effects 

`fit = ebps(df_seg_non0$obs_train,df_seg_non0$rl_rescaled_train,smooth_control = list(wave_trans='ndwt',ndwt_method='smash'), general_control = list(verbose=T,printevery=1, maxiter=50))`

* Predict the mutation rate on testing data, based on the rescaled Roulette estimation.

`df_per_window_data$test_pred <- df_per_window_data$randeff_est_train*df_per_window_data$rl_rescaled_test`



### 10kb

```{r}

load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_10kb_fitted_testpred_chr.rdata")

data <- df_per_window_data[complete.cases(df_per_window_data),]

p1 <- scatter_plot_mutct(data = data, xcol = "test_pred", ycol = "obs_test",title = "smashgen prediction")
p2 <- scatter_plot_mutct(data = data, xcol = "rl_rescaled_test", ycol = "obs_test",title = "Roulette baseline")

grid.arrange(p1, p2, ncol = 2)
```

### 30kb

```{r}

load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_30kb_fitted_testpred_chr.rdata")

data <- df_per_window_data[complete.cases(df_per_window_data),]

p1 <- scatter_plot_mutct(data = data, xcol = "test_pred", ycol = "obs_test",title = "smashgen prediction")
p2 <- scatter_plot_mutct(data = data, xcol = "rl_rescaled_test", ycol = "obs_test",title = "Roulette baseline")

grid.arrange(p1, p2, ncol = 2)
```


### 50kb

```{r}

load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_50kb_fitted_testpred_chr.rdata")

data <- df_per_window_data[complete.cases(df_per_window_data),]

p1 <- scatter_plot_mutct(data = data, xcol = "test_pred", ycol = "obs_test",title = "smashgen prediction")
p2 <- scatter_plot_mutct(data = data, xcol = "rl_rescaled_test", ycol = "obs_test",title = "Roulette baseline")

grid.arrange(p1, p2, ncol = 2)
```


### 100kb

```{r}

load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_100kb_fitted_testpred_chr.rdata")

data <- df_per_window_data[complete.cases(df_per_window_data),]

p1 <- scatter_plot_mutct(data = data, xcol = "test_pred", ycol = "obs_test",title = "smashgen prediction")
p2 <- scatter_plot_mutct(data = data, xcol = "rl_rescaled_test", ycol = "obs_test",title = "Roulette baseline")

grid.arrange(p1, p2, ncol = 2)
```





<!-- # Comparing the nugget effects estimated by smashgen for setting 1 and 2 -->

<!-- ### 10kb -->

<!-- ```{r message=FALSE, warning=FALSE} -->

<!-- df_seg <- get(load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_10kb_fitted_testpred.rdata")) -->
<!-- df_chr <- get(load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_10kb_fitted_testpred_chr.rdata")) -->


<!-- df_compare <- data.frame(seg_est = df_seg$randeff_est, -->
<!--                          chr_est = df_chr$randeff_est, -->
<!--                          fold_change = df_seg$foldchange, -->
<!--                          window_start = df_seg$Window_Start) -->


<!-- ggplot(data = df_compare) +aes(x=seg_est,y=chr_est) + -->
<!--   geom_point() + -->
<!--   labs(x = "nugget effects estimated from partitioned segments", -->
<!--        y= "nugget effects estimated from whole chromosome") + -->
<!--   geom_abline(slope = 1, intercept = 0, col="red") +  -->
<!--   theme_minimal() -->

<!-- ggplot(df_compare, aes(x =  window_start/1000000)) + -->
<!--   geom_line(aes(y = seg_est, color = "estimated from partitioned segments"), alpha = 0.4) + -->
<!--   geom_line(aes(y = chr_est, color = "estimated from whole chr"), alpha = 0.3) + -->
<!--   geom_line(aes(y = fold_change, color = "observed fold change"), alpha = 0.1) + -->
<!--   labs(x = "gemonic position (mb)", -->
<!--        y = "estimated random effects") + -->
<!--   scale_color_manual(name = "Group",  -->
<!--                      values = c("estimated from partitioned segments" = "blue", -->
<!--                                 "estimated from whole chr" = "green", "observed fold change" = "red"),  -->
<!--                      labels = c("estimated from partitioned segments","estimated from whole chr", "observed fold change")) + -->
<!--   theme_minimal() -->


<!-- ``` -->


<!-- ### 30kb -->

<!-- ```{r message=FALSE, warning=FALSE} -->

<!-- df_seg <- get(load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_30kb_fitted_testpred.rdata")) -->
<!-- df_chr <- get(load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_30kb_fitted_testpred_chr.rdata")) -->

<!-- df_compare <- data.frame(seg_est = df_seg$randeff_est, -->
<!--                          chr_est = df_chr$randeff_est) -->



<!-- ggplot(data = df_compare) +aes(x=seg_est,y=chr_est) + -->
<!--   geom_point() + -->
<!--   labs(x = "nugget effects estimated from partitioned segments", -->
<!--        y= "nugget effects estimated from whole chromosome") + -->
<!--   geom_abline(slope = 1, intercept = 0, col="red") +  -->
<!--   theme_minimal() -->

<!-- ``` -->


<!-- ### 50kb -->

<!-- ```{r message=FALSE, warning=FALSE} -->

<!-- df_seg <- get(load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_50kb_fitted_testpred.rdata")) -->
<!-- df_chr <- get(load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_50kb_fitted_testpred_chr.rdata")) -->

<!-- df_compare <- data.frame(seg_est = df_seg$randeff_est, -->
<!--                          chr_est = df_chr$randeff_est) -->



<!-- ggplot(data = df_compare) +aes(x=seg_est,y=chr_est) + -->
<!--   geom_point() + -->
<!--   labs(x = "nugget effects estimated from partitioned segments", -->
<!--        y= "nugget effects estimated from whole chromosome") + -->
<!--   geom_abline(slope = 1, intercept = 0, col="red") +  -->
<!--   theme_minimal() -->

<!-- ``` -->


<!-- ### 100kb -->

<!-- ```{r message=FALSE, warning=FALSE} -->

<!-- df_seg <- get(load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_100kb_fitted_testpred.rdata")) -->
<!-- df_chr <- get(load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_100kb_fitted_testpred_chr.rdata")) -->

<!-- df_compare <- data.frame(seg_est = df_seg$randeff_est, -->
<!--                          chr_est = df_chr$randeff_est) -->


<!-- ggplot(data = df_compare) +aes(x=seg_est,y=chr_est) + -->
<!--   geom_point() + -->
<!--   labs(x = "nugget effects estimated from partitioned segments", -->
<!--        y= "nugget effects estimated from whole chromosome") + -->
<!--   geom_abline(slope = 1, intercept = 0, col="red") +  -->
<!--   theme_minimal() -->

<!-- ``` -->







# Comparing the nugget effects estimated by smashgen for setting 1 and 2

### 10kb

```{r message=FALSE, warning=FALSE, fig.width=15, fig.height=7}

df_seg <- get(load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_10kb_fitted_testpred.rdata"))
df_chr <- get(load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_10kb_fitted_testpred_chr.rdata"))


df_compare <- data.frame(seg_est = df_seg$randeff_est,
                         chr_est = df_chr$randeff_est,
                         fold_change = df_seg$foldchange,
                         window_start = df_seg$Window_Start)

plot_scatter_randeff(data)
plot_randeff_genome(data)


```


### 30kb

```{r message=FALSE, warning=FALSE, fig.width=15, fig.height=7}

df_seg <- get(load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_30kb_fitted_testpred.rdata"))
df_chr <- get(load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_30kb_fitted_testpred_chr.rdata"))

df_compare <- data.frame(seg_est = df_seg$randeff_est,
                         chr_est = df_chr$randeff_est,
                         fold_change = df_seg$foldchange,
                         window_start = df_seg$Window_Start)

plot_scatter_randeff(data)
plot_randeff_genome(data)
```


### 50kb

```{r message=FALSE, warning=FALSE, fig.width=15, fig.height=7}

df_seg <- get(load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_50kb_fitted_testpred.rdata"))
df_chr <- get(load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_50kb_fitted_testpred_chr.rdata"))

df_compare <- data.frame(seg_est = df_seg$randeff_est,
                         chr_est = df_chr$randeff_est,
                         fold_change = df_seg$foldchange,
                         window_start = df_seg$Window_Start)

plot_scatter_randeff(data)
plot_randeff_genome(data)

```


### 100kb

```{r message=FALSE, warning=FALSE, fig.width=15, fig.height=7}

df_seg <- get(load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_100kb_fitted_testpred.rdata"))
df_chr <- get(load("/project/xinhe/xsun/mutation_rate/10.denovo_real/results/ssc_trost_per_window_chr1_100kb_fitted_testpred_chr.rdata"))

df_compare <- data.frame(seg_est = df_seg$randeff_est,
                         chr_est = df_chr$randeff_est,
                         fold_change = df_seg$foldchange,
                         window_start = df_seg$Window_Start)

plot_scatter_randeff(data)
plot_randeff_genome(data)

```

