---
title: "Calibrating An's de novo data with Roulette (Science 2018)"
author: "XSun"
date: "2023-07-04"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(data.table)
library(ggplot2)
```


Data source: 

[Genome-wide de novo risk score implicates promoter variation in autism spectrum disorder [science 2018]](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6432922/)

7,608 samples in 1,902 autism spectrum disorder (ASD) families, identifying 255,106 de novo mutations in total, 233,954 of them are de novo SNVs.

The samples are from Simons Simplex Collection.

# Data overview

```{r echo=FALSE, message=FALSE, warning=FALSE}
print("all 6 mutation types")
an <- fread("/project/xinhe/xsun/mutation_rate/data/an_denovo/processed/alltype.bed")
colnames(an) <-c("chr","pos","ref","alt")
an[1:10,]
```

| mutation types | # of mutations        |
|----------------|-----------------------|
| a->c t->g      | 7938 + 7927 = 15865   |
| a->g t->c      | 30829 + 30365 = 61194 |
| a->t t->a      | 7434 + 7717 = 15151   |
| c->a g->t      | 11027 + 11276 = 22303 |
| c->g g->c      | 10952 + 10833 = 21785 |
| c->t g->a      | 48750 + 48906 = 97656 |


```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=5}

data <- data.frame(
  category = c("a->c", "a->g", "a->t", "c->a", "c->g", "c->t"),
  percentage = c(6.78,26.16, 6.48, 9.53, 9.31,41.74)
)
data$angle <- cumsum(data$percentage) - (0.5 * data$percentage)
ggplot(data, aes(x = "", y = percentage, fill = category)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  theme_void() +
  theme(legend.position = "bottom") +
  # Adjust the position of the labels
  geom_label(
    aes(label = paste0(category, "\n", percentage, "%")),
    position = position_stack(vjust = 0.5),
    size = 3.5,
    show.legend = FALSE
  )
```

# Rescaling

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=5}

load("/project/xinhe/xsun/mutation_rate/4.de_novo_an/results/scaling_factor.rdata")


print("fitting: observed (An et al.) ~ expected (Roulette) + 0")
fit <- lm(denovo_per_chr ~ rl_per_chr + 0)
print(summary(fit))

data <- data.frame(expected = rl_per_chr, observed = denovo_per_chr)
print(data)

# Create the plot
ggplot(data, aes(x = expected, y = observed)) +
  geom_point() +
  geom_abline(slope = fit$coefficients[1], intercept = 0, color = "red") +
  xlab("Expected_Roulette") +
  ylab("Observed_An et al.") 

print(paste0("the scalling factor is ",round(fit$coefficients[1],digits = 5)))
print(paste0("obs = ",round(fit$coefficients[1],digits = 5),"* exp"))
```


# Random effect distribution (after rescaling)

## Window size: 10kb

### chr1

All 6 mutation types, random effect: obs/exp

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

data <- fread("/project/xinhe/xsun/mutation_rate/4.de_novo_an/data/per_chr_processed/chr1/rl_an_combine_chr1_10000.txt")
colnames(data) <- c("chr","start","end","obs","exp")
data$Mb <- data$start/1000000
data$randeff <- data$obs / data$exp

print("the number of observed de novo mutations per window")
table(data$obs)

ggplot(data, aes(x = Mb, y = obs)) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of observed de novo mutations") +
  ggtitle("Distribution of observed de novo mutations") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


ggplot(data, aes(x = Mb, y = randeff)) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("Random effects") +
  ggtitle("Distribution of random effects") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


```

### chr22

All 6 mutation types, random effect: obs/exp

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

data <- fread("/project/xinhe/xsun/mutation_rate/4.de_novo_an/data/per_chr_processed/chr22/rl_an_combine_chr22_10000.txt")
colnames(data) <- c("chr","start","end","obs","exp")
data$Mb <- data$start/1000000
data$randeff <- data$obs / data$exp

print("the number of observed de novo mutations per window")
table(data$obs)

ggplot(data, aes(x = Mb, y = obs)) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of observed de novo mutations") +
  ggtitle("Distribution of observed de novo mutations") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


ggplot(data, aes(x = Mb, y = randeff)) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("Random effects") +
  ggtitle("Distribution of random effects") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


```

## Window size: 50kb

### chr1

All 6 mutation types, random effect: obs/exp

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

data <- fread("/project/xinhe/xsun/mutation_rate/4.de_novo_an/data/per_chr_processed/chr1/rl_an_combine_chr1_50000.txt")
colnames(data) <- c("chr","start","end","obs","exp")
data$Mb <- data$start/1000000
data$randeff <- data$obs / data$exp

print("the number of observed de novo mutations per window")
table(data$obs)

ggplot(data, aes(x = Mb, y = obs)) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of observed de novo mutations") +
  ggtitle("Distribution of observed de novo mutations") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


ggplot(data, aes(x = Mb, y = randeff)) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("Random effects") +
  ggtitle("Distribution of random effects") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


```

### chr22

All 6 mutation types, random effect: obs/exp

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

data <- fread("/project/xinhe/xsun/mutation_rate/4.de_novo_an/data/per_chr_processed/chr22/rl_an_combine_chr22_50000.txt")
colnames(data) <- c("chr","start","end","obs","exp")
data$Mb <- data$start/1000000
data$randeff <- data$obs / data$exp

print("the number of observed de novo mutations per window")
table(data$obs)

ggplot(data, aes(x = Mb, y = obs)) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of observed de novo mutations") +
  ggtitle("Distribution of observed de novo mutations") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


ggplot(data, aes(x = Mb, y = randeff)) +
  geom_point(size=0.2) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("Random effects") +
  ggtitle("Distribution of random effects") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


```
