---
title: "Modelling the random effect with smash"
author: "XSun"
date: "2023-04-06"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
suppressMessages(library(ggplot2))
suppressMessages(library(gridExtra))
```


# Introduction

We grouped the chr22 into 100bp, 1kb, 10kb windows. In each window, we sampled 90% sites as training data, and the remaining 10% as the testing data. For comparison, the training and testing data here are the same with [that without smoothing](https://xsun1229.github.io/mutation_rate/gnomad_random_obsexp_cv.html)

We

1. computed the random effect (eff_rand) in this window using these training sites. 

2. did log transform for eff_rand and got lg_eff_rand

3. smoothed the lg_eff_rand across all windows using [smashr](https://github.com/stephenslab/smashr), got mean_lg_eff_rand

4. eff_rand_sm <- exp(mean_lg_eff_rand)

5. predicted the observed number of SNV with eff_rand_sm on the testing sites in each window.

6. fit SNV_obs ~ SNV_obs_smash + 0 

For the windows with 0 SNVs: we replaced the 0s with a pseudocount of 0.5. 

# 1kb

## All 6 mutation types
```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_1kb_pseudo05.rdata")

x_max <- max(combine$SNV_obs_unsam)
y_max <- max(combine$SNV_exp_unsam, combine$obs_randeff_umsam)

fit1 <- lm(obs_randeff_umsam ~ SNV_obs_unsam  + 0, data = combine)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

smash_plot <- ggplot(combine, aes(x = SNV_obs_unsam, y = obs_randeff_umsam)) +
  geom_point() +
  labs(title = "SMASH", x = "Observed", y = "Predicted_SMASH") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +  
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(SNV_exp_unsam ~ SNV_obs_unsam + 0, data = combine)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

roulette_plot <- ggplot(combine, aes(x =SNV_obs_unsam , y = SNV_exp_unsam)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Observed", y = "Predicted_Roulette") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) + 
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(smash_plot, roulette_plot, ncol = 2)
```


## A - G

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_ag_1kb_pseudo05.rdata")

x_max <- max(combine$SNV_obs_unsam)
y_max <- max(combine$SNV_exp_unsam, combine$obs_randeff_umsam)

fit1 <- lm(obs_randeff_umsam ~ SNV_obs_unsam  + 0, data = combine)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

smash_plot <- ggplot(combine, aes(x = SNV_obs_unsam, y = obs_randeff_umsam)) +
  geom_point() +
  labs(title = "SMASH", x = "Observed", y = "Predicted_SMASH") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +  
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(SNV_exp_unsam ~ SNV_obs_unsam + 0, data = combine)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

roulette_plot <- ggplot(combine, aes(x =SNV_obs_unsam , y = SNV_exp_unsam)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Observed", y = "Predicted_Roulette") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) + 
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(smash_plot, roulette_plot, ncol = 2)

```


## A - C

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_ac_1kb_pseudo05.rdata")

x_max <- max(combine$SNV_obs_unsam)
y_max <- max(combine$SNV_exp_unsam, combine$obs_randeff_umsam)

fit1 <- lm(obs_randeff_umsam ~ SNV_obs_unsam  + 0, data = combine)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

smash_plot <- ggplot(combine, aes(x = SNV_obs_unsam, y = obs_randeff_umsam)) +
  geom_point() +
  labs(title = "SMASH", x = "Observed", y = "Predicted_SMASH") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +  
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(SNV_exp_unsam ~ SNV_obs_unsam + 0, data = combine)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

roulette_plot <- ggplot(combine, aes(x =SNV_obs_unsam , y = SNV_exp_unsam)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Observed", y = "Predicted_Roulette") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) + 
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(smash_plot, roulette_plot, ncol = 2)

```

## A - T

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_at_1kb_pseudo05.rdata")

x_max <- max(combine$SNV_obs_unsam)
y_max <- max(combine$SNV_exp_unsam, combine$obs_randeff_umsam)

fit1 <- lm(obs_randeff_umsam ~ SNV_obs_unsam  + 0, data = combine)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

smash_plot <- ggplot(combine, aes(x = SNV_obs_unsam, y = obs_randeff_umsam)) +
  geom_point() +
  labs(title = "SMASH", x = "Observed", y = "Predicted_SMASH") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +  
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(SNV_exp_unsam ~ SNV_obs_unsam + 0, data = combine)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

roulette_plot <- ggplot(combine, aes(x =SNV_obs_unsam , y = SNV_exp_unsam)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Observed", y = "Predicted_Roulette") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) + 
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(smash_plot, roulette_plot, ncol = 2)

```

## C - A

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_ca_1kb_pseudo05.rdata")

x_max <- max(combine$SNV_obs_unsam)
y_max <- max(combine$SNV_exp_unsam, combine$obs_randeff_umsam)

fit1 <- lm(obs_randeff_umsam ~ SNV_obs_unsam  + 0, data = combine)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

smash_plot <- ggplot(combine, aes(x = SNV_obs_unsam, y = obs_randeff_umsam)) +
  geom_point() +
  labs(title = "SMASH", x = "Observed", y = "Predicted_SMASH") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +  
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(SNV_exp_unsam ~ SNV_obs_unsam + 0, data = combine)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

roulette_plot <- ggplot(combine, aes(x =SNV_obs_unsam , y = SNV_exp_unsam)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Observed", y = "Predicted_Roulette") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) + 
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(smash_plot, roulette_plot, ncol = 2)

```

## C - G
```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_cg_1kb_pseudo05.rdata")

x_max <- max(combine$SNV_obs_unsam)
y_max <- max(combine$SNV_exp_unsam, combine$obs_randeff_umsam)

fit1 <- lm(obs_randeff_umsam ~ SNV_obs_unsam  + 0, data = combine)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

smash_plot <- ggplot(combine, aes(x = SNV_obs_unsam, y = obs_randeff_umsam)) +
  geom_point() +
  labs(title = "SMASH", x = "Observed", y = "Predicted_SMASH") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +  
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(SNV_exp_unsam ~ SNV_obs_unsam + 0, data = combine)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

roulette_plot <- ggplot(combine, aes(x =SNV_obs_unsam , y = SNV_exp_unsam)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Observed", y = "Predicted_Roulette") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) + 
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(smash_plot, roulette_plot, ncol = 2)

```


## C - T

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_ct_1kb_pseudo05.rdata")

x_max <- max(combine$SNV_obs_unsam)
y_max <- max(combine$SNV_exp_unsam, combine$obs_randeff_umsam)

fit1 <- lm(obs_randeff_umsam ~ SNV_obs_unsam  + 0, data = combine)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

smash_plot <- ggplot(combine, aes(x = SNV_obs_unsam, y = obs_randeff_umsam)) +
  geom_point() +
  labs(title = "SMASH", x = "Observed", y = "Predicted_SMASH") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +  
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(SNV_exp_unsam ~ SNV_obs_unsam + 0, data = combine)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

roulette_plot <- ggplot(combine, aes(x =SNV_obs_unsam , y = SNV_exp_unsam)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Observed", y = "Predicted_Roulette") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) + 
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(smash_plot, roulette_plot, ncol = 2)

```


# 10kb

## All 6 mutation types
```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_10kb_pseudo05.rdata")

x_max <- max(combine$SNV_obs_unsam)
y_max <- max(combine$SNV_exp_unsam, combine$obs_randeff_umsam)

fit1 <- lm(obs_randeff_umsam ~ SNV_obs_unsam  + 0, data = combine)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

smash_plot <- ggplot(combine, aes(x = SNV_obs_unsam, y = obs_randeff_umsam)) +
  geom_point() +
  labs(title = "SMASH", x = "Observed", y = "Predicted_SMASH") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +  
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(SNV_exp_unsam ~ SNV_obs_unsam + 0, data = combine)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

roulette_plot <- ggplot(combine, aes(x =SNV_obs_unsam , y = SNV_exp_unsam)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Observed", y = "Predicted_Roulette") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) + 
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(smash_plot, roulette_plot, ncol = 2)

```


## A - G

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_ag_10kb_pseudo05.rdata")

x_max <- max(combine$SNV_obs_unsam)
y_max <- max(combine$SNV_exp_unsam, combine$obs_randeff_umsam)

fit1 <- lm(obs_randeff_umsam ~ SNV_obs_unsam  + 0, data = combine)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

smash_plot <- ggplot(combine, aes(x = SNV_obs_unsam, y = obs_randeff_umsam)) +
  geom_point() +
  labs(title = "SMASH", x = "Observed", y = "Predicted_SMASH") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +  
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(SNV_exp_unsam ~ SNV_obs_unsam + 0, data = combine)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

roulette_plot <- ggplot(combine, aes(x =SNV_obs_unsam , y = SNV_exp_unsam)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Observed", y = "Predicted_Roulette") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) + 
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(smash_plot, roulette_plot, ncol = 2)

```


## A - C

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_ac_10kb_pseudo05.rdata")

x_max <- max(combine$SNV_obs_unsam)
y_max <- max(combine$SNV_exp_unsam, combine$obs_randeff_umsam)

fit1 <- lm(obs_randeff_umsam ~ SNV_obs_unsam  + 0, data = combine)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

smash_plot <- ggplot(combine, aes(x = SNV_obs_unsam, y = obs_randeff_umsam)) +
  geom_point() +
  labs(title = "SMASH", x = "Observed", y = "Predicted_SMASH") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +  
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(SNV_exp_unsam ~ SNV_obs_unsam + 0, data = combine)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

roulette_plot <- ggplot(combine, aes(x =SNV_obs_unsam , y = SNV_exp_unsam)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Observed", y = "Predicted_Roulette") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) + 
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(smash_plot, roulette_plot, ncol = 2)
```

## A - T

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_at_10kb_pseudo05.rdata")

x_max <- max(combine$SNV_obs_unsam)
y_max <- max(combine$SNV_exp_unsam, combine$obs_randeff_umsam)

fit1 <- lm(obs_randeff_umsam ~ SNV_obs_unsam  + 0, data = combine)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

smash_plot <- ggplot(combine, aes(x = SNV_obs_unsam, y = obs_randeff_umsam)) +
  geom_point() +
  labs(title = "SMASH", x = "Observed", y = "Predicted_SMASH") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +  
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(SNV_exp_unsam ~ SNV_obs_unsam + 0, data = combine)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

roulette_plot <- ggplot(combine, aes(x =SNV_obs_unsam , y = SNV_exp_unsam)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Observed", y = "Predicted_Roulette") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) + 
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(smash_plot, roulette_plot, ncol = 2)

```

## C - A

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_ca_10kb_pseudo05.rdata")

x_max <- max(combine$SNV_obs_unsam)
y_max <- max(combine$SNV_exp_unsam, combine$obs_randeff_umsam)

fit1 <- lm(obs_randeff_umsam ~ SNV_obs_unsam  + 0, data = combine)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

smash_plot <- ggplot(combine, aes(x = SNV_obs_unsam, y = obs_randeff_umsam)) +
  geom_point() +
  labs(title = "SMASH", x = "Observed", y = "Predicted_SMASH") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +  
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(SNV_exp_unsam ~ SNV_obs_unsam + 0, data = combine)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

roulette_plot <- ggplot(combine, aes(x =SNV_obs_unsam , y = SNV_exp_unsam)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Observed", y = "Predicted_Roulette") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) + 
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(smash_plot, roulette_plot, ncol = 2)

```

## C - G
```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_cg_10kb_pseudo05.rdata")

x_max <- max(combine$SNV_obs_unsam)
y_max <- max(combine$SNV_exp_unsam, combine$obs_randeff_umsam)

fit1 <- lm(obs_randeff_umsam ~ SNV_obs_unsam  + 0, data = combine)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

smash_plot <- ggplot(combine, aes(x = SNV_obs_unsam, y = obs_randeff_umsam)) +
  geom_point() +
  labs(title = "SMASH", x = "Observed", y = "Predicted_SMASH") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +  
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(SNV_exp_unsam ~ SNV_obs_unsam + 0, data = combine)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

roulette_plot <- ggplot(combine, aes(x =SNV_obs_unsam , y = SNV_exp_unsam)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Observed", y = "Predicted_Roulette") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) + 
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(smash_plot, roulette_plot, ncol = 2)
```


## C - T

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_ct_10kb_pseudo05.rdata")

x_max <- max(combine$SNV_obs_unsam)
y_max <- max(combine$SNV_exp_unsam, combine$obs_randeff_umsam)

fit1 <- lm(obs_randeff_umsam ~ SNV_obs_unsam  + 0, data = combine)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

smash_plot <- ggplot(combine, aes(x = SNV_obs_unsam, y = obs_randeff_umsam)) +
  geom_point() +
  labs(title = "SMASH", x = "Observed", y = "Predicted_SMASH") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +  
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(SNV_exp_unsam ~ SNV_obs_unsam + 0, data = combine)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

roulette_plot <- ggplot(combine, aes(x =SNV_obs_unsam , y = SNV_exp_unsam)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Observed", y = "Predicted_Roulette") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) + 
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(smash_plot, roulette_plot, ncol = 2)

```

<!-- # 100bp -->

<!-- ## All 6 mutation types -->
<!-- ```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10} -->

<!-- load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_100bp.rdata") -->

<!-- # set up the 2x1 plot grid -->
<!-- par(mfrow=c(1,2)) -->

<!-- # plot the first figure -->
<!-- plot(combine$obs_randeff_umsam, combine$SNV_obs_unsam,  -->
<!--      xlab = "prediced_smash", ylab = "observed",  -->
<!--      main = "smash") -->
<!-- fit1 <- lm(combine$SNV_obs_unsam ~ combine$obs_randeff_umsam +0) -->
<!-- abline(fit1,col="red") -->
<!-- adj.rsq <-summary(fit1)$adj.r.squared -->
<!-- pval <- summary(fit1)$coefficients[,4] -->
<!-- print(paste0("SMASH: the ajusted rsq is ",adj.rsq)) -->
<!-- print(paste0("SMASH: the fitting pvalue is ",pval)) -->

<!-- # plot the second figure -->
<!-- plot(combine$SNV_exp_unsam, combine$SNV_obs_unsam,  -->
<!--      xlab = "predicted_Roulette", ylab = "observed", -->
<!--      main = "Roulette baseline") -->
<!-- fit2 <- lm(combine$SNV_obs_unsam ~ combine$SNV_exp_unsam +0) -->
<!-- abline(fit2,col="red") -->
<!-- adj.rsq <-summary(fit2)$adj.r.squared -->
<!-- pval <- summary(fit2)$coefficients[,4] -->
<!-- print(paste0("Roulette baseline: the ajusted rsq is ",adj.rsq)) -->
<!-- print(paste0("Roulette baseline: the fitting pvalue is ",pval)) -->

<!-- ``` -->


<!-- ## A - G -->

<!-- ```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10} -->

<!-- load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_ag_100bp.rdata") -->

<!-- # set up the 2x1 plot grid -->
<!-- par(mfrow=c(1,2)) -->

<!-- # plot the first figure -->
<!-- plot(combine$obs_randeff_umsam, combine$SNV_obs_unsam,  -->
<!--      xlab = "prediced_smash", ylab = "observed",  -->
<!--      main = "smash") -->
<!-- fit1 <- lm(combine$SNV_obs_unsam ~ combine$obs_randeff_umsam +0) -->
<!-- abline(fit1,col="red") -->
<!-- adj.rsq <-summary(fit1)$adj.r.squared -->
<!-- pval <- summary(fit1)$coefficients[,4] -->
<!-- print(paste0("SMASH: the ajusted rsq is ",adj.rsq)) -->
<!-- print(paste0("SMASH: the fitting pvalue is ",pval)) -->

<!-- # plot the second figure -->
<!-- plot(combine$SNV_exp_unsam, combine$SNV_obs_unsam,  -->
<!--      xlab = "predicted_Roulette", ylab = "observed", -->
<!--      main = "Roulette baseline") -->
<!-- fit2 <- lm(combine$SNV_obs_unsam ~ combine$SNV_exp_unsam +0) -->
<!-- abline(fit2,col="red") -->
<!-- adj.rsq <-summary(fit2)$adj.r.squared -->
<!-- pval <- summary(fit2)$coefficients[,4] -->
<!-- print(paste0("Roulette baseline: the ajusted rsq is ",adj.rsq)) -->
<!-- print(paste0("Roulette baseline: the fitting pvalue is ",pval)) -->

<!-- ``` -->


<!-- ## A - C -->

<!-- ```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10} -->

<!-- load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_ac_100bp.rdata") -->

<!-- # set up the 2x1 plot grid -->
<!-- par(mfrow=c(1,2)) -->

<!-- # plot the first figure -->
<!-- plot(combine$obs_randeff_umsam, combine$SNV_obs_unsam,  -->
<!--      xlab = "prediced_smash", ylab = "observed",  -->
<!--      main = "smash") -->
<!-- fit1 <- lm(combine$SNV_obs_unsam ~ combine$obs_randeff_umsam +0) -->
<!-- abline(fit1,col="red") -->
<!-- adj.rsq <-summary(fit1)$adj.r.squared -->
<!-- pval <- summary(fit1)$coefficients[,4] -->
<!-- print(paste0("SMASH: the ajusted rsq is ",adj.rsq)) -->
<!-- print(paste0("SMASH: the fitting pvalue is ",pval)) -->

<!-- # plot the second figure -->
<!-- plot(combine$SNV_exp_unsam, combine$SNV_obs_unsam,  -->
<!--      xlab = "predicted_Roulette", ylab = "observed", -->
<!--      main = "Roulette baseline") -->
<!-- fit2 <- lm(combine$SNV_obs_unsam ~ combine$SNV_exp_unsam +0) -->
<!-- abline(fit2,col="red") -->
<!-- adj.rsq <-summary(fit2)$adj.r.squared -->
<!-- pval <- summary(fit2)$coefficients[,4] -->
<!-- print(paste0("Roulette baseline: the ajusted rsq is ",adj.rsq)) -->
<!-- print(paste0("Roulette baseline: the fitting pvalue is ",pval)) -->

<!-- ``` -->

<!-- ## A - T -->

<!-- ```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10} -->

<!-- load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_at_100bp.rdata") -->

<!-- # set up the 2x1 plot grid -->
<!-- par(mfrow=c(1,2)) -->

<!-- # plot the first figure -->
<!-- plot(combine$obs_randeff_umsam, combine$SNV_obs_unsam,  -->
<!--      xlab = "prediced_smash", ylab = "observed",  -->
<!--      main = "smash") -->
<!-- fit1 <- lm(combine$SNV_obs_unsam ~ combine$obs_randeff_umsam +0) -->
<!-- abline(fit1,col="red") -->
<!-- adj.rsq <-summary(fit1)$adj.r.squared -->
<!-- pval <- summary(fit1)$coefficients[,4] -->
<!-- print(paste0("SMASH: the ajusted rsq is ",adj.rsq)) -->
<!-- print(paste0("SMASH: the fitting pvalue is ",pval)) -->

<!-- # plot the second figure -->
<!-- plot(combine$SNV_exp_unsam, combine$SNV_obs_unsam,  -->
<!--      xlab = "predicted_Roulette", ylab = "observed", -->
<!--      main = "Roulette baseline") -->
<!-- fit2 <- lm(combine$SNV_obs_unsam ~ combine$SNV_exp_unsam +0) -->
<!-- abline(fit2,col="red") -->
<!-- adj.rsq <-summary(fit2)$adj.r.squared -->
<!-- pval <- summary(fit2)$coefficients[,4] -->
<!-- print(paste0("Roulette baseline: the ajusted rsq is ",adj.rsq)) -->
<!-- print(paste0("Roulette baseline: the fitting pvalue is ",pval)) -->

<!-- ``` -->

<!-- ## C - A -->

<!-- ```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10} -->

<!-- load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_ca_100bp.rdata") -->

<!-- # set up the 2x1 plot grid -->
<!-- par(mfrow=c(1,2)) -->

<!-- # plot the first figure -->
<!-- plot(combine$obs_randeff_umsam, combine$SNV_obs_unsam,  -->
<!--      xlab = "prediced_smash", ylab = "observed",  -->
<!--      main = "smash") -->
<!-- fit1 <- lm(combine$SNV_obs_unsam ~ combine$obs_randeff_umsam +0) -->
<!-- abline(fit1,col="red") -->
<!-- adj.rsq <-summary(fit1)$adj.r.squared -->
<!-- pval <- summary(fit1)$coefficients[,4] -->
<!-- print(paste0("SMASH: the ajusted rsq is ",adj.rsq)) -->
<!-- print(paste0("SMASH: the fitting pvalue is ",pval)) -->

<!-- # plot the second figure -->
<!-- plot(combine$SNV_exp_unsam, combine$SNV_obs_unsam,  -->
<!--      xlab = "predicted_Roulette", ylab = "observed", -->
<!--      main = "Roulette baseline") -->
<!-- fit2 <- lm(combine$SNV_obs_unsam ~ combine$SNV_exp_unsam +0) -->
<!-- abline(fit2,col="red") -->
<!-- adj.rsq <-summary(fit2)$adj.r.squared -->
<!-- pval <- summary(fit2)$coefficients[,4] -->
<!-- print(paste0("Roulette baseline: the ajusted rsq is ",adj.rsq)) -->
<!-- print(paste0("Roulette baseline: the fitting pvalue is ",pval)) -->

<!-- ``` -->

<!-- ## C - G -->
<!-- ```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10} -->

<!-- load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_cg_100bp.rdata") -->

<!-- # set up the 2x1 plot grid -->
<!-- par(mfrow=c(1,2)) -->

<!-- # plot the first figure -->
<!-- plot(combine$obs_randeff_umsam, combine$SNV_obs_unsam,  -->
<!--      xlab = "prediced_smash", ylab = "observed",  -->
<!--      main = "smash") -->
<!-- fit1 <- lm(combine$SNV_obs_unsam ~ combine$obs_randeff_umsam +0) -->
<!-- abline(fit1,col="red") -->
<!-- adj.rsq <-summary(fit1)$adj.r.squared -->
<!-- pval <- summary(fit1)$coefficients[,4] -->
<!-- print(paste0("SMASH: the ajusted rsq is ",adj.rsq)) -->
<!-- print(paste0("SMASH: the fitting pvalue is ",pval)) -->

<!-- # plot the second figure -->
<!-- plot(combine$SNV_exp_unsam, combine$SNV_obs_unsam,  -->
<!--      xlab = "predicted_Roulette", ylab = "observed", -->
<!--      main = "Roulette baseline") -->
<!-- fit2 <- lm(combine$SNV_obs_unsam ~ combine$SNV_exp_unsam +0) -->
<!-- abline(fit2,col="red") -->
<!-- adj.rsq <-summary(fit2)$adj.r.squared -->
<!-- pval <- summary(fit2)$coefficients[,4] -->
<!-- print(paste0("Roulette baseline: the ajusted rsq is ",adj.rsq)) -->
<!-- print(paste0("Roulette baseline: the fitting pvalue is ",pval)) -->

<!-- ``` -->


<!-- ## C - T -->

<!-- ```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10} -->

<!-- load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_ct_100bp.rdata") -->

<!-- # set up the 2x1 plot grid -->
<!-- par(mfrow=c(1,2)) -->

<!-- # plot the first figure -->
<!-- plot(combine$obs_randeff_umsam, combine$SNV_obs_unsam,  -->
<!--      xlab = "prediced_smash", ylab = "observed",  -->
<!--      main = "smash") -->
<!-- fit1 <- lm(combine$SNV_obs_unsam ~ combine$obs_randeff_umsam +0) -->
<!-- abline(fit1,col="red") -->
<!-- adj.rsq <-summary(fit1)$adj.r.squared -->
<!-- pval <- summary(fit1)$coefficients[,4] -->
<!-- print(paste0("SMASH: the ajusted rsq is ",adj.rsq)) -->
<!-- print(paste0("SMASH: the fitting pvalue is ",pval)) -->

<!-- # plot the second figure -->
<!-- plot(combine$SNV_exp_unsam, combine$SNV_obs_unsam,  -->
<!--      xlab = "predicted_Roulette", ylab = "observed", -->
<!--      main = "Roulette baseline") -->
<!-- fit2 <- lm(combine$SNV_obs_unsam ~ combine$SNV_exp_unsam +0) -->
<!-- abline(fit2,col="red") -->
<!-- adj.rsq <-summary(fit2)$adj.r.squared -->
<!-- pval <- summary(fit2)$coefficients[,4] -->
<!-- print(paste0("Roulette baseline: the ajusted rsq is ",adj.rsq)) -->
<!-- print(paste0("Roulette baseline: the fitting pvalue is ",pval)) -->

<!-- ``` -->

