---
title: "Modelling the random effect with smash ---- first 90% base pairs in a window as training"
author: "XSun"
date: "2023-04-06"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
suppressMessages(library(ggplot2))
suppressMessages(library(gridExtra))
```


# Introduction

We grouped the chr22 into 100bp, 1kb, 10kb windows. In each window, we used the first 90% sites as training data, and the remaining 10% as the test_continuousing data. 

We

1. computed the random effect (eff_rand) in this window using these training sites. 

2. did log transform for eff_rand and got lg_eff_rand

3. smoothed the lg_eff_rand across all windows using [smashr](https://github.com/stephenslab/smashr), got mean_lg_eff_rand

4. eff_rand_sm <- exp(mean_lg_eff_rand)

5. predicted the observed number of SNV with eff_rand_sm on the test_continuousing sites in each window.

6. fit SNV_obs ~ SNV_obs_smash + 0 

For the windows with 0 SNVs: we replaced the 0s with a pseudocount of 0.5. 

# 1kb

## All 6 mutation types
```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_continuous_1kb_pseudo05.rdata")

x_max <- max(combine$SNV_exp_unsam, combine$obs_randeff_umsam)
y_max <- max(combine$SNV_obs_unsam)

fit1 <- lm(SNV_obs_unsam ~ obs_randeff_umsam + 0, data = combine)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

smash_plot <- ggplot(combine, aes(x = obs_randeff_umsam, y = SNV_obs_unsam)) +
  geom_point() +
  labs(title = "SMASH", x ="Predicted_SMASH" , y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(SNV_obs_unsam ~ SNV_exp_unsam+ 0, data = combine)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

roulette_plot <- ggplot(combine, aes(x = SNV_exp_unsam, y = SNV_obs_unsam)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Predicted_Roulette", y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(smash_plot, roulette_plot, ncol = 2)

combine$Mb <- combine$start_sam / 1000000
ggplot(combine, aes(x = Mb, y = SNV_obs_unsam)) +
  geom_point(size=0.1) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of SNVs in testing set/window -- SMASH") +
  geom_line(aes(x = Mb, y = obs_randeff_umsam), data = combine, color = "red",size=0.2) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

combine <- combine[combine$Mb >25 &combine$Mb <30, ]
ggplot(combine, aes(x = Mb, y = SNV_obs_unsam)) +
  geom_point(size=0.1) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of SNVs in testing set/window -- SMASH") +
  geom_line(aes(x = Mb, y = obs_randeff_umsam), data = combine, color = "red",size=0.2) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```


## A - G

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_continuous_ag_1kb_pseudo05.rdata")

x_max <- max(combine$SNV_exp_unsam, combine$obs_randeff_umsam)
y_max <- max(combine$SNV_obs_unsam)

fit1 <- lm(SNV_obs_unsam ~ obs_randeff_umsam + 0, data = combine)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

smash_plot <- ggplot(combine, aes(x = obs_randeff_umsam, y = SNV_obs_unsam)) +
  geom_point() +
  labs(title = "SMASH", x ="Predicted_SMASH" , y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(SNV_obs_unsam ~ SNV_exp_unsam+ 0, data = combine)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

roulette_plot <- ggplot(combine, aes(x = SNV_exp_unsam, y = SNV_obs_unsam)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Predicted_Roulette", y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(smash_plot, roulette_plot, ncol = 2)


combine$Mb <- combine$start_sam / 1000000
ggplot(combine, aes(x = Mb, y = SNV_obs_unsam)) +
  geom_point(size=0.1) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of SNVs in testing set/window -- SMASH") +
  geom_line(aes(x = Mb, y = obs_randeff_umsam), data = combine, color = "red",size=0.2) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

combine <- combine[combine$Mb >25 &combine$Mb <30, ]
ggplot(combine, aes(x = Mb, y = SNV_obs_unsam)) +
  geom_point(size=0.1) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of SNVs in testing set/window -- SMASH") +
  geom_line(aes(x = Mb, y = obs_randeff_umsam), data = combine, color = "red",size=0.2) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```


## A - C

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_continuous_ac_1kb_pseudo05.rdata")

x_max <- max(combine$SNV_exp_unsam, combine$obs_randeff_umsam)
y_max <- max(combine$SNV_obs_unsam)

fit1 <- lm(SNV_obs_unsam ~ obs_randeff_umsam + 0, data = combine)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

smash_plot <- ggplot(combine, aes(x = obs_randeff_umsam, y = SNV_obs_unsam)) +
  geom_point() +
  labs(title = "SMASH", x ="Predicted_SMASH" , y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(SNV_obs_unsam ~ SNV_exp_unsam+ 0, data = combine)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

roulette_plot <- ggplot(combine, aes(x = SNV_exp_unsam, y = SNV_obs_unsam)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Predicted_Roulette", y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(smash_plot, roulette_plot, ncol = 2)

combine$Mb <- combine$start_sam / 1000000
ggplot(combine, aes(x = Mb, y = SNV_obs_unsam)) +
  geom_point(size=0.1) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of SNVs in testing set/window -- SMASH") +
  geom_line(aes(x = Mb, y = obs_randeff_umsam), data = combine, color = "red",size=0.2) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

combine <- combine[combine$Mb >25 &combine$Mb <30, ]
ggplot(combine, aes(x = Mb, y = SNV_obs_unsam)) +
  geom_point(size=0.1) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of SNVs in testing set/window -- SMASH") +
  geom_line(aes(x = Mb, y = obs_randeff_umsam), data = combine, color = "red",size=0.2) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

```

## A - T

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_continuous_at_1kb_pseudo05.rdata")

x_max <- max(combine$SNV_exp_unsam, combine$obs_randeff_umsam)
y_max <- max(combine$SNV_obs_unsam)

fit1 <- lm(SNV_obs_unsam ~ obs_randeff_umsam + 0, data = combine)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

smash_plot <- ggplot(combine, aes(x = obs_randeff_umsam, y = SNV_obs_unsam)) +
  geom_point() +
  labs(title = "SMASH", x ="Predicted_SMASH" , y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(SNV_obs_unsam ~ SNV_exp_unsam+ 0, data = combine)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

roulette_plot <- ggplot(combine, aes(x = SNV_exp_unsam, y = SNV_obs_unsam)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Predicted_Roulette", y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(smash_plot, roulette_plot, ncol = 2)

combine$Mb <- combine$start_sam / 1000000
ggplot(combine, aes(x = Mb, y = SNV_obs_unsam)) +
  geom_point(size=0.1) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of SNVs in testing set/window -- SMASH") +
  geom_line(aes(x = Mb, y = obs_randeff_umsam), data = combine, color = "red",size=0.2) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

combine <- combine[combine$Mb >25 &combine$Mb <30, ]
ggplot(combine, aes(x = Mb, y = SNV_obs_unsam)) +
  geom_point(size=0.1) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of SNVs in testing set/window -- SMASH") +
  geom_line(aes(x = Mb, y = obs_randeff_umsam), data = combine, color = "red",size=0.2) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

```

## C - A

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_continuous_ca_1kb_pseudo05.rdata")

x_max <- max(combine$SNV_exp_unsam, combine$obs_randeff_umsam)
y_max <- max(combine$SNV_obs_unsam)

fit1 <- lm(SNV_obs_unsam ~ obs_randeff_umsam + 0, data = combine)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

smash_plot <- ggplot(combine, aes(x = obs_randeff_umsam, y = SNV_obs_unsam)) +
  geom_point() +
  labs(title = "SMASH", x ="Predicted_SMASH" , y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(SNV_obs_unsam ~ SNV_exp_unsam+ 0, data = combine)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

roulette_plot <- ggplot(combine, aes(x = SNV_exp_unsam, y = SNV_obs_unsam)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Predicted_Roulette", y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(smash_plot, roulette_plot, ncol = 2)

combine$Mb <- combine$start_sam / 1000000
ggplot(combine, aes(x = Mb, y = SNV_obs_unsam)) +
  geom_point(size=0.1) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of SNVs in testing set/window -- SMASH") +
  geom_line(aes(x = Mb, y = obs_randeff_umsam), data = combine, color = "red",size=0.2) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

combine <- combine[combine$Mb >25 &combine$Mb <30, ]
ggplot(combine, aes(x = Mb, y = SNV_obs_unsam)) +
  geom_point(size=0.1) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of SNVs in testing set/window -- SMASH") +
  geom_line(aes(x = Mb, y = obs_randeff_umsam), data = combine, color = "red",size=0.2) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```

## C - G
```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_continuous_cg_1kb_pseudo05.rdata")

x_max <- max(combine$SNV_exp_unsam, combine$obs_randeff_umsam)
y_max <- max(combine$SNV_obs_unsam)

fit1 <- lm(SNV_obs_unsam ~ obs_randeff_umsam + 0, data = combine)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

smash_plot <- ggplot(combine, aes(x = obs_randeff_umsam, y = SNV_obs_unsam)) +
  geom_point() +
  labs(title = "SMASH", x ="Predicted_SMASH" , y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(SNV_obs_unsam ~ SNV_exp_unsam+ 0, data = combine)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

roulette_plot <- ggplot(combine, aes(x = SNV_exp_unsam, y = SNV_obs_unsam)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Predicted_Roulette", y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(smash_plot, roulette_plot, ncol = 2)

combine$Mb <- combine$start_sam / 1000000
ggplot(combine, aes(x = Mb, y = SNV_obs_unsam)) +
  geom_point(size=0.1) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of SNVs in testing set/window -- SMASH") +
  geom_line(aes(x = Mb, y = obs_randeff_umsam), data = combine, color = "red",size=0.2) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

combine <- combine[combine$Mb >25 &combine$Mb <30, ]
ggplot(combine, aes(x = Mb, y = SNV_obs_unsam)) +
  geom_point(size=0.1) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of SNVs in testing set/window -- SMASH") +
  geom_line(aes(x = Mb, y = obs_randeff_umsam), data = combine, color = "red",size=0.2) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```


## C - T

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_continuous_ct_1kb_pseudo05.rdata")

x_max <- max(combine$SNV_exp_unsam, combine$obs_randeff_umsam)
y_max <- max(combine$SNV_obs_unsam)

fit1 <- lm(SNV_obs_unsam ~ obs_randeff_umsam + 0, data = combine)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

smash_plot <- ggplot(combine, aes(x = obs_randeff_umsam, y = SNV_obs_unsam)) +
  geom_point() +
  labs(title = "SMASH", x ="Predicted_SMASH" , y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(SNV_obs_unsam ~ SNV_exp_unsam+ 0, data = combine)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

roulette_plot <- ggplot(combine, aes(x = SNV_exp_unsam, y = SNV_obs_unsam)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Predicted_Roulette", y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(smash_plot, roulette_plot, ncol = 2)

combine$Mb <- combine$start_sam / 1000000
ggplot(combine, aes(x = Mb, y = SNV_obs_unsam)) +
  geom_point(size=0.1) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of SNVs in testing set/window -- SMASH") +
  geom_line(aes(x = Mb, y = obs_randeff_umsam), data = combine, color = "red",size=0.2) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

combine <- combine[combine$Mb >25 &combine$Mb <30, ]
ggplot(combine, aes(x = Mb, y = SNV_obs_unsam)) +
  geom_point(size=0.1) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of SNVs in testing set/window -- SMASH") +
  geom_line(aes(x = Mb, y = obs_randeff_umsam), data = combine, color = "red",size=0.2) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```


# 10kb

## All 6 mutation types
```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_continuous_10kb_pseudo05.rdata")

x_max <- max(combine$SNV_exp_unsam, combine$obs_randeff_umsam)
y_max <- max(combine$SNV_obs_unsam)

fit1 <- lm(SNV_obs_unsam ~ obs_randeff_umsam + 0, data = combine)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

smash_plot <- ggplot(combine, aes(x = obs_randeff_umsam, y = SNV_obs_unsam)) +
  geom_point() +
  labs(title = "SMASH", x ="Predicted_SMASH" , y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(SNV_obs_unsam ~ SNV_exp_unsam+ 0, data = combine)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

roulette_plot <- ggplot(combine, aes(x = SNV_exp_unsam, y = SNV_obs_unsam)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Predicted_Roulette", y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(smash_plot, roulette_plot, ncol = 2)

combine$Mb <- combine$start_sam / 1000000
ggplot(combine, aes(x = Mb, y = SNV_obs_unsam)) +
  geom_point(size=0.1) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of SNVs in testing set/window -- SMASH") +
  geom_line(aes(x = Mb, y = obs_randeff_umsam), data = combine, color = "red",size=0.2) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

combine <- combine[combine$Mb >25 &combine$Mb <30, ]
ggplot(combine, aes(x = Mb, y = SNV_obs_unsam)) +
  geom_point(size=0.1) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of SNVs in testing set/window -- SMASH") +
  geom_line(aes(x = Mb, y = obs_randeff_umsam), data = combine, color = "red",size=0.2) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

```


## A - G

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_continuous_ag_10kb_pseudo05.rdata")

x_max <- max(combine$SNV_exp_unsam, combine$obs_randeff_umsam)
y_max <- max(combine$SNV_obs_unsam)

fit1 <- lm(SNV_obs_unsam ~ obs_randeff_umsam + 0, data = combine)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

smash_plot <- ggplot(combine, aes(x = obs_randeff_umsam, y = SNV_obs_unsam)) +
  geom_point() +
  labs(title = "SMASH", x ="Predicted_SMASH" , y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(SNV_obs_unsam ~ SNV_exp_unsam+ 0, data = combine)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

roulette_plot <- ggplot(combine, aes(x = SNV_exp_unsam, y = SNV_obs_unsam)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Predicted_Roulette", y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(smash_plot, roulette_plot, ncol = 2)

combine$Mb <- combine$start_sam / 1000000
ggplot(combine, aes(x = Mb, y = SNV_obs_unsam)) +
  geom_point(size=0.1) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of SNVs in testing set/window -- SMASH") +
  geom_line(aes(x = Mb, y = obs_randeff_umsam), data = combine, color = "red",size=0.2) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

combine <- combine[combine$Mb >25 &combine$Mb <30, ]
ggplot(combine, aes(x = Mb, y = SNV_obs_unsam)) +
  geom_point(size=0.1) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of SNVs in testing set/window -- SMASH") +
  geom_line(aes(x = Mb, y = obs_randeff_umsam), data = combine, color = "red",size=0.2) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

```


## A - C

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_continuous_ac_10kb_pseudo05.rdata")

x_max <- max(combine$SNV_exp_unsam, combine$obs_randeff_umsam)
y_max <- max(combine$SNV_obs_unsam)

fit1 <- lm(SNV_obs_unsam ~ obs_randeff_umsam + 0, data = combine)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

smash_plot <- ggplot(combine, aes(x = obs_randeff_umsam, y = SNV_obs_unsam)) +
  geom_point() +
  labs(title = "SMASH", x ="Predicted_SMASH" , y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(SNV_obs_unsam ~ SNV_exp_unsam+ 0, data = combine)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

roulette_plot <- ggplot(combine, aes(x = SNV_exp_unsam, y = SNV_obs_unsam)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Predicted_Roulette", y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(smash_plot, roulette_plot, ncol = 2)

combine$Mb <- combine$start_sam / 1000000
ggplot(combine, aes(x = Mb, y = SNV_obs_unsam)) +
  geom_point(size=0.1) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of SNVs in testing set/window -- SMASH") +
  geom_line(aes(x = Mb, y = obs_randeff_umsam), data = combine, color = "red",size=0.2) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

combine <- combine[combine$Mb >25 &combine$Mb <30, ]
ggplot(combine, aes(x = Mb, y = SNV_obs_unsam)) +
  geom_point(size=0.1) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of SNVs in testing set/window -- SMASH") +
  geom_line(aes(x = Mb, y = obs_randeff_umsam), data = combine, color = "red",size=0.2) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```

## A - T

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_continuous_at_10kb_pseudo05.rdata")

x_max <- max(combine$SNV_exp_unsam, combine$obs_randeff_umsam)
y_max <- max(combine$SNV_obs_unsam)

fit1 <- lm(SNV_obs_unsam ~ obs_randeff_umsam + 0, data = combine)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

smash_plot <- ggplot(combine, aes(x = obs_randeff_umsam, y = SNV_obs_unsam)) +
  geom_point() +
  labs(title = "SMASH", x ="Predicted_SMASH" , y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(SNV_obs_unsam ~ SNV_exp_unsam+ 0, data = combine)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

roulette_plot <- ggplot(combine, aes(x = SNV_exp_unsam, y = SNV_obs_unsam)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Predicted_Roulette", y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(smash_plot, roulette_plot, ncol = 2)

combine$Mb <- combine$start_sam / 1000000
ggplot(combine, aes(x = Mb, y = SNV_obs_unsam)) +
  geom_point(size=0.1) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of SNVs in testing set/window -- SMASH") +
  geom_line(aes(x = Mb, y = obs_randeff_umsam), data = combine, color = "red",size=0.2) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

combine <- combine[combine$Mb >25 &combine$Mb <30, ]
ggplot(combine, aes(x = Mb, y = SNV_obs_unsam)) +
  geom_point(size=0.1) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of SNVs in testing set/window -- SMASH") +
  geom_line(aes(x = Mb, y = obs_randeff_umsam), data = combine, color = "red",size=0.2) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```

## C - A

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_continuous_ca_10kb_pseudo05.rdata")

x_max <- max(combine$SNV_exp_unsam, combine$obs_randeff_umsam)
y_max <- max(combine$SNV_obs_unsam)

fit1 <- lm(SNV_obs_unsam ~ obs_randeff_umsam + 0, data = combine)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

smash_plot <- ggplot(combine, aes(x = obs_randeff_umsam, y = SNV_obs_unsam)) +
  geom_point() +
  labs(title = "SMASH", x ="Predicted_SMASH" , y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(SNV_obs_unsam ~ SNV_exp_unsam+ 0, data = combine)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

roulette_plot <- ggplot(combine, aes(x = SNV_exp_unsam, y = SNV_obs_unsam)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Predicted_Roulette", y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(smash_plot, roulette_plot, ncol = 2)

combine$Mb <- combine$start_sam / 1000000
ggplot(combine, aes(x = Mb, y = SNV_obs_unsam)) +
  geom_point(size=0.1) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of SNVs in testing set/window -- SMASH") +
  geom_line(aes(x = Mb, y = obs_randeff_umsam), data = combine, color = "red",size=0.2) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

combine <- combine[combine$Mb >25 &combine$Mb <30, ]
ggplot(combine, aes(x = Mb, y = SNV_obs_unsam)) +
  geom_point(size=0.1) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of SNVs in testing set/window -- SMASH") +
  geom_line(aes(x = Mb, y = obs_randeff_umsam), data = combine, color = "red",size=0.2) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

```

## C - G
```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_continuous_cg_10kb_pseudo05.rdata")

x_max <- max(combine$SNV_exp_unsam, combine$obs_randeff_umsam)
y_max <- max(combine$SNV_obs_unsam)

fit1 <- lm(SNV_obs_unsam ~ obs_randeff_umsam + 0, data = combine)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

smash_plot <- ggplot(combine, aes(x = obs_randeff_umsam, y = SNV_obs_unsam)) +
  geom_point() +
  labs(title = "SMASH", x ="Predicted_SMASH" , y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(SNV_obs_unsam ~ SNV_exp_unsam+ 0, data = combine)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

roulette_plot <- ggplot(combine, aes(x = SNV_exp_unsam, y = SNV_obs_unsam)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Predicted_Roulette", y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(smash_plot, roulette_plot, ncol = 2)

combine$Mb <- combine$start_sam / 1000000
ggplot(combine, aes(x = Mb, y = SNV_obs_unsam)) +
  geom_point(size=0.1) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of SNVs in testing set/window -- SMASH") +
  geom_line(aes(x = Mb, y = obs_randeff_umsam), data = combine, color = "red",size=0.2) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

combine <- combine[combine$Mb >25 &combine$Mb <30, ]
ggplot(combine, aes(x = Mb, y = SNV_obs_unsam)) +
  geom_point(size=0.1) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of SNVs in testing set/window -- SMASH") +
  geom_line(aes(x = Mb, y = obs_randeff_umsam), data = combine, color = "red",size=0.2) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```


## C - T

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=5, fig.width=10}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_train_test_continuous_ct_10kb_pseudo05.rdata")

x_max <- max(combine$SNV_exp_unsam, combine$obs_randeff_umsam)
y_max <- max(combine$SNV_obs_unsam)

fit1 <- lm(SNV_obs_unsam ~ obs_randeff_umsam + 0, data = combine)
adj.rsq <- summary(fit1)$adj.r.squared
pval <- summary(fit1)$coefficients[,4]

smash_plot <- ggplot(combine, aes(x = obs_randeff_umsam, y = SNV_obs_unsam)) +
  geom_point() +
  labs(title = "SMASH", x ="Predicted_SMASH" , y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))


fit2 <- lm(SNV_obs_unsam ~ SNV_exp_unsam+ 0, data = combine)
adj.rsq <- summary(fit2)$adj.r.squared
pval <- summary(fit2)$coefficients[,4]

roulette_plot <- ggplot(combine, aes(x = SNV_exp_unsam, y = SNV_obs_unsam)) +
  geom_point() +
  labs(title = "Roulette Baseline", x = "Predicted_Roulette", y = "Observed") +
  geom_smooth(method = lm, formula = y ~ x + 0, color = "red") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw() +
  annotate("text", x = x_max, y = y_max, label = paste0("Adjusted R-Squared: ", round(adj.rsq, 2)), vjust = 1, hjust = 1)+
  ylim(0, y_max) +
  xlim(0, x_max) +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(smash_plot, roulette_plot, ncol = 2)

combine$Mb <- combine$start_sam / 1000000
ggplot(combine, aes(x = Mb, y = SNV_obs_unsam)) +
  geom_point(size=0.1) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of SNVs in testing set/window -- SMASH") +
  geom_line(aes(x = Mb, y = obs_randeff_umsam), data = combine, color = "red",size=0.2) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

combine <- combine[combine$Mb >25 &combine$Mb <30, ]
ggplot(combine, aes(x = Mb, y = SNV_obs_unsam)) +
  geom_point(size=0.1) +
  scale_x_continuous(name = "Genomic position (Mb)") +
  ylab("# of SNVs in testing set/window -- SMASH") +
  geom_line(aes(x = Mb, y = obs_randeff_umsam), data = combine, color = "red",size=0.2) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```



