---
title: "Comparing variance estimated by Delta method and numerical method"
author: "XSun"
date: "2023-04-13"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
suppressMessages(library(data.table))
suppressMessages(library(ggplot2))
```


# Introduction

We compare the variance estimated by delta method and numerical method here. In numerical method, we sampled 1000 times.

estimating the variance of the local random effects numerically: we can directly sample log ^ui many
times, and compute the variance. Specifically, we obtain the samples by:

- Sample from $Pois(y_{i})$, then divided by $T_i$. This is our sample of $\hat{\mu _{i}}$.

- Compute $log(\hat{\mu _{i}})$

After we obtain these samples, we can compute the variance of the sampled $log(\hat{\mu _{i}})$. We will need to
do this numerical computation for each window.


# 1kb

## All mutation types

```{r echo=F, message=FALSE, warning=FALSE,fig.height=5, fig.width=5}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_1kb_pseudo05_wholewindow_numerical_var_1000.rdata")

g <- function(x) log(x)
g_prime <- function(x) 1/x
data$randeff_var <- data$SNV_obs/(data$SNV_exp^2)
data$var_log_re_delta <- (g_prime(data$randeff)^2) * data$randeff_var

ggplot(data, aes(x = var_log_re_delta, y = var_log_randeff_sample)) +
  geom_point() +
  labs(x = "Variance estimated by Delta method", y = "Variance estimated by numerical method") +
  theme_bw() 


```

## A-G

```{r echo=F, message=FALSE, warning=FALSE,fig.height=5, fig.width=5}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_1kb_pseudo05_wholewindow_numerical_var_1000_ag.rdata")

g <- function(x) log(x)
g_prime <- function(x) 1/x
data$randeff_var <- data$SNV_obs/(data$SNV_exp^2)
data$var_log_re_delta <- (g_prime(data$randeff)^2) * data$randeff_var

ggplot(data, aes(x = var_log_re_delta, y = var_log_randeff_sample)) +
  geom_point() +
  labs(x = "Variance estimated by Delta method", y = "Variance estimated by numerical method") +
  theme_bw() 


```

## A-C

```{r echo=F, message=FALSE, warning=FALSE,fig.height=5, fig.width=5}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_1kb_pseudo05_wholewindow_numerical_var_1000_ac.rdata")

g <- function(x) log(x)
g_prime <- function(x) 1/x
data$randeff_var <- data$SNV_obs/(data$SNV_exp^2)
data$var_log_re_delta <- (g_prime(data$randeff)^2) * data$randeff_var

ggplot(data, aes(x = var_log_re_delta, y = var_log_randeff_sample)) +
  geom_point() +
  labs(x = "Variance estimated by Delta method", y = "Variance estimated by numerical method") +
  theme_bw() 


```

## A-T

```{r echo=F, message=FALSE, warning=FALSE,fig.height=5, fig.width=5}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_1kb_pseudo05_wholewindow_numerical_var_1000_at.rdata")

g <- function(x) log(x)
g_prime <- function(x) 1/x
data$randeff_var <- data$SNV_obs/(data$SNV_exp^2)
data$var_log_re_delta <- (g_prime(data$randeff)^2) * data$randeff_var

ggplot(data, aes(x = var_log_re_delta, y = var_log_randeff_sample)) +
  geom_point() +
  labs(x = "Variance estimated by Delta method", y = "Variance estimated by numerical method") +
  theme_bw() 


```

## C-A

```{r echo=F, message=FALSE, warning=FALSE,fig.height=5, fig.width=5}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_1kb_pseudo05_wholewindow_numerical_var_1000_ca.rdata")

g <- function(x) log(x)
g_prime <- function(x) 1/x
data$randeff_var <- data$SNV_obs/(data$SNV_exp^2)
data$var_log_re_delta <- (g_prime(data$randeff)^2) * data$randeff_var

ggplot(data, aes(x = var_log_re_delta, y = var_log_randeff_sample)) +
  geom_point() +
  labs(x = "Variance estimated by Delta method", y = "Variance estimated by numerical method") +
  theme_bw() 


```

## C-G

```{r echo=F, message=FALSE, warning=FALSE,fig.height=5, fig.width=5}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_1kb_pseudo05_wholewindow_numerical_var_1000_cg.rdata")

g <- function(x) log(x)
g_prime <- function(x) 1/x
data$randeff_var <- data$SNV_obs/(data$SNV_exp^2)
data$var_log_re_delta <- (g_prime(data$randeff)^2) * data$randeff_var

ggplot(data, aes(x = var_log_re_delta, y = var_log_randeff_sample)) +
  geom_point() +
  labs(x = "Variance estimated by Delta method", y = "Variance estimated by numerical method") +
  theme_bw() 


```

## C-T

```{r echo=F, message=FALSE, warning=FALSE,fig.height=5, fig.width=5}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_1kb_pseudo05_wholewindow_numerical_var_1000_ct.rdata")

g <- function(x) log(x)
g_prime <- function(x) 1/x
data$randeff_var <- data$SNV_obs/(data$SNV_exp^2)
data$var_log_re_delta <- (g_prime(data$randeff)^2) * data$randeff_var

ggplot(data, aes(x = var_log_re_delta, y = var_log_randeff_sample)) +
  geom_point() +
  labs(x = "Variance estimated by Delta method", y = "Variance estimated by numerical method") +
  theme_bw() 


```

# 10kb

## All mutation types

```{r echo=F, message=FALSE, warning=FALSE,fig.height=5, fig.width=5}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_10kb_pseudo05_wholewindow_numerical_var_1000.rdata")

g <- function(x) log(x)
g_prime <- function(x) 1/x
data$randeff_var <- data$SNV_obs/(data$SNV_exp^2)
data$var_log_re_delta <- (g_prime(data$randeff)^2) * data$randeff_var

ggplot(data, aes(x = var_log_re_delta, y = var_log_randeff_sample)) +
  geom_point() +
  labs(x = "Variance estimated by Delta method", y = "Variance estimated by numerical method") +
  theme_bw() 


```

## A-G

```{r echo=F, message=FALSE, warning=FALSE,fig.height=5, fig.width=5}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_10kb_pseudo05_wholewindow_numerical_var_1000_ag.rdata")

g <- function(x) log(x)
g_prime <- function(x) 1/x
data$randeff_var <- data$SNV_obs/(data$SNV_exp^2)
data$var_log_re_delta <- (g_prime(data$randeff)^2) * data$randeff_var

ggplot(data, aes(x = var_log_re_delta, y = var_log_randeff_sample)) +
  geom_point() +
  labs(x = "Variance estimated by Delta method", y = "Variance estimated by numerical method") +
  theme_bw() 


```

## A-C

```{r echo=F, message=FALSE, warning=FALSE,fig.height=5, fig.width=5}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_10kb_pseudo05_wholewindow_numerical_var_1000_ac.rdata")

g <- function(x) log(x)
g_prime <- function(x) 1/x
data$randeff_var <- data$SNV_obs/(data$SNV_exp^2)
data$var_log_re_delta <- (g_prime(data$randeff)^2) * data$randeff_var

ggplot(data, aes(x = var_log_re_delta, y = var_log_randeff_sample)) +
  geom_point() +
  labs(x = "Variance estimated by Delta method", y = "Variance estimated by numerical method") +
  theme_bw() 


```

## A-T

```{r echo=F, message=FALSE, warning=FALSE,fig.height=5, fig.width=5}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_10kb_pseudo05_wholewindow_numerical_var_1000_at.rdata")

g <- function(x) log(x)
g_prime <- function(x) 1/x
data$randeff_var <- data$SNV_obs/(data$SNV_exp^2)
data$var_log_re_delta <- (g_prime(data$randeff)^2) * data$randeff_var

ggplot(data, aes(x = var_log_re_delta, y = var_log_randeff_sample)) +
  geom_point() +
  labs(x = "Variance estimated by Delta method", y = "Variance estimated by numerical method") +
  theme_bw() 


```

## C-A

```{r echo=F, message=FALSE, warning=FALSE,fig.height=5, fig.width=5}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_10kb_pseudo05_wholewindow_numerical_var_1000_ca.rdata")

g <- function(x) log(x)
g_prime <- function(x) 1/x
data$randeff_var <- data$SNV_obs/(data$SNV_exp^2)
data$var_log_re_delta <- (g_prime(data$randeff)^2) * data$randeff_var

ggplot(data, aes(x = var_log_re_delta, y = var_log_randeff_sample)) +
  geom_point() +
  labs(x = "Variance estimated by Delta method", y = "Variance estimated by numerical method") +
  theme_bw() 


```

## C-G

```{r echo=F, message=FALSE, warning=FALSE,fig.height=5, fig.width=5}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_10kb_pseudo05_wholewindow_numerical_var_1000_cg.rdata")

g <- function(x) log(x)
g_prime <- function(x) 1/x
data$randeff_var <- data$SNV_obs/(data$SNV_exp^2)
data$var_log_re_delta <- (g_prime(data$randeff)^2) * data$randeff_var

ggplot(data, aes(x = var_log_re_delta, y = var_log_randeff_sample)) +
  geom_point() +
  labs(x = "Variance estimated by Delta method", y = "Variance estimated by numerical method") +
  theme_bw() 


```

## C-T

```{r echo=F, message=FALSE, warning=FALSE,fig.height=5, fig.width=5}

load("/project/xinhe/xsun/mutation_rate/1.exp_obs_gnomad/data_processed/chr22/smash_10kb_pseudo05_wholewindow_numerical_var_1000_ct.rdata")

g <- function(x) log(x)
g_prime <- function(x) 1/x
data$randeff_var <- data$SNV_obs/(data$SNV_exp^2)
data$var_log_re_delta <- (g_prime(data$randeff)^2) * data$randeff_var

ggplot(data, aes(x = var_log_re_delta, y = var_log_randeff_sample)) +
  geom_point() +
  labs(x = "Variance estimated by Delta method", y = "Variance estimated by numerical method") +
  theme_bw() 


```